### 2019-10-27 tdabench-paper/tdabench-paper.Rmd
Skeleton of RJournal submission

---
title: Comparing efficiency of R packages implementing calculation of persistent homology
author:
  - name: Author One
    affiliation: Affiliation
    address:
    - line 1
    - line 2
    email:  author1@work
  - name: Author Two
    affiliation: Affiliation
    address:
    - line 1
    - line 2
    email:  author2@work
abstract: >
  An abstract of less than 150 words.
output:
  rticles::rjournal_article:
    includes:
      in_header: preamble.tex
---

## Introduction

Introductory section which may include references in parentheses
\citep{R}, or cite a reference such as \citet{R} in the text.

## Methods

This section may contain a figure such as Figure \ref{figure:rlogo}.

\begin{figure}[htbp]
  \centering
  \includegraphics{Rlogo}
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}

## Results

There will likely be several sections, perhaps including code snippets, such as:

```{r}
x <- 1:10
x
```

## Discussion

This file is only a basic article template. For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
\bibliography{RJreferences}

### 2019-10-27 tdabench-paper/tdabench-paper.Rmd
Added tentative figure descriptions

@@
 
-There will likely be several sections, perhaps including code snippets, such as:
+Figure 1: Comparison of persistent homology calculation runtimes across R packages for a torus.
+Panel 2.
+Horizontal axis: number of points in torus.
+Vertical axis: runtime.
+Point chart.
+Point color: underlying engine.
+Point shape: same as color.
+Panel 1.
+Torus showing users visual representation of dataset/point cloud.
+Goal: overview showing speed differences across engines on a canonical shape.
 
-```{r}
-x <- 1:10
-x
-```
+Figure 2: Comparison of persistent homology calculation runtimes across R packages for n-dimensional boxes (n = 2, 3, 4, 5).
+For each dimension, features up to n-1 dimensions are calculated.
+Panel 1.
+Visual representation of 2-d box (maybe points, maybe box, maybe scattered both).
+Panel 2.
+Visual representation of 3-d box (maybe points, maybe box, maybe both scattered).
+Panels 3-6 (differing n={2,3,4,5} for dimension).
+Point shape/color: underlying engine.
+Horizontal axis: number of points.
+Vertical axis: runtime.
+
+Figure 3: Runtime of persistent homology calculation as a function of underlying engine (4d noisy circle/annulus).
+Panel 1: visualize noisy circle - use imagination for 4d.
+Panel 2.
+Horizontal axis: number of points.
+Vertical axis: runtime.
+Point color/shape: underlying engine.
+Goal: show that TDAstats and GUDHIalpha are sufficiently faster than GUDHI and Dionysus, so the latter pair can be ignored for rest of figures.
+
+Figure 4: Runtime of persistent homology calculation as a function of underlying engine and feature dimension (4d noisy circle/annulus).
+Panel 1 = GUDHI alpha engine.
+Horizontal axis = number of points.
+Vertical axis = runtime.
+Point color/shape = feature dimension.
+Panel 2 = panel 1, but with TDAstats engine.
+Goal: show how increasing dimension slows down the calculation time.
+
+Figure 5: Memory use of Rips complex vs alpha complex.
+Engine = GUDHI(alpha).
+Panel 1 = 4-sphere.
+Panel 2 = 3-annulus.
+Panel 3 = torus.
+Panel 4 = 5-box.
+For each panel: point color/shape = engine (GUDHI vs alpha); horizontal axis = number of points; vertical axis = memory use; feature dimension = data dimension - 1 for all panels.
+Goal: compare memory
+POINT OUT THAT RIPSER DOESN'T MEASURE MEMORY B/C NO BOUNDARY MATRIX, BUT LIKELY SOMEWHERE IN BETWEEN THE TWO
+
+Figure 6: Same as Figure 5, but TDAstats vs GUDHIalpha for runtime.
 

### 2019-10-27 FullTDABench.R
Writes output time/mem to file

@@ -119,6 +119,10 @@ times.all <- mapply(TDA_bench, vars.all$measure, vars.all$data.type,
                       vars.all$num.iteration)
 
 vars.all$time <- times.all
+
+# write to file
+write.csv(vars.all, file = "time-output.csv",
+          row.names = FALSE)
 ##Finish measuring all times
 
 
@@ -132,24 +136,19 @@ memory.all <- mapply(TDA_bench, mvars.all$measure, mvars.all$data.type,
 mvars.all$memory <- memory.all
 ##Finish measuring all memories
 
+# write to file
+write.csv(mvars.all, file = "mem-output.csv",
+          row.names = FALSE)
 
 #testmvars
-mvars.test.circle <- as_tibble(expand.grid(measure = "memory", data.type = "circle",
-                                      data.dimensions = 2:4, num.points = seq(50, 150, 50),
-                                      feature.dimensions = 1, 
-                                      TDA.library = c("GUDHI", "GUDHIalpha")
-)) %>% subset(feature.dimensions < data.dimensions)
+#mvars.test.circle <- as_tibble(expand.grid(measure = "memory", data.type = "circle",
+#                                      data.dimensions = 2:4, num.points = seq(50, 150, 50),
+#                                      feature.dimensions = 1, 
+#                                      TDA.library = c("GUDHI", "GUDHIalpha")
+#)) %>% subset(feature.dimensions < data.dimensions)
 
-memory.test <- mapply(TDA_bench, mvars.test.circle$measure, mvars.test.circle$data.type,
-                      mvars.test.circle$data.dimensions, mvars.test.circle$num.points,
-                      mvars.test.circle$feature.dimensions, mvars.test.circle$TDA.library)
+#memory.test <- mapply(TDA_bench, mvars.test.circle$measure, mvars.test.circle$data.type,
+#                      mvars.test.circle$data.dimensions, mvars.test.circle$num.points,
+#                      mvars.test.circle$feature.dimensions, mvars.test.circle$TDA.library)
 
-mvars.test.circle$memory <- memory.test
-
-
-
-
-
-
-
-
+#mvars.test.circle$memory <- memory.test

### 2019-12-14 CCF_Time.R
commented out if there were same runs in data, ignore previous runs

@@ -100,15 +100,15 @@ vars.all <- vars.all[!(vars.all$data.dimensions > 3
                        & vars.all$TDA.library == "GUDHIalpha"), ]
 
 #Reads in collected data so far
-vars.tested <- read.csv("timeccf.csv")
-colnames(vars.tested) <- c("measure", "data.type", "data.dimensions",
-                           "num.points", "feature.dimensions", "TDA.library", "time1",
-                           "time2", "time3", "time4", "time5", "time6", "time7", "time8",
-                           "time9", "time10")
+#vars.tested <- read.csv("timeccf.csv")
+#colnames(vars.tested) <- c("measure", "data.type", "data.dimensions",
+#                           "num.points", "feature.dimensions", "TDA.library", "time1",
+ #                          "time2", "time3", "time4", "time5", "time6", "time7", "time8",
+ #                          "time9", "time10")
 
 #From the variables, removes what has already been collected from each session
-vars.tested$feature.dimensions <- as.integer(vars.tested$feature.dimensions)
-vars.all <- vars.all %>% anti_join(vars.tested)
+#vars.tested$feature.dimensions <- as.integer(vars.tested$feature.dimensions)
+#vars.all <- vars.all %>% anti_join(vars.tested)
 
 mapply(TDA_bench, vars.all$measure, vars.all$data.type,
        vars.all$data.dimensions, vars.all$num.points,

### 2019-12-14 Functions.R
For benchmarking, removed the enclosing radius, resulted in highly variable results (mainly because the max enclosing radius in a given point cloud can vary so widely). However, if a point cloud has a small max enclosing radius by chance (even if it's a big point cloud), the run time reduces greatly. Will discuss this separately. In the mean time, set all Rips complex (Including stats) to have max radius of 2 as that is the maximum enclosing radius among all the data points (torus r1=1 r2=1)

@@ -116,14 +116,7 @@ noisycircle <- function(num.points, data.dimensions,
   return(to.calc.hom)
 }
 
-#####Function to make enclosing radius for TDA GUDHI and Dionysus (makes runtime fairer)#####
-enclosing_radius <- function(X){
-  # function which finds radius beyond which no homology changes
-  # X is a point cloud data frame
-  d = dist(X)
-  n = nrow(X)
-  return(min(unlist(lapply(X = 1:(nrow(X) - 1), FUN = function(X){ return(max(d[(1+(X-1)*n-(X-1)*X/2):(X*n-X*(X+1)/2)])) }))))
-}
+
 
 #####bench time#####
 # NB: maxscale = 5 is used in TDA package examples, so used here (no default)
@@ -133,19 +126,16 @@ enclosing_radius <- function(X){
 bench <- function(pointdata, TDA.library, featdim, num.iterations) {
   # TDAstats
   if (TDA.library == "stats") {
-    enc.rad <- enclosing_radius(pointdata)
-    time <- mark(calculate_homology(pointdata, dim = featdim, threshold = enc.rad),
+    time <- mark(calculate_homology(pointdata, dim = featdim, threshold = 2),
                  iterations = num.iterations)
     # TDA - Dionysus
   } else if (TDA.library == "Dionysus") {
-    enc.rad <- enclosing_radius(pointdata)
-    time <- mark(ripsDiag(pointdata, maxdimension = featdim, maxscale = enc.rad,
+    time <- mark(ripsDiag(pointdata, maxdimension = featdim, maxscale = 2,
                           location = FALSE, library = "Dionysus"),
                  iterations = num.iterations)
     # TDA - GUDHI
   } else if (TDA.library == "GUDHI") {
-    enc.rad <- enclosing_radius(pointdata)
-    time <- mark(ripsDiag(pointdata, maxdimension = featdim, maxscale = enc.rad,
+    time <- mark(ripsDiag(pointdata, maxdimension = featdim, maxscale = 2,
                           location = FALSE, library = "GUDHI"),
                  iterations = num.iterations)
     # TDA - GUDHI (alpha complex)

### 2019-12-14 Enclosing_Radius_Improvement.R
File to analyze effect of specifying the enclosing radius first

@@ -1,0 +1,10 @@
+#####Function to make enclosing radius for TDA GUDHI and Dionysus (makes runtime fairer)#####
+enclosing_radius <- function(X){
+  # function which finds radius beyond which no homology changes
+  # X is a point cloud data frame
+  d = dist(X)
+  n = nrow(X)
+  return(min(unlist(lapply(X = 1:(nrow(X) - 1), FUN = function(X){ return(max(d[(1+(X-1)*n-(X-1)*X/2):(X*n-X*(X+1)/2)])) }))))
+}
+
+

### 2019-12-14 Enclosing_Radius_Improvement.R
Have a plan as to how to discuss this

@@ -1,3 +1,14 @@
+library(plyr)
+library(readr)
+library(dplyr)
+library(TDA)
+library(TDAstats)
+library(bench)
+library(pryr)
+library(ggplot2)
+
+source("Functions.R")
+
 #####Function to make enclosing radius for TDA GUDHI and Dionysus (makes runtime fairer)#####
 enclosing_radius <- function(X){
   # function which finds radius beyond which no homology changes
@@ -7,4 +18,77 @@ enclosing_radius <- function(X){
   return(min(unlist(lapply(X = 1:(nrow(X) - 1), FUN = function(X){ return(max(d[(1+(X-1)*n-(X-1)*X/2):(X*n-X*(X+1)/2)])) }))))
 }
 
+#####alternate bench function
+bench.enc <- function(pointdata, TDA.library, featdim, num.iterations) {
+  rad <- enclosing_radius(pointdata)
+  str.lib <- paste(TDA.library)
+  # TDAstats
+  if (TDA.library == "stats") {
+    time <- mark(calculate_homology(pointdata, dim = featdim, threshold = rad),
+                 iterations = num.iterations)
+    # TDA - Dionysus
+  } else if (TDA.library == "Dionysus") {
+    time <- mark(ripsDiag(pointdata, maxdimension = featdim, maxscale = rad,
+                          location = FALSE, library = "Dionysus"),
+                 iterations = num.iterations)
+    # TDA - GUDHI
+  } else if (TDA.library == "GUDHI") {
+    time <- mark(ripsDiag(pointdata, maxdimension = featdim, maxscale = rad,
+                          location = FALSE, library = "GUDHI"),
+                 iterations = num.iterations)
+    # TDA - GUDHI (alpha complex)
+  } else if (TDA.library == "GUDHIalpha") {
+    time <- mark(alphaComplexDiag(pointdata, maxdimension = featdim,
+                                  location = FALSE, library = "GUDHI"),
+                 iterations = num.iterations)
+    # none of the above
+  } else {
+    stop("Invalid TDA engine selected")
+  }
+  
+  # row 1 column 12 is list of times
+  time.list <- time[1,12]
+  time.runs <- unlist(time.list[[1]])
+  appnd <- tibble()
+  row <- c(rad, time.runs, str.lib)
+  row.appnd <- rbind(appnd, row)
+  colnames(row.appnd) <- seq(1:ncol(row.appnd))
+  return(row.appnd)
+}
+
+
+list <- replicate(100, unifcircle(2, 2), simplify = FALSE)
[…]

### 2019-12-14 Functions.R
Actual max scale is 4

@@ -126,16 +126,16 @@ noisycircle <- function(num.points, data.dimensions,
 bench <- function(pointdata, TDA.library, featdim, num.iterations) {
   # TDAstats
   if (TDA.library == "stats") {
-    time <- mark(calculate_homology(pointdata, dim = featdim, threshold = 2),
+    time <- mark(calculate_homology(pointdata, dim = featdim, threshold = 4),
                  iterations = num.iterations)
     # TDA - Dionysus
   } else if (TDA.library == "Dionysus") {
-    time <- mark(ripsDiag(pointdata, maxdimension = featdim, maxscale = 2,
+    time <- mark(ripsDiag(pointdata, maxdimension = featdim, maxscale = 4,
                           location = FALSE, library = "Dionysus"),
                  iterations = num.iterations)
     # TDA - GUDHI
   } else if (TDA.library == "GUDHI") {
-    time <- mark(ripsDiag(pointdata, maxdimension = featdim, maxscale = 2,
+    time <- mark(ripsDiag(pointdata, maxdimension = featdim, maxscale = 4,
                           location = FALSE, library = "GUDHI"),
                  iterations = num.iterations)
     # TDA - GUDHI (alpha complex)

### 2019-12-16 CCF_Time.R
from lab mac

@@ -28,8 +28,8 @@ vars.noisycircle <- as_tibble(expand.grid(measure = "time", data.type = "annulus
 
 #Making Box Variables for time
 vars.box <- as_tibble(expand.grid(measure = "time", data.type = "uniform",
-                                  data.dimensions = 2:10, num.points = seq(10, 500, 5),
-                                  feature.dimensions = 1:9, 
+                                  data.dimensions = 2:8, num.points = seq(10, 500, 5),
+                                  feature.dimensions = 1:7, 
                                   TDA.library = c("stats", "Dionysus", "GUDHI", "GUDHIalpha"),
                                   num.iteration = 10, file.name = "timeccf.csv")) %>% subset(feature.dimensions < data.dimensions)
 
@@ -100,15 +100,15 @@ vars.all <- vars.all[!(vars.all$data.dimensions > 3
                        & vars.all$TDA.library == "GUDHIalpha"), ]
 
 #Reads in collected data so far
-#vars.tested <- read.csv("timeccf.csv")
-#colnames(vars.tested) <- c("measure", "data.type", "data.dimensions",
-#                           "num.points", "feature.dimensions", "TDA.library", "time1",
- #                          "time2", "time3", "time4", "time5", "time6", "time7", "time8",
- #                          "time9", "time10")
+vars.tested <- read.csv("timeccf.csv")
+colnames(vars.tested) <- c("measure", "data.type", "data.dimensions",
+                           "num.points", "feature.dimensions", "TDA.library", "time1",
+                          "time2", "time3", "time4", "time5", "time6", "time7", "time8",
+                          "time9", "time10")
 
 #From the variables, removes what has already been collected from each session
-#vars.tested$feature.dimensions <- as.integer(vars.tested$feature.dimensions)
-#vars.all <- vars.all %>% anti_join(vars.tested)
+vars.tested$feature.dimensions <- as.integer(vars.tested$feature.dimensions)
+vars.all <- vars.all %>% anti_join(vars.tested)
 
 mapply(TDA_bench, vars.all$measure, vars.all$data.type,
        vars.all$data.dimensions, vars.all$num.points,

### 2019-12-16 Cluster_Practice.R
Added to git ignore and cluster script

@@ -1,0 +1,343 @@
+####calling libraries, scrips, and making function####
+library(plyr)
+library(readr)
+library(dplyr)
+library(TDA)
+library(TDAstats)
+library(bench)
+library(pryr)
+
+# add if statements to check for parameter validity
+
+#####BOX DATA#####
+# generates an n-dimension box (0, 1) of uniformly distributed points
+unifbox <- function(num.points, data.dimensions) {
+  # empty 
+  to.calc.hom <- matrix(NA, nrow = num.points, ncol = data.dimensions)
+  
+  # adds a column of randomaly generated points 0 to 1 for how many ever dimensions specified
+  for (i in 1:data.dimensions) {
+    col <- runif(num.points, 0, 1)
+    to.calc.hom[, i] <- col
+  }
+  
+  return(to.calc.hom)
+}
+
+#####TORUS DATA#####
+# very simple, uses native uniform torus function from TDA
+torus <- function(num.points) {
+  torusUnif(num.points, 1, 1)
+}
+
+#####UNIFORM CIRCLE DATA#####
+# Uses the sphere picking tactic to make uniform distribution 
+# Cite Marsaglia paper from Wolfram Alpha page
+unifcircle <- function(num.points, data.dimensions) {
+  # var that stores result (empty df setup)
+  to.calc.hom <- matrix(NA, nrow = num.points, ncol = data.dimensions)
+  #to.calc.hom <- as.data.frame(to.calc.hom)
+  
+  # returns 2-d circle data
+  if (data.dimensions == 2) {
+    angles <- runif(num.points, 0, 2*pi)
+    to.calc.hom <- cbind(cos(angles), sin(angles))
+  }
+  
+  # returns 3-d circle data
+  if (data.dimensions == 3) {
+    
+    # each loop generates one row of data
+    for (curr_row in 1:num.points) {
+      
+      # generate valid x1 and x2
+      x1 <- runif(1, -1, 1)
+      x2 <- runif(1, -1, 1)
+      while (x1 ^ 2 + x2 ^ 2 >= 1) {
+        x1 <- runif(1, -1, 1)
+        x2 <- runif(1, -1, 1)
+      }
+      
+      # generate coordinates of sphere
+      x <- 2 * x1 * sqrt(1 - x1 ^ 2 - x2 ^ 2)
+      y <- 2 * x2 * sqrt(1 - x1 ^ 2 - x2 ^ 2)
[…]

### 2019-12-16 tdabench-paper/tdabench.Rmd
Updated paper title/authors

--- 2019-10-27 tdabench-paper/tdabench-paper.Rmd
@@
 ---
-title: Comparing efficiency of R packages implementing calculation of persistent homology
+title: Comparing R packages for calculation of persistent homology
 author:
-  - name: Author One
+  - name: Eashwar V. Somasundaram
     affiliation: Affiliation
@@
     email:  author1@work
-  - name: Author Two
+  - name: Shael E. Brown
     affiliation: Affiliation
@@
     email:  author2@work
+  - name: Adam Litzler
+    affiliation: Affiliation
+    address:
+    - line 1
+    - line 2
+    email:  author1@work
+  - name: Raoul R. Wadhwa
+    affiliation: Affiliation
+    address:
+    - line 1
+    - line 2
+    email:  author1@work
+  - name: Jacob G. Scott
+    affiliation: Affiliation
+    address:
+    - line 1
+    - line 2
+    email:  author1@work
 abstract: >
@@
 For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
-\bibliography{RJreferences}
+\bibliography{tdabench-paper}

### 2019-12-16 tdabench-paper/tdabench.Rmd
Add most affiliations

--- 2019-10-27 tdabench-paper/tdabench-paper.Rmd
@@
 ---
-title: Comparing efficiency of R packages implementing calculation of persistent homology
+title: Comparing R packages for calculation of persistent homology
 author:
-  - name: Author One
+  - name: Eashwar V. Somasundaram
+    affiliation: School of Medicine
+    address:
+    - Case Western Reserve University
+    - Cleveland, OH 44016
+    email: eashwar.somasundaram@case.edu
+  - name: Shael E. Brown
     affiliation: Affiliation
@@
     - line 2
-    email:  author1@work
-  - name: Author Two
-    affiliation: Affiliation
+    email: shael.brown@mail.mcgill.ca
+  - name: Adam Litzler
+    affiliation: Department of Translational Hematology and Oncology Research
     address:
-    - line 1
-    - line 2
-    email:  author2@work
+    - Cleveland Clinic Foundation
+    - Cleveland, OH 44195
+    email:  author@work
+  - name: Raoul R. Wadhwa
+    affiliation: Cleveland Clinic Lerner College of Medicine
+    address:
+    - Case Western Reserve University
+    - Cleveland, OH 44195
+    email: raoul.wadhwa@case.edu
+  - name: Jacob G. Scott
+    affiliation: Department of Translational Hematology and Oncology Research
+    address:
+    - Cleveland Clinic Foundation
+    - Cleveland, OH 44195
+    email: ScottJ10@ccf.org
 abstract: >
@@
 Introductory section which may include references in parentheses
-\citep{R}, or cite a reference such as \citet{R} in the text.
+\citep{Rlang}, or cite a reference such as \citet{Rlang} in the text.
 
@@
 For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
-\bibliography{RJreferences}
+\bibliography{tdabench}

### 2019-12-16 tdabench-paper/tdabench.Rmd
edited affiliations

--- 2019-10-27 tdabench-paper/tdabench-paper.Rmd
@@
 ---
-title: Comparing efficiency of R packages implementing calculation of persistent homology
+title: Comparing R packages for calculation of persistent homology
 author:
-  - name: Author One
+  - name: Eashwar V. Somasundaram
+    affiliation: School of Medicine
+    address:
+    - Case Western Reserve University
+    - Cleveland, OH 44016, United States
+    email: eashwar.somasundaram@case.edu
+  - name: Shael E. Brown
     affiliation: Affiliation
@@
     - line 2
-    email:  author1@work
-  - name: Author Two
-    affiliation: Affiliation
+    email: shael.brown@mail.mcgill.ca
+  - name: Adam Litzler
+    affiliation: Department of Translational Hematology and Oncology Research
     address:
-    - line 1
-    - line 2
-    email:  author2@work
+    - Cleveland Clinic Foundation
+    - Cleveland, OH 44195, United States
+    email:  author@work
+  - name: Raoul R. Wadhwa
+    affiliation: Cleveland Clinic Lerner College of Medicine
+    address:
+    - Case Western Reserve University
+    - Cleveland, OH 44195, United States
+    email: raoul.wadhwa@case.edu
+  - name: Jacob G. Scott
+    affiliation: Department of Translational Hematology and Oncology Research
+    address:
+    - Cleveland Clinic Foundation
+    - Cleveland, OH 44195, United States
+    email: ScottJ10@ccf.org
 abstract: >

### 2019-12-16 tdabench-paper/tdabench.Rmd
added methods section

@@
 
-This section may contain a figure such as Figure \ref{figure:rlogo}.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics{Rlogo}
-  \caption{The logo of R.}
-  \label{figure:rlogo}
-\end{figure}
+An exclusively R (v3.6.2) code pipeline was built for this project \citep{Rlang,ihaka1996}.
+We use \texttt{readr} v1.3.1 to read rectangular data \cite{readr}, \texttt{ggplot2} v3.2.1 \citep{ggplot2} and \texttt{magick} v2.2 \citep{magick} to visualize data, \texttt{bench} v1.0.4 to collect benchmark data \citep{bench}, \texttt{TDA} v1.6.9 \citep{TDA} and \texttt{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \texttt{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
+Runtime calculations were averaged over 10 iterations and are visualized as mean $\pm$ SD.
+Datasets were generated by sampling functions in base R to generate points uniformly distributed over circles (dimension = 2), spheres (dimension = 3), filled squares (dimension = 2), filled cubes (dimension = 3, 4), and tori (dimension = 3).
+The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after pragmatic considerations of available computational resources.
+For consistency, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the TDA and TDAstats R packages.
+Within the TDA package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the TDAstats package, benchmark data was collected for the Ripser \citep{Ripser} library.
+As alpha complex calculation was only implemented in GUDHI, alpha complex benchmark data was naturally only collected for the single library.
+Measuring memory usage proved challenging since all the libraries calculating persistent homology were implemented in another programming language and then wrapped in R.
+Thus, memory burden was indirectly measured by using boundary matrix size as a proxy.
+Given that Ripser optimizes persistent homology calculation by avoiding calculation of a boundary matrix, memory use benchmarks are not provided for Ripser and, consequently, TDAstats.
+Fully reproducible code for all numerical results and figures in this report can be found at https://github.com/eashwarsoma/TDA-benchmark.
 

### 2019-12-16 tdabench-paper/tdabench.Rmd
corrected format for CRAN packages

@@
 
-Introductory section which may include references in parentheses
-\citep{Rlang}, or cite a reference such as \citet{Rlang} in the text.
-
 ## Methods
 
-An exclusively R (v3.6.2) code pipeline was built for this project \citep{Rlang,ihaka1996}.
-We use \texttt{readr} v1.3.1 to read rectangular data \cite{readr}, \texttt{ggplot2} v3.2.1 \citep{ggplot2} and \texttt{magick} v2.2 \citep{magick} to visualize data, \texttt{bench} v1.0.4 to collect benchmark data \citep{bench}, \texttt{TDA} v1.6.9 \citep{TDA} and \texttt{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \texttt{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
+<!--An exclusively R (v3.6.2) code pipeline was built for this project \citep{Rlang,ihaka1996}.-->
+We use \CRANpkg{readr} v1.3.1 to read rectangular data \cite{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2} and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
 Runtime calculations were averaged over 10 iterations and are visualized as mean $\pm$ SD.
@@
 The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after pragmatic considerations of available computational resources.
-For consistency, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the TDA and TDAstats R packages.
-Within the TDA package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the TDAstats package, benchmark data was collected for the Ripser \citep{Ripser} library.
+For consistency, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} R packages.
+Within the \CRANpkg{TDA} package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the \CRANpkg{TDAstats} package, benchmark data was collected for the Ripser \citep{Ripser} library.
 As alpha complex calculation was only implemented in GUDHI, alpha complex benchmark data was naturally only collected for the single library.
@@
 Thus, memory burden was indirectly measured by using boundary matrix size as a proxy.
-Given that Ripser optimizes persistent homology calculation by avoiding calculation of a boundary matrix, memory use benchmarks are not provided for Ripser and, consequently, TDAstats.
+Given that Ripser optimizes persistent homology calculation by avoiding calculation of a boundary matrix, memory use benchmarks are not provided for Ripser and, consequently, \CRANpkg{TDAstats}.
 Fully reproducible code for all numerical results and figures in this report can be found at https://github.com/eashwarsoma/TDA-benchmark.
@@
 
-This file is only a basic article template.
-For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
 \bibliography{tdabench}

### 2019-12-16 Cluster_TDA_bench.R
Run this code for cluster

@@ -333,8 +333,6 @@ format.data <- do.call(rbind, all.data)
 
 write.csv(format.data, "cluster.csv")
 
-
-
 format.data
 
 

### 2019-12-17 tdabench-paper/tdabench.Rmd
Added temporary placeholder for discussion
Added affiliation

@@
   - name: Shael E. Brown
-    affiliation: Affiliation
+    affiliation: Department of Quantitative Life Sciences
     address:
-    - line 1
-    - line 2
+    - McGill University
+    - Montreal, QC H3A 0G4, Canada
     email: shael.brown@mail.mcgill.ca
@@
 
+Need to make a better version of this figure, but essential idea remains the same (recommendation for which R package to use depending on what you want to do).
+
+\begin{figure}
+  \centering
+  \includegraphics[height=2.5in]{decision}
+\end{figure}
+
 \bibliography{tdabench}

### 2020-01-11 tdabench-paper/tdabench.Rmd
updated author order

@@
     email:  author@work
+  - name: Jacob G. Scott
+    affiliation: Department of Translational Hematology and Oncology Research
+    address:
+    - Cleveland Clinic Foundation
+    - Cleveland, OH 44195, United States
+    email: ScottJ10@ccf.org
   - name: Raoul R. Wadhwa
@@
     email: raoul.wadhwa@case.edu
-  - name: Jacob G. Scott
-    affiliation: Department of Translational Hematology and Oncology Research
-    address:
-    - Cleveland Clinic Foundation
-    - Cleveland, OH 44195, United States
-    email: ScottJ10@ccf.org
 abstract: >

### 2020-01-11 tdabench-paper/tdabench.Rmd
added visualization package citations

@@
 <!--An exclusively R (v3.6.2) code pipeline was built for this project \citep{Rlang,ihaka1996}.-->
-We use \CRANpkg{readr} v1.3.1 to read rectangular data \cite{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2} and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
+We use \CRANpkg{readr} v1.3.1 to read rectangular data \cite{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2}, \CRANpkg{scatterplot3d} v0.3-41 \citep{scatterplot3d}, \CRANpkg{recexcavAAR} v0.3.0 \citep{recexcavAAR}, and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
 Runtime calculations were averaged over 10 iterations and are visualized as mean $\pm$ SD.

### 2020-01-11 tdabench-paper/tdabench.Rmd
fixed zip code

@@
     - Case Western Reserve University
-    - Cleveland, OH 44016, United States
+    - Cleveland, OH 44106, United States
     email: eashwar.somasundaram@case.edu

### 2020-01-11 tdabench-paper/tdabench.Rmd
grammar fix

@@
 Datasets were generated by sampling functions in base R to generate points uniformly distributed over circles (dimension = 2), spheres (dimension = 3), filled squares (dimension = 2), filled cubes (dimension = 3, 4), and tori (dimension = 3).
-The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after pragmatic considerations of available computational resources.
+The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after considering available computational resources.
 For consistency, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} R packages.

### 2020-01-11 tdabench-paper/tdabench.Rmd
made sentence more specific

@@
 The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after considering available computational resources.
-For consistency, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} R packages.
+For consistency between software libraries, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} R packages.
 Within the \CRANpkg{TDA} package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the \CRANpkg{TDAstats} package, benchmark data was collected for the Ripser \citep{Ripser} library.

### 2020-01-11 tdabench-paper/tdabench.Rmd
fixed double adverb
added results structure

@@
     - Cleveland, OH 44195, United States
-    email:  author@work
+    email:  adli4611@colorado.edu
   - name: Jacob G. Scott
@@
 Within the \CRANpkg{TDA} package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the \CRANpkg{TDAstats} package, benchmark data was collected for the Ripser \citep{Ripser} library.
-As alpha complex calculation was only implemented in GUDHI, alpha complex benchmark data was naturally only collected for the single library.
+As alpha complex calculation was only implemented in GUDHI, alpha complex benchmark data was only collected for the single library.
 Measuring memory usage proved challenging since all the libraries calculating persistent homology were implemented in another programming language and then wrapped in R.
@@
 ## Results
+
+### Runtime comparison between packages for calculation of V-R complex
+
+#### Prototype/Canonical dataset (Torus)
+
+#### Runtime as a function of data dimension (feature = n - 1)
+
+#### Runtime as a function of feature dimension
+
+### Comparison of alpha and Vietoris-Rips complexes within GUDHI
+
+#### Runtime as a function of data dimension (const feature)
+
+#### Memory benchmarks
+
+### Computational Limitations (advice/guidance for users)
+
+Include Ripser doesn't calcualte boundary matrix (in comparison of GUDHI rips vs. alpha)
 

### 2020-01-11 Regression_analysis.R
Completed Regression for all Figures

@@ -1,0 +1,154 @@
+library(dplyr)
+library(ggplot2)
+library(reshape)
+library(ggforce)
+library(knitr)
+
+source("./Figures/Creating_ggplot_Figures.R")
+
+
+#Create Rips Power Reg Function
+#Returns coefficient, exponent, and R^2 from log log linear regression
+reg.fit.rip <- function (data, strat, y, x) {
+  #subset the data
+  data.sub <- split(data, data[strat])
+  data.sub <- data.sub[lapply(data.sub, nrow) != 0]
+  
+  #lin.regression to have a list of m and b predict values
+  fmla.lin <- as.formula(paste("log(", y, ")"," ", "~", " ", "log(", x, ")", sep=""))
+  lin.list <- lapply(data.sub, lm, formula = fmla.lin)
+  #return(lin.list)
+  #Extract the relevant coefficients
+  lin.coef <- lapply(lin.list, function(z) c(z$coefficients, summary(z)$r.squared))
+  #return(lin.coef)
+  #Return a nice table in the for
+  bval <- unlist(lapply(lin.coef, function(z) exp(z[1])))
+  zval <- unlist(lapply(lin.coef, function(z) z[2]))
+  rval <- unlist(lapply(lin.coef, function(z) z[3]))
+  vals <- cbind(coef = bval, exp = zval, `R^2` = rval)
+  row.names(vals) <- names(data.sub)
+  return(vals)
+}
+
+#Create Alpha Linear Reg Function
+#Returns coefficient, exponent, and R^2 from linear regression
+reg.fit.alp <- function (data, strat, y, x) {
+  #subset the data
+  data.sub <- split(data, data[strat])
+  data.sub <- data.sub[lapply(data.sub, nrow) != 0]
+  
+  #lin.regression to have a list of m and b predict values
+  fmla.lin <- as.formula(paste(y," ", "~", " ", x, sep=""))
+  lin.list <- lapply(data.sub, lm, formula = fmla.lin)
+  #return(lin.list)
+  #Extract the relevant coefficients
+  lin.coef <- lapply(lin.list, function(z) c(z$coefficients, summary(z)$r.squared))
+  #return(lin.coef)
+  #Return a nice table in the for
+  bval <- unlist(lapply(lin.coef, function(z) z[1]))
+  zval <- unlist(lapply(lin.coef, function(z) z[2]))
[…]

### 2020-01-11 Figures/Creating_ggplot_Figures.R
Added Rips Complex Intro Figure

@@ -564,11 +564,96 @@ ggsave("./Figures/Unrasterized_Images/fig6mac.png", plot = fig.6.mac,
        dpi = 400, limitsize = TRUE)
 
 
+####Intro Figures####
+####Rips Complex
+library("ggtda")
+
+n <- 5;
+set.seed(3)
+d <- data.frame(
+  x = runif(n, 0, .8),
+  y = runif(n, 0, .8)
+)
+# compute the persistent homology
+ph <- as.data.frame(TDAstats::calculate_homology(as.matrix(d), dim = 1))
+print(head(ph, n = 12))
+ph <- transform(ph, dim = as.factor(dimension))
+
+prox <- .25
+p_d1 <- ggplot(d, aes(x = x, y = y)) +
+  theme_classic() +
+  coord_fixed() +
+  stat_disk(radius = prox/2, fill = "aquamarine3") +
+  geom_point() +theme(axis.title = element_blank(),
+                      axis.text = element_blank(),
+                      axis.ticks = element_blank(),
+                      axis.line=element_blank())
+p_sc1 <- ggplot(d, aes(x = x, y = y)) +
+  theme_classic() +
+  coord_fixed() +
+  stat_vietoris2(diameter = prox, fill = "darkgoldenrod") +
+  stat_vietoris1(diameter = prox, alpha = .25) +
+  stat_vietoris0() + theme(axis.title = element_blank(),
+                            axis.text = element_blank(),
+                            axis.ticks = element_blank(),
+                            axis.line=element_blank())
+
+prox <- .5
+p_d2 <- ggplot(d, aes(x = x, y = y)) +
+  theme_classic() +
+  coord_fixed() +
+  stat_disk(radius = prox/2, fill = "aquamarine3") +
+  geom_point() +theme(axis.title = element_blank(),
+                      axis.text = element_blank(),
+                      axis.ticks = element_blank(),
+                      axis.line=element_blank())
+p_sc2 <- ggplot(d, aes(x = x, y = y)) +
+  theme_classic() +
+  coord_fixed() +
+  stat_vietoris2(diameter = prox, fill = "darkgoldenrod") +
+  stat_vietoris1(diameter = prox, alpha = .25) +
+  stat_vietoris0() + theme(axis.title = element_blank(),
+                           axis.text = element_blank(),
+                           axis.ticks = element_blank(),
+                           axis.line=element_blank())
+
+prox <- .75
[…]

### 2020-01-11 Figures/Creating_ggplot_Figures.R
Added code for alpha complex

@@ -636,9 +636,7 @@ p_sc3 <- ggplot(d, aes(x = x, y = y)) +
                            axis.ticks = element_blank(),
                            axis.line=element_blank())
 
-# combine the plots
-
-
+# combine and save the plots
 png(filename = "./Figures/Final_Figures/Intro_Rips.png",
     width = 6, height = 4, units = "in", res = 450)
 
@@ -653,19 +651,117 @@ gridExtra::grid.arrange(
 
 dev.off()
 
-####Finish here!
+####Alpha Complex
+library(deldir)
+library(ggplot2)
+
+df <- d
+
+#This creates the voronoi line segments
+voronoi <- deldir(d$x, d$y)
+voronoi$delsgs %>% mutate(
+  distance = sqrt((x1-x2)^2+(y1-y2)^2),
+)  -> voronoi$delsgs
 
 
+#Now we can make a plot
+alp.1 <- ggplot(data=df, aes(x=x,y=y)) +
+  #Plot the voronoi lines
+  geom_segment(
+    aes(x = x1, y = y1, xend = x2, yend = y2),
+    size = .5,
+    data = voronoi$dirsgs,
+    linetype = 1,
+    color= "#FFB958") + 
+  #Plot the points
+  geom_point(
+    fill=rgb(70,130,180,255,maxColorValue=255),
+    pch=21,
+    size = 4,
+    color="#333333") +
+  #Plot the voronoi lines
+  geom_segment(
+    aes(x = x1, y = y1, xend = x2, yend = y2),
+    size = 1,
+    data = subset(voronoi$delsgs, distance <= .25),
+    linetype = 1,
+    color= "black") + 
+  geom_circle(data= df, aes(x0=x,y0=y, r = .25/2), fill = "aquamarine", alpha = .35, inherit.aes = FALSE, color = NA) + 
+  coord_fixed() + theme_classic() + theme(axis.title = element_blank(),
+                        axis.text = element_blank(),
+                        axis.ticks = element_blank(),
+                        axis.line=element_blank()) 
 
+alp.1
 
+alp.2 <- ggplot(data=df, aes(x=x,y=y)) +
+  #Plot the voronoi lines
+  geom_segment(
+    aes(x = x1, y = y1, xend = x2, yend = y2),
+    size = .5,
+    data = voronoi$dirsgs,
+    linetype = 1,
+    color= "#FFB958") + 
+  #Plot the points
+  geom_point(
[…]

### 2020-01-12 Regression_analysis.R
added kable function to make latex code

@@ -66,6 +66,15 @@ fig.2.reg.rip <- reg.fit.rip(data = data.fig.2,
                          strat = c("library", "point.cloud.dim"),
                          y = "avg.time",
                          x = "num.points")
+rownames(fig.2.reg.rip) <- c("TDAstats Circle",
+                             "Dionysus Circle",
+                             "GUDHI Circle",
+                             "TDAstats Sphere",
+                             "Dionysus Sphere",
+                             "GUDHI Sphere",
+                             "TDAstats Hypersphere",
+                             "Dionysus Hypersphere",
+                             "GUDHI Hypersphere")
 kable(fig.2.reg.rip)
 
 #Figure 3 (#exclude, data set is too small to tell reliable)
@@ -75,11 +84,17 @@ fig.4.reg.rip <- reg.fit.rip(data = subset(data.fig.4, library == "GUDHI Rips"),
                          strat = "point.cloud.dim",
                          y = "avg.time",
                          x = "num.points")
+rownames(fig.4.reg.rip) <- c("2-annulus",
+                             "3-annulus",
+                             "4-annulus")
 
 fig.4.reg.alp <- reg.fit.alp(data = subset(data.fig.4, library == "GUDHI Alpha"),
                              strat = "point.cloud.dim",
                              y = "avg.time",
                              x = "num.points")
+rownames(fig.4.reg.alp) <- c("2-annulus",
+                             "3-annulus")
+
 kable(fig.4.reg.rip, digits = 9)
 kable(fig.4.reg.alp)
 
@@ -111,6 +126,13 @@ fig.5c.reg.alp <- reg.fit.alp(data = data.fig.5c,
                           strat = c("point.cloud", "point.cloud.dim"),
                           y = "memory",
                           x = "num.points")
+rownames(fig.5c.reg.alp) <- c("2-annulus",
+                             "Circle",
+                             "2-box",
+                             "3-annulus",
+                             "sphere",
+                             "torus",
+                             "3-box")
 
 kable(fig.5c.reg.alp)
 
@@ -152,3 +174,5 @@ names(all.reg.list) <- c("fig.1.reg.rip",
                            "fig.5c.reg.alp",
                            "fig.6.reg.rip",
                            "fig.6.reg.alp")
+
+lapply(all.reg.list, kable, format = "latex")

### 2020-01-16 Figures/Creating_ggplot_Figures.R
Edited code and intro alpha complex figure to look much better

@@ -666,35 +666,12 @@ voronoi$delsgs %>% mutate(
 
 #Now we can make a plot
 alp.1 <- ggplot(data=df, aes(x=x,y=y)) +
-  #Plot the voronoi lines
-  geom_segment(
-    aes(x = x1, y = y1, xend = x2, yend = y2),
-    size = .5,
-    data = voronoi$dirsgs,
-    linetype = 1,
-    color= "#FFB958") + 
-  #Plot the points
-  geom_point(
-    fill=rgb(70,130,180,255,maxColorValue=255),
-    pch=21,
-    size = 4,
-    color="#333333") +
-  #Plot the voronoi lines
-  geom_segment(
-    aes(x = x1, y = y1, xend = x2, yend = y2),
-    size = 1,
-    data = subset(voronoi$delsgs, distance <= .25),
-    linetype = 1,
-    color= "black") + 
-  geom_circle(data= df, aes(x0=x,y0=y, r = .25/2), fill = "aquamarine", alpha = .35, inherit.aes = FALSE, color = NA) + 
+  #Plot the union balls
+  geom_circle(data= df, aes(x0=x,y0=y, r = .25/2), fill = "aquamarine", alpha = 1.0, inherit.aes = FALSE, color = NA) + 
   coord_fixed() + theme_classic() + theme(axis.title = element_blank(),
                         axis.text = element_blank(),
                         axis.ticks = element_blank(),
-                        axis.line=element_blank()) 
-
-alp.1
-
-alp.2 <- ggplot(data=df, aes(x=x,y=y)) +
+                        axis.line=element_blank()) +
   #Plot the voronoi lines
   geom_segment(
     aes(x = x1, y = y1, xend = x2, yend = y2),
@@ -702,28 +679,27 @@ alp.2 <- ggplot(data=df, aes(x=x,y=y)) +
     data = voronoi$dirsgs,
     linetype = 1,
     color= "#FFB958") + 
-  #Plot the points
-  geom_point(
-    fill=rgb(70,130,180,255,maxColorValue=255),
-    pch=21,
-    size = 4,
-    color="#333333") +
-  #Plot the voronoi lines
-  geom_segment(
-    aes(x = x1, y = y1, xend = x2, yend = y2),
-    size = 1,
-    data = subset(voronoi$delsgs, distance <= .35),
-    linetype = 1,
-    color= "black") + 
[…]

### 2020-01-17 Figures/Rasterize_Figures.R
Reran figure code, compiled fig 5a,5b,5c into one fig5

@@ -156,6 +156,28 @@ image_write(fig4, path = './Figures/Final_Figures/fig4.png', format = 'png')
 fig5b <- image_read("./Figures/Unrasterized_Images/fig5b.png") 
 image_write(fig5b, path = './Figures/Final_Figures/fig5b.png', format = 'png')
 
+####Combine Figure 5 into one grid
+library(png)
+library(grid)
+library(gridExtra)
+
+plot1 <- readPNG('./Figures/Final_Figures/fig5a.png')
+plot2 <- readPNG('./Figures/Final_Figures/fig5b.png')
+plot3 <- readPNG('./Figures/Final_Figures/fig5c.png')
+
+png("./Figures/Final_Figures/fig5t.png", width = 6, height = 4, 
+    units = "in", res = 450)
+grid.arrange(rasterGrob(plot1, interpolate = T),
+             rasterGrob(plot2, interpolate = T), 
+             rasterGrob(plot3, interpolate = T), 
+             layout_matrix = rbind(c(1,2), c(3, 2)),
+             heights=c(2,2, 1))
+dev.off()
+
+#Crop white space
+fig5 <- image_read("./Figures/Final_Figures/fig5t.png") 
+fig5 <- image_crop(fig5, "2700x1400", gravity = "North")
+image_write(fig5, path = './Figures/Final_Figures/fig5.png', format = 'png')
 
 ####Mac Data Figure 1####
 #Read in Figure

### 2020-01-17 tdabench-paper/tdabench.Rmd
Added figures and results text

--- 2020-01-11 tdabench-paper/tdabench.Rmd
@@
     - Case Western Reserve University
-    - Cleveland, OH 44106, United States
+    - Cleveland, OH 44016, United States
     email: eashwar.somasundaram@case.edu
@@
 
-## Introduction
+\section{Introduction}
 
-## Methods
+\section{Mathematics of Persistent Homology}
 
+\begin{figure}
+  \centering
+  \includegraphics[height=2.5in]{../Figures/Final_Figures/Intro_Rips.png}
+  \caption{As $\delta$ continues to grow, more circles intersect filling out the simplicial complex.
+Features on the simplicial complex appear and die as $\delta$ increases.
+Eventually, the full convex hull is drawn ending the "filtration" process.}
+\end{figure}
+\begin{figure}
+  \centering
+  \includegraphics[height=2.5in]{../Figures/Final_Figures/Intro_Alpha.png}
+  \caption{As $\alpha$ continues to grow, more circles fill out the Voronoi cells.
+Once two circles in adjacent Voronoi cells fill out the cell, a connecting segment can be drawn.
+Eventually, once $\alpha$ is large enough, the Delaunay Triangulation is formed.}
+\end{figure}
+
+\section{Methods}
 <!--An exclusively R (v3.6.2) code pipeline was built for this project \citep{Rlang,ihaka1996}.-->
-We use \CRANpkg{readr} v1.3.1 to read rectangular data \cite{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2}, \CRANpkg{scatterplot3d} v0.3-41 \citep{scatterplot3d}, \CRANpkg{recexcavAAR} v0.3.0 \citep{recexcavAAR}, and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
+We use \CRANpkg{readr} v1.3.1 to read rectangular data \cite{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2} and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
 Runtime calculations were averaged over 10 iterations and are visualized as mean $\pm$ SD.
 Datasets were generated by sampling functions in base R to generate points uniformly distributed over circles (dimension = 2), spheres (dimension = 3), filled squares (dimension = 2), filled cubes (dimension = 3, 4), and tori (dimension = 3).
-The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after considering available computational resources.
+The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after pragmatic considerations of available computational resources.
 For consistency, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} R packages.
@@
 
-## Results
+\section{Results}
 
-Figure 1: Comparison of persistent homology calculation runtimes across R packages for a torus.
+\subsection{Runtime comparison between packages for calculation of V-R complex}
+In our analysis, we first compare the three Rips complexes libraries among each other. 
+We examine the run time performance of persistence homology calculation under different conditions. 
+We then focus our analysis on comparing Rips complexes with Alpha complexes using the GUDHI package.
+In all runtime calculations, 10 iterations were performed and data are presented as mean +/- standard deviation.
+
+\subsubsection{Prototype/Canonical dataset (Torus)}
+The torus point cloud is a classical shape used to demonstrate persistent homology calculations. 
+A persistent homology calculation of a uniformly distributed torus would reveal a large 2 dimensional feature (void) in the middle representing the “hole” of the torus. 
+In figure \ref{fig:tor}, we calculated up to 2 dimensional features of a torus 3d point cloud for an increasing number of points across all TDA libraries.
+
+\begin{figure}
+  \centering
+  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig1.png}
+  \caption{Comparison of Rips Complex persistent homology calculation runtimes across R packages for a torus.
+The x-axis represents the number of points in a 3D torus point cloud, and y-axis represents average run time.
+For these given parameters, the Rips libraries show a polynomial growth curve for increasing point dimension.
+Further details about regression parameters is found in supplemental material.}
+\label{fig:tor}
+\end{figure}
+
+\subsubsection{Runtime as a function of data dimension (feature = n - 1)}
+From the torus comparison, we see a clear difference in performance between the three Rips libraries. 
+We wanted to see if this performance difference changed with more complex point clouds. 
+Each panel in figure \ref{fig:cir} corresponds to an n-dimensional circle (circle, sphere, and hypersphere). 
+For each point cloud, n-1 dimensional features were calculated.
+
+A similar growth pattern was seen among the TDA libraries. 
+As the sphere dimensions and feature dimensions increased, the execution time growth becomes apparent sooner. 
+At these higher dimensional features, some of the curves seem to cut off. 
+This is due to certain point cloud sizes not being computable on our devices. 
+These limitations are laid out in the "Computation Limits" section.
+
+\begin{figure}
+  \centering
+  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig2.png}
+  \caption{Runtime of persistent homology calculation as a function of underlying engine for different dimension circles.
+Panel 1.
+Comparing persistent homology calculation run times for a 2D circle.
 Panel 2.
-Horizontal axis: number of points in torus.
-Vertical axis: runtime.
-Point chart.
-Point color: underlying engine.
-Point shape: same as color.
+Comparing persistent homology calculation run times for a 3D sphere.
+Panel 3.
+Comparing persistent homology calculation run times for a 4D hypersphere.
+Maximum feature dimensions were calculated for each point cloud.}
+\label{fig:cir}
+\end{figure}
+
+\subsubsection{Runtime as a function of feature dimension}
+For an n-dimensional point cloud, up to n-1 feature dimensions can be calculated. 
+In the field, most work has only included calculations up to 1 or 2 feature dimensions. 
+Perhaps this due to difficulty of conceptualizing what very high dimensional features represent. 
+High dimensional feature calculations are also very taxing computationally. 
+For Figure \ref{fig:box}, we restricted the calculations to very small point clouds due to computational limits.
+
+As feature dimension increased, Rips complexes show a significant rise in run time with increasing feature dimension within an 8 dimensional point cloud. 
+With this small sample set, it is not clear whether the growth is exponential or polynomial.
+
+\begin{figure}
+  \centering
+  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig3.png}
+  \caption{Comparison of Rips Complex persistent homology calculation as a function of feature dimension.
 Panel 1.
-Torus showing users visual representation of dataset/point cloud.
-Goal: overview showing speed differences across engines on a canonical shape.
+Comparing run times for TDAstats.
+Panel 2.
+Comparing run times for Dionysus.
+Panel 3.
+Comparing run times for GUDHI.
+Limited point clouds were tested for high dimensional features due to computational limits.}
+\label{fig:box}
+\end{figure}
 
-Figure 2: Comparison of persistent homology calculation runtimes across R packages for n-dimensional boxes (n = 2, 3, 4, 5).
-For each dimension, features up to n-1 dimensions are calculated.
-Panel 1.
-Visual representation of 2-d box (maybe points, maybe box, maybe scattered both).
-Panel 2.
-Visual representation of 3-d box (maybe points, maybe box, maybe both scattered).
-Panels 3-6 (differing n={2,3,4,5} for dimension).
-Point shape/color: underlying engine.
-Horizontal axis: number of points.
-Vertical axis: runtime.
+\subsection{Comparison of alpha and Vietoris-Rips complexes within GUDHI}
 
-Figure 3: Runtime of persistent homology calculation as a function of underlying engine (4d noisy circle/annulus).
-Panel 1: visualize noisy circle - use imagination for 4d.
-Panel 2.
-Horizontal axis: number of points.
-Vertical axis: runtime.
-Point color/shape: underlying engine.
-Goal: show that TDAstats and GUDHIalpha are sufficiently faster than GUDHI and Dionysus, so the latter pair can be ignored for rest of figures.
+The GUDHI library provided functionality to calculate persistent homology with Rips complexes and with Alpha complexes. 
+The next set of results use only GUDHI to focus the comparison on alpha vs rips complex usage.
 
-Figure 4: Runtime of persistent homology calculation as a function of underlying engine and feature dimension (4d noisy circle/annulus).
-Panel 1 = GUDHI alpha engine.
-Horizontal axis = number of points.
-Vertical axis = runtime.
-Point color/shape = feature dimension.
-Panel 2 = panel 1, but with TDAstats engine.
-Goal: show how increasing dimension slows down the calculation time.
+\subsubsection{Runtime as a function of data dimension (const feature)}
 
-Figure 5: Memory use of Rips complex vs alpha complex.
-Engine = GUDHI(alpha).
-Panel 1 = 4-sphere.
-Panel 2 = 3-annulus.
-Panel 3 = torus.
-Panel 4 = 5-box.
-For each panel: point color/shape = engine (GUDHI vs alpha); horizontal axis = number of points; vertical axis = memory use; feature dimension = data dimension - 1 for all panels.
-Goal: compare memory
-POINT OUT THAT RIPSER DOESN'T MEASURE MEMORY B/C NO BOUNDARY MATRIX, BUT LIKELY SOMEWHERE IN BETWEEN THE TWO
+In our study, our computers could not calculate persistent homology for alpha complexes when the point cloud dimension exceeded 3. 
+This is understandable as alpha complex theoretical computational complexity increases exponentially for increasing point cloud dimension.
 
-Figure 6: Same as Figure 5, but TDAstats vs GUDHIalpha for runtime.
+This difference is seen when comparing run times as a function of changing point cloud dimension. 
+For calculating a 1 dimensional feature with Rips Complexes, the run time does not change for different dimensional point clouds.
+Even though alpha complexes perform better overall, their computation time increases with increasing point cloud dimension even for calculating the same feature dimension.
+This is shown in figure \ref{fig:ann}
 
-## Discussion
+\begin{figure}
+  \centering
+  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig4.png}
+  \caption{Comparing homology calculation as a function of point cloud dimension for Rips vs Alpha Complexes.
+Panel 1 = Comparing run times using GUDHI's Rips Complex library.
+Panel 2 = Comparing run times using GUDHI's Alpha Complex library.
+In all calculations, only 1 dimensional features were calculated.
+For our parameters, Alpha Complex run times are more linear in contrast to Rips more polynomial run times.
+Exact regression data is provided in Supplemental material.}
+\label{fig:ann}
+\end{figure}
+
+\subsubsection{Memory Benchmarks}
+As point cloud number increases, Rips complexes show a more dramatic rise in run time than alpha complexes. 
+This is predicted by Rips theoretical exponential complexity and alpha complex's theoretical polynomial complexity. 
+In both complexes, we use the boundary matrix to calculate of persistent homology. 
+The library GUDHI provides functions to return the boundary matrix, which allows us to measure object size.
+
+Rips complexes show a rapid growth in object size as the point cloud number increases in figure \label{fig:mem}, panel 1. 
+The growth pattern differences mirror our findings for run time growth with increasing point cloud size. 
+Alpha complexes show a linear growth in memory complexity whereas Rips complexes show a polynomial growth. 
+However, calculating the object size reveals some interesting differences in what parameters determine the object size of Rips vs Alpha Complexes.
+
+Figure \label{fig:mem}, panel 2 shows that Alpha Complex object size depends on the feature dimension, point cloud size, and point cloud shape with torus and uniform shapes having the higher object size.
+Figure \label{fig:mem}, panel 3 shows that Rips complex object size depends only on the point cloud size and feature dimensions.
+
+\begin{figure}
+  \centering
+  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig5.png}
+  \caption{Persistent Homology Boundary Matrix for Rips vs Alpha Complex.
+Shape of used point cloud is demonstrated by each image.
+Panel 1 = Rips Complex vs Alpha Complex for object size.
+Panel 2 = Alpha Complex object size depends on feature size, point cloud number, and point cloud shape.
+Panel 3 = Rips Complex object size depends on feature size and point cloud number alone.
+Point cloud shape is not defined as it had no impact on boundary matrix size.
+For a given n point cloud, n-1 dimension features were calculated in these examples.}
+\label{fig:mem}
+\end{figure}
+
+\subsubsection{Computational Limitations (advice/guidance for users)}
+In benchmarking all of our parameters, each point cloud was given up to 16 GB of RAM. 
+However, memory allocation errors as reported by R caused several of our parameters to not run. 
+This table describes the point cloud size limit for calculating a given feature dimension. 
+Notably, TDAstats has a higher threshold for point cloud size at high dimensions when compared to TDA's GUDHI or Dionysus libraries. 
+Alpha complex's limitation is better described as the point cloud dimension being the limiting factor, which is why 3 and 4 dimensional features were not computable.
+
+\begin{table}
+\centering
+\small
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+Feature & TDA Stats & Dionysus & Gudhi Rips & Gudhi Complex \\
+Dimension & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit\\
+\hline
+1 & Not Reached & Not Reached & Not Reached & Not Reached \\ 
+2 & Not Reached & 200 points & 280 points & Not Reached \\ 
+3 & 300 points & 100 points & 125 points & Could Not Compute \\
+4 & 100 points & 50 points & 75 points & Could Not Compute \\ 
+\hline
+\end{tabular}
+\caption{During the benchmarking, certain libraries were not able to calculate persistent homology for sufficiently complex point clouds. 
+Limiting factors were point cloud number and feature dimension calculation. 
+For Rips complexes, TDA stats had the highest performing capacity among the three Rips libraries. 
+Alpha complexes were not computable for any point cloud greater than 3 dimensions, so dimensional features 3 and higher could not be computed.}
+\end{table}
+
+\section{Discussion}
+As persistent homology calculations continue to become a more popular tool to analyze complex multidimensional data, it will be important to understand from a computing perspective which method to use.
+In this paper, we examined two forms of persistent homology complexes: Vietoris Rips and Alpha Complexes. 
+Both algorithms describe the topological features through the generation of simplicial complexes. 
+The advantage in saving computational time in choosing a particular algorithm depends on the point cloud characteristics.
+
+Figure \ref{fig:fin} shows that at high point cloud sizes, GUDHI’s alpha complex calculations outperforms TDAstats. 
+Alpha complexes gain polynomial complexity as the number of points increases whereas Rips complexes gain exponentially. 
+Specifically, alpha complexes are $O(n^\frac{d}{2})$ and Rips complexes are $O(2^n)$ where n is the number of points on a point cloud and d is the dimensions on the point cloud.
+
+Based on this theoretical complexity and our results, alpha complexes are superior for small dimensional point clouds (3 dimensions and below). 
+This advantage becomes markedly clear at high number of points. 
+If data is more than three dimensions, then it should be dimensionally reduced by a procedure such as PCA for efficient alpha complex usage.
+
+However, if data dimension cannot be reduced, then Rips complexes should be used. Among the different libraries, TDAstats has the fastest calculation time. 
+GUDHI and Dionysus significantly fall behind for high feature dimension calculations and high numbers of points. 
+GUDHI and Dionysus libraries, however, do have the option of returning the boundary matrix, which Ripser does not do as one of its time saving features.
+
+\begin{figure}
+  \centering
+  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig6.png}
+  \caption{TDAstats Rips Complexes vs GUDHI Alpha Complex average run times across all 3D shapes for 2D features.
+Panel 1 = Execution time on annulus.
+Panel 2 = Execution time on sphere.
+Panel 3 = Execution time on torus.
+Panel 4 = Execution time on uniform box}
+\label{fig:fin}
+\end{figure}
+
+While Rips Complexes can handle high dimensional data well, it still significantly slows down when looking for higher dimensional topological features. 
+The size of the boundary matrix rapidly increases and saturates the RAM even for small numbers of points (e.g. ~100 points when looking for 4 dimensional features). 
+Looking for high dimensional features in a large high dimensional point cloud remains a challenge. 
+Other persistent homology complex methods exist such as the Delaunay complex, Witness complex, etc. but they are not currently implemented in R [cite]. 
+Future challenges would be creating and implementing algorithms that reduce the computational complexity of higher dimensional topological feature calculations for R.
 
 Need to make a better version of this figure, but essential idea remains the same (recommendation for which R package to use depending on what you want to do).
-
 \begin{figure}
@@
 
+\section{Acknowledgements}
+Emily Dolson, PhD for assistance in using the HPC for collecting data
+
 \bibliography{tdabench}

### 2020-01-18 tdabench-paper/tdabench.Rmd
Imported all other text from Overleaf, Revised Intro

@@
 \section{Introduction}
+Topological data analysis (TDA) is a broad set of methodologies that characterize structural features of data sets inspired by topological principles.
+Within the umbrella of TDA, persistent homology represents an algebraic approach to understanding the number, characteristics, and persistence of structural features in an n-dimensional point cloud.
+
+In the basic workflow of persistent homology, a series of simplicial complexes are generated on point clouds to characterize the topological features. 
+Simplicial complexes are connected sets of points, lines, and n-dimensional shapes.
+There are several methods to generate these complexes on point clouds. 
+In this paper, we are using the Rips and Alpha Complex for our persistent homology calculations. 
+In both methods, a series of data structures, called simplicial complexes, are built on top of a point cloud.
+These complexes can be viewed as approximating real topologic relationships in the data.
+The exact method of constructing these complexes is described in the Mathematics section. 
+However, the key takeaway is that the simplicial complex attempts to reveal the underlying topology of the point cloud. 
+These features can be described by dimension and "persistence." 
+Long lasting (persistent) features that exist across the series of simplicial complexes intuitively represent "real" geometrical patterns as opposed to noise.
+
+One limitation of persistent homology is its run time limits for processing large numbers of points ( $> 10^9$ points) and calculating of high dimensional features ($>$ 2 dimensions). 
+Alpha complex calculations have a complexity of $O(n^\frac{d}{2})$ and Rips complex calculations have a complexity of $O(2^n)$ where n is the number of points and d is the dimension of the point cloud \cite{roadmap}.
+The limits and computational resources needed to run complex data sets with major persistent homology C++ software have been well characterized \cite{roadmap}.
+
+There are several C++ libraries that calculate Alpha and Rips complexes.
+Dionysus provides Alpha and Rips complex functionality \cite{Dionysus}.
+GUDHI is another C++ library that also provides both functionalities as well as other TDA methodologies \cite{gudhi}.
+These libraries have been implemented in R with the TDA package \cite{TDA} providing an easier interface to utilize persistent homology calculations.
+
+A new C++ library, Ripser, was developed as a more efficient way to calculate Rips Complexes.
+The code is supposed to outperform other similar libraries (e.g.
+Dionysus and GUDHI) by a factor over 40 in runtime \cite{Ripser}.
+This library was given an R interface in the package TDAstats \cite{TDAstats}. 
+TDAstats represents a more specialized package for topological data analysis compared to the TDA package.
+It calculates persistent homology using the more efficient Ripser algorithm and provides data visualization with integration of ggplot.
+While Ripser is known to be faster than the other libraries for Rips complex calculations in C++, how this translates in the R interface is not clear.
+While C++ generally runs faster than R, R provides higher level functionality, and appeals to a broad audience of data science users.
+
+Topological data analysis has had a broad range of usage from viral evolution to physical chemistry~\cite{TDA-Viral,TDA-PChem}. 
+Given that persistent homology can apply to any n-dimensional data set, it has wide ranging applicability. 
+Its main limiting factor is computational for calculating high dimensional features in large multi-dimensional data sets. 
+We believe lack of analysis of these higher dimensional features may under utilize persistent homology's analysis potential.
+In this project, we explore different methodologies to find efficient options that could solve this issue. 
+We specifically aim to characterize the computational performance (run time) and limits of persistent homology calculations between Rips complexes in the R packages TDAstats and TDA.
+We also explore the differences in computational performance between Rips and Alpha complexes in R to assess which datasets are most appropriate for each algorithm.
 
 \section{Mathematics of Persistent Homology}
+This section summarizes the mathematics behind generating simplicial complexes on point clouds in the workflow of persistent homology. 
+An n-dimensional simplex is the convex hull of n+1 vertices.
+More intuitively, n-dimensional simplex is the simplest n-dimensional object (e.g. a 1 simplex is a line, a 2-simplex is a triangle, 3-simplex is a tetrahedron, etc.).
+These simplices can be glued together on a common subsimplices to form a simplicial complex (e.g. two triangles sharing a common side).
+In a simplicial complex, certain features will arise that can be characterized by dimension number (Betti number): 0-dimensional features refer to connected components, 1-dimensional features refer to holes, 2-dimensional features refer to cavities, etc.
+~\cite{phom-survey}
+
+There are several different methods to construct a simplicial complex on a given point cloud S, but this paper focuses on the Rips complex and Alpha complex. 
+The Rips complex is perhaps the most common method for constructing a simplicial complex under persistent homology ~\cite{Rips-Complex}.
+In a point cloud of k points, a distance parameter, $\delta$, can be used to draw a circle of diameter $\delta$ around every single point in S.
+If the distance parameter is non-zero, then some of the resulting circles may intersect.
+If they intersect, a line is drawn to connect the points (which represent the centers of the growing circles).
+This group of points and lines form a simplicial complex.
+For each distance parameter, $\delta$, there will be some simplicial complex associated with it.
+As $\delta$ increases, different topological features may appear, persist, then disappear once $\delta$ has reached a threshold.
+Eventually, once $\delta$ reaches the maximum euclidean distance between any pair of points in the point cloud, a convex hull will form around all k points creating a (k-1)-dimensional simplex.
+By this stage, all features should have disappeared.
+A three column matrix can be created recording the Betti number of a certain feature, the distance, $\delta$, at which that feature appeared, and the $\delta$ at which that feature disappeared.
+This matrix characterizes the persistent homology of that point cloud.
 
@@
 \end{figure}
+
+Alpha complexes provide another method to generate simplicial complexes on the point cloud S.
+Alpha complexes are closely linked with Voronoi diagrams and Delaunay triangulations.
+The Voronoi diagram gives each point in S a cell such that each every point in the cell is closest to that point in S.
+The nerve of a Voronoi diagram is equivalent to the Delaunay Triangulation~\cite{alpha-complex}.
+Alpha complexes are simplicial complexes that are subsets of the Delaunay Triangulation.
+The parameter, $\alpha$, can describe the radius of a ball of each point in the point cloud S much like $\delta$ describes the diameter of a circle in the Rips Complex.
+An alpha complex at radius $\alpha$ is formed through connecting centers of intersecting regions formed by the intersection of $\alpha$ radius balls and the Voronoi diagrams. 
+Eventually, once $\alpha$ is large enough, the full Delaunay complex is formed.
+In between these stages, the birth and death of features at certain values of $\alpha$ can be captured in a three column persistent homology matrix much like the Rips Complex.
+
+In both methods, the boundary matrix records all simplicial complexes for a given parameter value. 
+From this matrix, the persistent homology matrix is calculated. 
+The size of the boundary matrix can describe the memory complexity on the RAM for persistent homology calculations. 
+We compare this memory complexity between Alpha and Rips complexes for this project.
+
 \begin{figure}

### 2020-01-18 tdabench-paper/diagram.R
Updated decision flow chart to reflect results

@@ -1,39 +1,37 @@
 #####DECISION FLOW DIAGRAM#####
 library("DiagrammeR")
 
-decision <- create_graph() %>%
-  add_node(label = "TDA method?",
-           node_aes = node_aes(shape = "rectangle",
-                               width = 1.1)) %>%               # 1
-  add_node(label = "GUDHI",
-           node_aes = node_aes(shape = "rectangle",
-                               width = 0.65)) %>%                     # 2
-  add_node(label = "Approx?",
-           node_aes = node_aes(shape = "rectangle",
-                               width = 0.75)) %>%      # 3
-  add_node(label = "TDAstats",
-           node_aes = node_aes(shape = "rectangle",
-                               width = 0.8)) %>%                  # 4
-  add_node(label = "Dim?",
-           node_aes = node_aes(shape = "rectangle",
-                               width = 0.5)) %>%           # 5
-  add_node(label = "GUDHI",
-           node_aes = node_aes(shape = "rectangle",
-                               width = 0.65)) %>%                     # 6
-  add_node(label = "TDAstats",
-           node_aes = node_aes(shape = "rectangle",
-                               width = 0.8)) %>%                  # 7
-  add_edge(from = 1, to = 2,
-           edge_aes = edge_aes(label = "Other")) %>%
-  add_edge(from = 1, to = 3,
-           edge_aes = edge_aes(label = "PHom")) %>%
-  add_edge(from = 3, to = 4,
-           edge_aes = edge_aes(label = "No")) %>%
-  add_edge(from = 3, to = 5,
-           edge_aes = edge_aes(label = "Yes")) %>%
-  add_edge(from = 5, to = 6,
-           edge_aes = edge_aes(label = "<= 3")) %>%
-  add_edge(from = 5, to = 7,
-           edge_aes = edge_aes(label = "> 3"))
 
-decision %>% render_graph(layout = "tree")
+diagram <- DiagrammeR::grViz("
+digraph graph2 {
+
+graph [layout = dot]
+
+# node definitions with substituted label text
+node [shape = ellipse, width = 2, height = 1, fillcolor = Biege]
+a [label = 'TDA method']
+b [label = 'Data set \u2264 3 dim']
+c [label = 'Data dim reducible']
+d [label = 'TDAstats \n(Rips Complex)']
+
[…]

### 2020-01-18 tdabench-paper/tdabench.Rmd
Edited and updated remaining sections

@@
 There are several methods to generate these complexes on point clouds. 
-In this paper, we are using the Rips and Alpha Complex for our persistent homology calculations. 
+In this paper, we are using the Rips and Alpha complex for our persistent homology calculations. 
 In both methods, a series of data structures, called simplicial complexes, are built on top of a point cloud.
@@
 
-A new C++ library, Ripser, was developed as a more efficient way to calculate Rips Complexes.
+A new C++ library, Ripser, was developed as a more efficient way to calculate Rips complexes.
 The code is supposed to outperform other similar libraries (e.g.
@@
 
-Topological data analysis has had a broad range of usage from viral evolution to physical chemistry~\cite{TDA-Viral,TDA-PChem}. 
+Topological data analysis has had a broad range of usage from viral evolution to physical chemistry\cite{TDA-Viral,TDA-PChem}. 
 Given that persistent homology can apply to any n-dimensional data set, it has wide ranging applicability. 
@@
 In a simplicial complex, certain features will arise that can be characterized by dimension number (Betti number): 0-dimensional features refer to connected components, 1-dimensional features refer to holes, 2-dimensional features refer to cavities, etc.
-~\cite{phom-survey}
+\cite{phom-survey}
 
-There are several different methods to construct a simplicial complex on a given point cloud S, but this paper focuses on the Rips complex and Alpha complex. 
-The Rips complex is perhaps the most common method for constructing a simplicial complex under persistent homology ~\cite{Rips-Complex}.
+There are several different methods to construct a simplicial complex on a given point cloud S, but this paper focuses on the Rips and Alpha complexes. 
+The Rips complex is perhaps the most common method for constructing a simplicial complex under persistent homology \cite{Rips-Complex}.
 In a point cloud of k points, a distance parameter, $\delta$, can be used to draw a circle of diameter $\delta$ around every single point in S.
@@
 Eventually, once $\delta$ reaches the maximum euclidean distance between any pair of points in the point cloud, a convex hull will form around all k points creating a (k-1)-dimensional simplex.
-By this stage, all features should have disappeared.
+At this stage, all features should have disappeared.
 A three column matrix can be created recording the Betti number of a certain feature, the distance, $\delta$, at which that feature appeared, and the $\delta$ at which that feature disappeared.
@@
 The Voronoi diagram gives each point in S a cell such that each every point in the cell is closest to that point in S.
-The nerve of a Voronoi diagram is equivalent to the Delaunay Triangulation~\cite{alpha-complex}.
+The nerve of a Voronoi diagram is equivalent to the Delaunay Triangulation\cite{alpha-complex}.
 Alpha complexes are simplicial complexes that are subsets of the Delaunay Triangulation.
-The parameter, $\alpha$, can describe the radius of a ball of each point in the point cloud S much like $\delta$ describes the diameter of a circle in the Rips Complex.
-An alpha complex at radius $\alpha$ is formed through connecting centers of intersecting regions formed by the intersection of $\alpha$ radius balls and the Voronoi diagrams. 
-Eventually, once $\alpha$ is large enough, the full Delaunay complex is formed.
-In between these stages, the birth and death of features at certain values of $\alpha$ can be captured in a three column persistent homology matrix much like the Rips Complex.
+The parameter, $\alpha$, can describe the radius of a ball of each point in the point cloud S much like $\delta$ describes the diameter of a circle in the Rips complex.
+An Alpha complex at radius $\alpha$ is formed through connecting centers of intersecting regions formed by intersecting the union of $\alpha$ radius balls and the Voronoi diagram. 
+Eventually, once $\alpha$ is large enough, the full Delaunay Triangulation is formed.
+In between these stages, the birth and death of features at certain values of $\alpha$ can be captured in a three column persistent homology matrix much like the Rips complex.
 
@@
 Within the \CRANpkg{TDA} package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the \CRANpkg{TDAstats} package, benchmark data was collected for the Ripser \citep{Ripser} library.
-As alpha complex calculation was only implemented in GUDHI, alpha complex benchmark data was naturally only collected for the single library.
+As Alpha complex calculation was only implemented in GUDHI, Alpha complex benchmark data was naturally only collected for the single library.
 Measuring memory usage proved challenging since all the libraries calculating persistent homology were implemented in another programming language and then wrapped in R.
@@
 \subsection{Runtime comparison between packages for calculation of V-R complex}
-In our analysis, we first compare the three Rips complexes libraries among each other. 
-We examine the run time performance of persistence homology calculation under different conditions. 
+In our analysis, we compare runtime performances of the three Rips complexes libraries for persistence homology calculation under different conditions.
 We then focus our analysis on comparing Rips complexes with Alpha complexes using the GUDHI package.
-In all runtime calculations, 10 iterations were performed and data are presented as mean +/- standard deviation.
 
@@
 The torus point cloud is a classical shape used to demonstrate persistent homology calculations. 
-A persistent homology calculation of a uniformly distributed torus would reveal a large 2 dimensional feature (void) in the middle representing the “hole” of the torus. 
-In figure \ref{fig:tor}, we calculated up to 2 dimensional features of a torus 3d point cloud for an increasing number of points across all TDA libraries.
+A persistent homology calculation of a uniformly distributed torus would reveal a large 1 dimensional feature representing the middle circle of the torus.
+There would also be a large 2 dimensional feature representing the inner cavity of the torus.
+In figure \ref{fig:tor}, we calculated up to 2 dimensional features of torus 3d points cloud for an increasing number of points across all TDA libraries.
 
@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/fig1.png}
-  \caption{Comparison of Rips Complex persistent homology calculation runtimes across R packages for a torus.
+  \caption{Comparison of Rips complex persistent homology calculation runtimes across R packages for a torus.
 The x-axis represents the number of points in a 3D torus point cloud, and y-axis represents average run time.
@@
 \subsubsection{Runtime as a function of data dimension (feature = n - 1)}
-From the torus comparison, we see a clear difference in performance between the three Rips libraries. 
-We wanted to see if this performance difference changed with more complex point clouds. 
+For small point cloud numbers, Gudhi and TDAstats performed similarly while Dionysus lagged behind. 
+At larger point clouds, TDAstats outperforms Gudhi, and the difference among the Rips libraries is more visible.
+This result is especially true for calculating higher feature dimensions on large point clouds. 
+While the runtime values changed for each point cloud condition, the growth pattern with respect to increasing number of points was similar for all Rips complex libraries. 
+For the conditions tested in our analysis, all Rips libraries resulted in an apparent polynomial growth for run time. 
+The exact values for the log-log regression of this growth pattern is provided in supplemental data.
+
 Each panel in figure \ref{fig:cir} corresponds to an n-dimensional circle (circle, sphere, and hypersphere). 
 For each point cloud, n-1 dimensional features were calculated.
-
-A similar growth pattern was seen among the TDA libraries. 
-As the sphere dimensions and feature dimensions increased, the execution time growth becomes apparent sooner. 
+As the sphere dimensions and feature dimensions increased, the polynomial time growth becomes apparent sooner. 
 At these higher dimensional features, some of the curves seem to cut off. 
@@
 In the field, most work has only included calculations up to 1 or 2 feature dimensions. 
-Perhaps this due to difficulty of conceptualizing what very high dimensional features represent. 
-High dimensional feature calculations are also very taxing computationally. 
+Perhaps this is due to difficulty of conceptualizing what high dimensional features represent. 
+However, it might also be due to computational restrictions of calculating high dimensional features. 
 For Figure \ref{fig:box}, we restricted the calculations to very small point clouds due to computational limits.
@@
 As feature dimension increased, Rips complexes show a significant rise in run time with increasing feature dimension within an 8 dimensional point cloud. 
-With this small sample set, it is not clear whether the growth is exponential or polynomial.
+With this small sample set, it is not clear whether the growth is exponential or polynomial, so no regression values are provided in the supplement.
 
@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/fig3.png}
-  \caption{Comparison of Rips Complex persistent homology calculation as a function of feature dimension.
+  \caption{Comparison of Rips complex persistent homology calculation as a function of feature dimension.
 Panel 1.
@@
 \subsection{Comparison of alpha and Vietoris-Rips complexes within GUDHI}
-
-The GUDHI library provided functionality to calculate persistent homology with Rips complexes and with Alpha complexes. 
-The next set of results use only GUDHI to focus the comparison on alpha vs rips complex usage.
+The GUDHI library provided functionality to calculate persistent homology with Rips and Alpha complexes. 
+Therefore, we only used the GUDHI library to compare performance between Rips and Alpha complexes.
 
 \subsubsection{Runtime as a function of data dimension (const feature)}
+In our study, our computers could not calculate persistent homology for Alpha complexes when the point cloud dimension exceeded 3. 
+This is understandable as Alpha complex theoretical complexity increases exponentially for increasing point cloud dimension.
 
-In our study, our computers could not calculate persistent homology for alpha complexes when the point cloud dimension exceeded 3. 
-This is understandable as alpha complex theoretical computational complexity increases exponentially for increasing point cloud dimension.
+Changes in computational performance from changing point cloud dimension is one major difference between calculation of Rips and Alpha complexes.
+This difference can be seen when comparing run times as a function of changing point cloud dimension.
+In figure \ref{fig:ann}, we show that Rips complexes  run time does not change for different dimensional point clouds as long as feature dimension is held constant.
+Even though Alpha complexes perform better overall, their computation time increases with increasing point cloud dimension even when calculating the same feature dimension.
 
-This difference is seen when comparing run times as a function of changing point cloud dimension. 
-For calculating a 1 dimensional feature with Rips Complexes, the run time does not change for different dimensional point clouds.
-Even though alpha complexes perform better overall, their computation time increases with increasing point cloud dimension even for calculating the same feature dimension.
-This is shown in figure \ref{fig:ann}
+Another key difference is that Alpha complex runtime seems to have an apparent linear rise in the conditions for our analysis. 
+We provide the linear regression coefficients in the supplemental data.
 
@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/fig4.png}
-  \caption{Comparing homology calculation as a function of point cloud dimension for Rips vs Alpha Complexes.
-Panel 1 = Comparing run times using GUDHI's Rips Complex library.
-Panel 2 = Comparing run times using GUDHI's Alpha Complex library.
+  \caption{Comparing homology calculation as a function of point cloud dimension for Rips vs Alpha complexes.
+Panel 1 = Comparing run times using GUDHI's Rips complex library.
+Panel 2 = Comparing run times using GUDHI's Alpha complex library.
 In all calculations, only 1 dimensional features were calculated.
-For our parameters, Alpha Complex run times are more linear in contrast to Rips more polynomial run times.
+For our parameters, Alpha complex run times are more linear in contrast to Rips more polynomial run times.
 Exact regression data is provided in Supplemental material.}
@@
 \subsubsection{Memory Benchmarks}
-As point cloud number increases, Rips complexes show a more dramatic rise in run time than alpha complexes. 
-This is predicted by Rips theoretical exponential complexity and alpha complex's theoretical polynomial complexity. 
-In both complexes, we use the boundary matrix to calculate of persistent homology. 
+As point cloud number increases, Rips complexes show a more dramatic rise in run time than Alpha complexes. 
+This is predicted by Rips theoretical exponential complexity and Alpha complex's theoretical polynomial complexity. 
+In both complexes, the boundary matrix is used to calculate persistent homology (explained in detail in the Mathematics section). 
 The library GUDHI provides functions to return the boundary matrix, which allows us to measure object size.
@@
 Alpha complexes show a linear growth in memory complexity whereas Rips complexes show a polynomial growth. 
-However, calculating the object size reveals some interesting differences in what parameters determine the object size of Rips vs Alpha Complexes.
+However, calculating the object size reveals some interesting differences in what parameters determine the object size of Rips vs Alpha complexes.
 
-Figure \label{fig:mem}, panel 2 shows that Alpha Complex object size depends on the feature dimension, point cloud size, and point cloud shape with torus and uniform shapes having the higher object size.
+Figure \label{fig:mem}, panel 2 shows that Alpha complex object size depends on the feature dimension, point cloud size, and point cloud shape with torus and uniform shapes having the higher object size.
+Noisier shapes like the box and annulus increase object size compared to more uniform shapes like the sphere.
 Figure \label{fig:mem}, panel 3 shows that Rips complex object size depends only on the point cloud size and feature dimensions.
@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/fig5.png}
-  \caption{Persistent Homology Boundary Matrix for Rips vs Alpha Complex.
+  \caption{Persistent Homology Boundary Matrix for Rips vs Alpha complex.
 Shape of used point cloud is demonstrated by each image.
-Panel 1 = Rips Complex vs Alpha Complex for object size.
-Panel 2 = Alpha Complex object size depends on feature size, point cloud number, and point cloud shape.
-Panel 3 = Rips Complex object size depends on feature size and point cloud number alone.
+Panel 1 = Rips complex vs Alpha complex for object size.
+Panel 2 = Alpha complex object size depends on feature size, point cloud number, and point cloud shape.
+Panel 3 = Rips complex object size depends on feature size and point cloud number alone.
 Point cloud shape is not defined as it had no impact on boundary matrix size.
@@
 \hline
-Feature & TDA Stats & Dionysus & Gudhi Rips & Gudhi Complex \\
+Feature & TDA Stats & Dionysus & Gudhi Rips & Gudhi Alpha \\
 Dimension & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit\\
@@
 \section{Discussion}
-As persistent homology calculations continue to become a more popular tool to analyze complex multidimensional data, it will be important to understand from a computing perspective which method to use.
-In this paper, we examined two forms of persistent homology complexes: Vietoris Rips and Alpha Complexes. 
+As persistent homology calculations continue to become a more popular tool to analyze complex multidimensional data, it will be important to understand from a computational perspective which method to use.
+In this paper, we examined two forms of persistent homology complexes: Vietoris Rips and Alpha complexes. 
 Both algorithms describe the topological features through the generation of simplicial complexes. 
@@
 
-Figure \ref{fig:fin} shows that at high point cloud sizes, GUDHI’s alpha complex calculations outperforms TDAstats. 
-Alpha complexes gain polynomial complexity as the number of points increases whereas Rips complexes gain exponentially. 
-Specifically, alpha complexes are $O(n^\frac{d}{2})$ and Rips complexes are $O(2^n)$ where n is the number of points on a point cloud and d is the dimensions on the point cloud.
+Figure \ref{fig:fin} shows that at high point cloud sizes, GUDHI’s Alpha complex calculations outperforms TDAstats. 
+Theoritically, Alpha complexes gain polynomial complexity as the number of points increases whereas Rips complexes gain exponentially. 
+Specifically, Alpha complexes are $O(n^\frac{d}{2})$ and Rips complexes are $O(2^n)$ where n is the number of points on a point cloud and d is the dimensions on the point cloud. 
+For the conditions in our paper, Rips and Alpha complexes both performed better than their theoritical maximums. 
+Rips complex calculations consistently had a polynomial growth for both runtime and boundary matrix size while Alpha complexes had linear growth.
 
-Based on this theoretical complexity and our results, alpha complexes are superior for small dimensional point clouds (3 dimensions and below). 
-This advantage becomes markedly clear at high number of points. 
-If data is more than three dimensions, then it should be dimensionally reduced by a procedure such as PCA for efficient alpha complex usage.
+Based on this theoretical complexity and our results, Alpha complexes are superior for small dimensional point clouds (3 dimensions and below). 
+This advantage becomes especially clear at high number of points.
+However, with out computational resources, Alpha complexes were not usable for point cloud dimensions greater than 3. 
+If data is more than three dimensions, then it should be dimensionally reduced by a procedure such as PCA for before using Alpha complexes.
 
@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/fig6.png}
-  \caption{TDAstats Rips Complexes vs GUDHI Alpha Complex average run times across all 3D shapes for 2D features.
+  \caption{TDAstats Rips complexes vs GUDHI Alpha complex average run times across all 3D shapes for 2D features.
 Panel 1 = Execution time on annulus.
@@
 
-While Rips Complexes can handle high dimensional data well, it still significantly slows down when looking for higher dimensional topological features. 
+While Rips complexes can handle high dimensional data well, it still significantly slows down when looking for higher dimensional topological features. 
 The size of the boundary matrix rapidly increases and saturates the RAM even for small numbers of points (e.g. ~100 points when looking for 4 dimensional features). 
@@
 
-Need to make a better version of this figure, but essential idea remains the same (recommendation for which R package to use depending on what you want to do).
 \begin{figure}
   \centering
-  \includegraphics[height=2.5in]{decision}
+  \includegraphics[height=2.5in]{decision.png}
+    \caption{Basic decision chart for choosing TDA library based on dataset.}
+\label{decision}
 \end{figure}

### 2020-01-20 tdabench-paper/tdabench.Rmd
minor text changes

@@
 One limitation of persistent homology is its run time limits for processing large numbers of points ( $> 10^9$ points) and calculating of high dimensional features ($>$ 2 dimensions). 
-Alpha complex calculations have a complexity of $O(n^\frac{d}{2})$ and Rips complex calculations have a complexity of $O(2^n)$ where n is the number of points and d is the dimension of the point cloud \cite{roadmap}.
+Alpha complex calculations have a complexity of $O(n^\frac{d}{2})$ and Rips complex calculations have a complexity of $O(2^n)$ where n is the number of points and d is the dimension of the point cloud \citep{roadmap}.
 The limits and computational resources needed to run complex data sets with major persistent homology C++ software have been well characterized \cite{roadmap}.
@@
 <!--An exclusively R (v3.6.2) code pipeline was built for this project \citep{Rlang,ihaka1996}.-->
-We use \CRANpkg{readr} v1.3.1 to read rectangular data \cite{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2} and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
+We use \CRANpkg{readr} v1.3.1 to read rectangular data \citep{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2} and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
 Runtime calculations were averaged over 10 iterations and are visualized as mean $\pm$ SD.

### 2020-01-20 tdabench-paper/tdabench.Rmd

@@
 <!--An exclusively R (v3.6.2) code pipeline was built for this project \citep{Rlang,ihaka1996}.-->
-We use \CRANpkg{readr} v1.3.1 to read rectangular data \citep{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2} and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
+We use \CRANpkg{readr} v1.3.1 to read rectangular data \citep{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2}, \CRANpkg{scatterplot3d} v0.3-41 \citep{scatterplot3d}, \CRANpkg{recexcavAAR} v0.3.0 \citep{recexcavAAR}, and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
 Runtime calculations were averaged over 10 iterations and are visualized as mean $\pm$ SD.
 Datasets were generated by sampling functions in base R to generate points uniformly distributed over circles (dimension = 2), spheres (dimension = 3), filled squares (dimension = 2), filled cubes (dimension = 3, 4), and tori (dimension = 3).
-The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after pragmatic considerations of available computational resources.
-For consistency, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} R packages.
+The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after considering available computational resources.
+For consistency between software libraries, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} R packages.
 Within the \CRANpkg{TDA} package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the \CRANpkg{TDAstats} package, benchmark data was collected for the Ripser \citep{Ripser} library.

### 2020-01-21 tdabench-paper/tdabench.Rmd
changed figures slightly to update new figure 5 arrangement

@@
 
-Rips complexes show a rapid growth in object size as the point cloud number increases in figure \label{fig:mem}, panel 1. 
+Rips complexes show a rapid growth in object size as the point cloud number increases in figure \label{fig:mem1}. 
 The growth pattern differences mirror our findings for run time growth with increasing point cloud size. 
@@
 
-Figure \label{fig:mem}, panel 2 shows that Alpha complex object size depends on the feature dimension, point cloud size, and point cloud shape with torus and uniform shapes having the higher object size.
+Figure \label{fig:mem2}, panel 2 shows that Alpha complex object size depends on the feature dimension, point cloud size, and point cloud shape with torus and uniform shapes having the higher object size.
 Noisier shapes like the box and annulus increase object size compared to more uniform shapes like the sphere.
-Figure \label{fig:mem}, panel 3 shows that Rips complex object size depends only on the point cloud size and feature dimensions.
+Figure \label{fig:mem2}, panel 1 shows that Rips complex object size depends only on the point cloud size and feature dimensions.
 
@@
   \centering
-  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig5.png}
-  \caption{Persistent Homology Boundary Matrix for Rips vs Alpha complex.
-Shape of used point cloud is demonstrated by each image.
-Panel 1 = Rips complex vs Alpha complex for object size.
+  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig5a.png}
+  \caption{Persistent Homology Boundary Matrix for Rips vs Alpha complex. Shape of used point cloud is demonstrated by each image. Rips complex vs Alpha complex for object size.}
+\label{fig:mem1}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig5t.png}
+  \caption{Different factors change boundary matrix size for each complex.
+Panel 1 = Rips complex object size depends on feature size and point cloud number alone.
 Panel 2 = Alpha complex object size depends on feature size, point cloud number, and point cloud shape.
-Panel 3 = Rips complex object size depends on feature size and point cloud number alone.
 Point cloud shape is not defined as it had no impact on boundary matrix size.
 For a given n point cloud, n-1 dimension features were calculated in these examples.}
-\label{fig:mem}
+\label{fig:mem2}
 \end{figure}

### 2020-01-21 Figures/Creating_ggplot_Figures.R
Widened images with three panels, cropped out white space, split figure 5 into two figures (5a and 5t...5t = 5b+ 5c)

@@ -119,7 +119,7 @@ fig.2 <- fig.2 + theme_classic() +
 fig.2
 
 ggsave("./Figures/Unrasterized_Images/fig2.png", plot = fig.2,
-       scale = 1, width = 8, height = 4, units = "in",
+       scale = 1, width = 12, height = 4, units = "in",
        dpi = 400, limitsize = TRUE)
 
 
@@ -157,7 +157,7 @@ fig.3 <- fig.3 + theme_classic() +
 fig.3
 
 ggsave("./Figures/Unrasterized_Images/fig3.png", plot = fig.3,
-       scale = 1, width = 6, height = 4, units = "in",
+       scale = 1, width = 12, height = 4, units = "in",
        dpi = 400, limitsize = TRUE)
 
 
@@ -199,7 +199,7 @@ fig.4 <- fig.4 + theme_classic() +
 fig.4
 
 ggsave("./Figures/Unrasterized_Images/fig4.png", plot = fig.4,
-       scale = 1, width = 6, height = 4, units = "in",
+       scale = 1, width = 8, height = 4, units = "in",
        dpi = 400, limitsize = TRUE)
 
 ####Figure 5: Object Size use of Rips complex vs alpha complex. Engine = GUDHI(alpha).#### 
@@ -219,15 +219,15 @@ fig.5a <- ggplot(data.fig.5a, aes(x=num.points, y=memory, color=library)) +
 
 #Editing the colors (Making everything blank)
 fig.5a <- fig.5a + theme_classic() +
-                    theme(legend.position = c(1.10, 0.550),
+                    theme(legend.position = c(.915, 0.915),
                                  legend.title = element_text(size = 9),
                                  legend.text = element_text(size = 7),
                                  plot.title = element_text(hjust = 0.5),
                                  legend.text.align = 0,
-                                 legend.title.align = 0,
+                                 legend.title.align = 0.8,
                                  axis.text.x = element_text(angle=0, hjust=.025),
                                  strip.background = element_blank(),
-                                 plot.margin = unit(c(1,3,1,1), "cm"),
+                                 #plot.margin = unit(c(1,3,1,1), "cm"),
                                  strip.text.x = element_blank())
 
 fig.5a
@@ -261,11 +261,12 @@ fig.5b <- fig.5b + theme_classic() +
                legend.title.align = 0,
[…]

### 2020-01-21 Figures/Rasterize_Figures.R

@@ -52,12 +52,14 @@ image_write(fig.1.rast, path = './Figures/Final_Figures/fig1.png', format = 'png
 fig2.unrast <- image_read("./Figures/Unrasterized_Images/fig2.png")
 #Combine Figure and Image
 fig2.rast1 <- circle %>% image_scale("200") %>%
-                         image_composite(fig2.unrast, ., offset = "+2125+205", 
+                         image_composite(fig2.unrast, ., offset = "+3225+205", 
                               gravity = "northeast")
 
 fig2.rast2 <- sphere %>% image_scale("200") %>%
-                         image_composite(fig2.rast1, ., offset = "+1100+205", 
+                         image_composite(fig2.rast1, ., offset = "+1700+205", 
                   gravity = "northeast")
+
+fig2.rast2
 
 #Write Out File
 image_write(fig2.rast2, path = './Figures/Final_Figures/fig2.png', format = 'png')
@@ -66,22 +68,22 @@ image_write(fig2.rast2, path = './Figures/Final_Figures/fig2.png', format = 'png
 fig5a.unrast <- image_read("./Figures/Unrasterized_Images/fig5a.png")
 #Combine Figure and Images
 fig5a.rast1 <- annulus.3 %>% image_scale("200") %>%
-  image_composite(fig5a.unrast, ., offset = "+1325+300", 
+  image_composite(fig5a.unrast, ., offset = "+1875+175", 
                   gravity = "northeast")
 
 fig5a.rast2 <- sphere %>% image_scale("190") %>%
-  image_composite(fig5a.rast1, ., offset = "+525+300", 
+  image_composite(fig5a.rast1, ., offset = "+720+175", 
                   gravity = "northeast")
 
 fig5a.rast3 <- torus %>% image_scale("210") %>%
-  image_composite(fig5a.rast2, ., offset = "+1325+850", 
+  image_composite(fig5a.rast2, ., offset = "+1865+850", 
                   gravity = "northeast")
 
 fig5a.rast4 <- cube %>% image_scale("210") %>%
-  image_composite(fig5a.rast3, ., offset = "+525+850", 
+  image_composite(fig5a.rast3, ., offset = "+710+850", 
                   gravity = "northeast")
 
-
+fig5a.rast4
 #Write Out File
[…]

### 2020-01-24 Figures/Creating_Point_Clouds.R
Code has been consolidated to one Reproducible Script

@@ -1,124 +1,0 @@
-library(plyr)
-library(readr)
-library(dplyr)
-library(TDA)
-library(TDAstats)
-library(bench)
-library(pryr)
-library(scatterplot3d)
-library(recexcavAAR)
-
-
-source("Functions.R")
-
-####Circles####
-#Create circle, zero the y so that the circle is looking normal
-circle.2d <- unifcircle(200, 2)
-circle.2d <- cbind(circle.2d, 0)
-circle.2d <- circle.2d[,c(1,3,2)]
-
-#Save Plot As Image
-png(filename = "./Figures/Unrasterized_Images/circle.png",
-    width = 1.36*5, height = 1*5, units = "in", res = 300)
-
-scatterplot3d(circle.2d, pch = 20, 
-              grid=F, box=F,
-              axis = F, angle = 0, highlight.3d = T)
-dev.off()
-
-
-#Create sphere
-circle.3d <- unifcircle(900, 3)
-
-#Save Plot As Image
-png(filename = "./Figures/Unrasterized_Images/sphere.png",
-    width = 1.16*5, height = 1*5, units = "in", res = 300)
-
-scatterplot3d(circle.3d, pch = 20, 
-                                grid=F, box=F,
-                                axis = F, angle = 5, highlight.3d = T)
-
-dev.off()
-
-
-####Annuluses####
-#Create Annulus, zero the y so that the circle is looking normal
-annulus.2d <- noisycircle(400, 2)
-annulus.2d <- cbind(annulus.2d, 0)
-annulus.2d <- annulus.2d[,c(1,3,2)]
-
-#Save Plot As Image
-png(filename = "./Figures/Unrasterized_Images/2annulus.png",
-    width = 1.26*5, height = 1*5, units = "in", res = 300)
-
-scatterplot3d(annulus.2d, pch = 20, 
-                                 highlight.3d = T,  
-                                grid=F, box=F,
-                                axis = F, angle = 0)
-dev.off()
-
-
-#Create 3d annulus
-annulus.3d <- noisycircle(1300, 3)
-
-#Save Plot As Image
-png(filename = "./Figures/Unrasterized_Images/3annulus.png",
-    width = 1.16*5, height = 1*5, units = "in", res = 300)
-
-scatterplot3d(annulus.3d, pch = 20, 
-                                grid=F, box=F,
-                                axis = F, angle = 5, highlight.3d = T)
-
-dev.off()
-
-
-####Torus####
[…]

### 2020-01-24 Figures/Creating_ggplot_Figures.R

@@ -1,768 +1,0 @@
-library(dplyr)
-library(ggplot2)
-library(reshape)
-library(ggforce)
-
-#Reading in the time data
-data.time <- rbind(read.csv("./Cluster_data/annulus.csv", header = TRUE),
-                   read.csv("./Cluster_data/circle.csv", header = TRUE),
-                   read.csv("./Cluster_data/torus.csv", header = TRUE),
-                   read.csv("./Cluster_data/uniform.csv", header = TRUE))
-#Adding Column Names
-colnames(data.time) <- c("row", "measure.type", "point.cloud", "point.cloud.dim", 
-                         "num.points", "feat.dim", "library", 
-                         "time1", "time2", "time3", "time4", 
-                         "time5", "time6", "time7",
-                         "time8", "time9", "time10")
-
-#Remove extraneaous row variable
-data.time <- data.time %>% select(-"row")
-
-#Creating average and standard deviation columns for time data
-data.time$avg.time <- apply(data.time[,7:16],1,mean)
-data.time$std <- apply(data.time[,7:16],1,sd)
-data.time$min.time <- apply(data.time[,7:16],1,min)
-data.time$max.time <- apply(data.time[,7:16],1,max)
-
-#Create Factor labels for libraries
-data.time$library <- factor(data.time$library,
-                            levels = c("stats","Dionysus","GUDHI", "GUDHIalpha"),
-                            labels = c("TDAstats", "Dionysus", "GUDHI Rips", "GUDHI Alpha"))
-
-#Read in object size data
-data.mem <- read.csv("./mem1.csv", header = FALSE)
-
-#Name columns
-colnames(data.mem) <- c("measure.type", "point.cloud", "point.cloud.dim", 
-                        "num.points", "feat.dim", "library", "memory")
-
-#Create Factor labels for libraries
-data.mem$library <- factor(data.mem$library,
-                            levels = c("stats","Dionysus","GUDHI", "GUDHIalpha"),
-                            labels = c("TDAstats", "Dionysus", "GUDHI Rips", "GUDHI Alpha"))
-
-
-####Figure_1####
-#Goal: overview showing speed differences across engines on a canonical shape.
-
-#Selected torus data scanning for two dimensional features
-data.fig.1 <- subset(data.time, point.cloud == "torus" & feat.dim == 2 & 
-                       library != "GUDHI Alpha") 
-
-#Point graph; X: Num.points; Y: Runtime; Color: TDA library
[…]

### 2020-01-24 Figures/Rasterize_Figures.R

@@ -1,257 +1,0 @@
-library(plyr)
-library(readr)
-library(dplyr)
-library(TDA)
-library(TDAstats)
-library(bench)
-library(pryr)
-library(ggplot2)
-library(magick)
-
-####Reading in Point Clouds and Cropping them####
-torus <- image_read("./Figures/Unrasterized_Images/torus.png") %>% 
-         image_scale("1000") %>% 
-        image_crop("525x410+225+150")
-
-circle <- image_read("./Figures/Unrasterized_Images/circle.png") %>% 
-  image_scale("1000") %>% 
-  image_crop("525x450+225+125")
-
-sphere <- image_read("./Figures/Unrasterized_Images/sphere.png") %>% 
-  image_scale("1000") %>% 
-  image_crop("525x525+225+150")
-
-annulus.2 <- image_read("./Figures/Unrasterized_Images/2annulus.png") %>% 
-  image_scale("1200") %>% 
-  image_crop("525x450+315+240")
-
-annulus.3 <- image_read("./Figures/Unrasterized_Images/3annulus.png") %>% 
-  image_scale("1400") %>% 
-  image_crop("525x525+410+320")
-
-square <- image_read("./Figures/Unrasterized_Images/square.png") %>% 
-  image_scale("900") %>% 
-  image_crop("525x450+175+75")
-
-cube <- image_read("./Figures/Unrasterized_Images/cube.png") %>% 
-  image_scale("1000") %>% 
-  image_crop("525x475+225+200")
-
-
-####Figure_1####
-#Read in Figure
-fig1.unrast <- image_read("./Figures/Unrasterized_Images/fig1.png")
-#Combine Figure and Image
-fig.1.rast <- image_composite(fig1.unrast, torus, offset = "+25+125", gravity = "northeast")
-#Write Out File
-image_write(fig.1.rast, path = './Figures/Final_Figures/fig1.png', format = 'png')
-
-
-####Figure_2####
-#Read in Figure
-fig2.unrast <- image_read("./Figures/Unrasterized_Images/fig2.png")
-#Combine Figure and Image
-fig2.rast1 <- circle %>% image_scale("200") %>%
-                         image_composite(fig2.unrast, ., offset = "+3225+205", 
-                              gravity = "northeast")
-
-fig2.rast2 <- sphere %>% image_scale("200") %>%
-                         image_composite(fig2.rast1, ., offset = "+1700+205", 
[…]

### 2020-01-24 Figures/Extra_Mac_Figures.R
Separated the mac figures to indicate Mac data is extra

@@ -1,0 +1,219 @@
+library(dplyr)
+library(ggplot2)
+library(reshape)
+library(ggforce)
+
+
+####Alternative Figures from Extra Mac Desktop Time Data####
+#This data was run on a 16 GB lab mac computer
+#While we did not present this data in the official paper
+#We ran it just to be sure that the benchmark results were similar
+#To the High powered cluster data
+
+#Figure 5 was left out since object size is invariant and always the same  
+
+#Reading in the time data
+data.time.mac <- read.csv("./time_mac_desktop.csv", header = FALSE)
+#Adding Column Names
+colnames(data.time.mac) <- c("measure.type", "point.cloud", "point.cloud.dim", 
+                         "num.points", "feat.dim", "library", 
+                         "time1", "time2", "time3", "time4", 
+                         "time5", "time6", "time7",
+                         "time8", "time9", "time10")
+
+#Creating average and standard deviation columns for time data
+data.time.mac$avg.time <- apply(data.time.mac[,7:16],1,mean)
+data.time.mac$std <- apply(data.time.mac[,7:16],1,sd)
+data.time.mac$min.time <- apply(data.time.mac[,7:16],1,min)
+data.time.mac$max.time <- apply(data.time.mac[,7:16],1,max)
+
+#Create Factor labels for libraries
+data.time.mac$library <- factor(data.time.mac$library,
+                            levels = c("stats","Dionysus","GUDHI", "GUDHIalpha"),
+                            labels = c("TDAstats", "Dionysus", "GUDHI Rips", "GUDHI Alpha"))
+
+
+####Figure 1 Mac####
+data.fig.1.mac <- subset(data.time.mac, point.cloud == "torus" & feat.dim == 2 & 
+                       library != "GUDHI Alpha") 
+
+#Point graph; X: Num.points; Y: Runtime; Color: TDA library
+fig.1.mac <- ggplot(data.fig.1.mac, aes(x=num.points, y=avg.time, color=library)) + 
+  geom_point()+
+  geom_errorbar(data.fig.1.mac, mapping = aes(x=num.points, 
+                                          ymin=avg.time - std, 
+                                          ymax=avg.time + std)) +
+  labs(color = "TDA Library",
+       x = "Number of Points on Torus",
+       y = "Average Run Time",
[…]

### 2020-01-24 Figures/Final_Figures_Reproducible.R
All Figure and Data Code has been consolidated to this reproducible script

@@ -1,0 +1,852 @@
+library(dplyr)
+library(ggplot2)
+library(reshape)
+library(ggforce)
+
+####Reading in Data####
+#Reading in the time data
+data.time <- rbind(read.csv("./Cluster_data/annulus.csv", header = TRUE),
+                   read.csv("./Cluster_data/circle.csv", header = TRUE),
+                   read.csv("./Cluster_data/torus.csv", header = TRUE),
+                   read.csv("./Cluster_data/uniform.csv", header = TRUE))
+#Adding Column Names
+colnames(data.time) <- c("row", "measure.type", "point.cloud", "point.cloud.dim", 
+                         "num.points", "feat.dim", "library", 
+                         "time1", "time2", "time3", "time4", 
+                         "time5", "time6", "time7",
+                         "time8", "time9", "time10")
+
+#Remove extraneaous row variable
+data.time <- data.time %>% select(-"row")
+
+#Creating average and standard deviation columns for time data
+data.time$avg.time <- apply(data.time[,7:16],1,mean)
+data.time$std <- apply(data.time[,7:16],1,sd)
+data.time$min.time <- apply(data.time[,7:16],1,min)
+data.time$max.time <- apply(data.time[,7:16],1,max)
+
+#Create Factor labels for libraries
+data.time$library <- factor(data.time$library,
+                            levels = c("stats","Dionysus","GUDHI", "GUDHIalpha"),
+                            labels = c("TDAstats", "Dionysus", "GUDHI Rips", "GUDHI Alpha"))
+
+#Read in object size data
+data.mem <- read.csv("./mem1.csv", header = FALSE)
+
+#Name columns
+colnames(data.mem) <- c("measure.type", "point.cloud", "point.cloud.dim", 
+                        "num.points", "feat.dim", "library", "memory")
+
+#Create Factor labels for libraries
+data.mem$library <- factor(data.mem$library,
+                           levels = c("stats","Dionysus","GUDHI", "GUDHIalpha"),
+                           labels = c("TDAstats", "Dionysus", "GUDHI Rips", "GUDHI Alpha"))
+
+
+####Figure_1####
+#Goal: overview showing speed differences across engines on a canonical shape.
+
+#Selected torus data scanning for two dimensional features
+data.fig.1 <- subset(data.time, point.cloud == "torus" & feat.dim == 2 & 
[…]

### 2020-01-24 Benchmark_Object_Size.R
Renamed Scripts

@@ -16,14 +16,14 @@ vars.circle <- as_tibble(expand.grid(measure = "memory", data.type = "circle",
                                      data.dimensions = 2:4, num.points = seq(50, 500, 25),
                                      feature.dimensions = 1:3, 
                                      TDA.library = c("GUDHI", "GUDHIalpha"),
-                                     num.iteration = 1, file.name = "mem_CCF.csv")) %>% subset(feature.dimensions < data.dimensions)
+                                     num.iteration = 1, file.name = "mem.csv")) %>% subset(feature.dimensions < data.dimensions)
 
 #Making Annulus Variables for time
 vars.noisycircle <- as_tibble(expand.grid(measure = "memory", data.type = "annulus",
                                           data.dimensions = 2:4, num.points = seq(50, 500, 25),
                                           feature.dimensions = 1:3, 
                                           TDA.library = c("GUDHI", "GUDHIalpha"),
-                                          num.iteration = 1, file.name = "mem_CCF.csv")) %>% subset(feature.dimensions < data.dimensions)
+                                          num.iteration = 1, file.name = "mem.csv")) %>% subset(feature.dimensions < data.dimensions)
 
 
 #Making Box Variables for time
@@ -31,14 +31,14 @@ vars.box <- as_tibble(expand.grid(measure = "memory", data.type = "uniform",
                                   data.dimensions = 2:5, num.points = seq(50, 500, 25),
                                   feature.dimensions = 1:4, 
                                   TDA.library = c("GUDHI", "GUDHIalpha"),
-                                  num.iteration = 1, file.name = "mem_CCF.csv")) %>% subset(feature.dimensions < data.dimensions)
+                                  num.iteration = 1, file.name = "mem.csv")) %>% subset(feature.dimensions < data.dimensions)
 
 #Making Torus Variables for time
 vars.torus <- as_tibble(expand.grid(measure = "memory", data.type = "torus",
                                     data.dimensions = 3, num.points = seq(50, 500, 25),
                                     feature.dimensions = 1:2, 
                                     TDA.library = c("GUDHI", "GUDHIalpha"),
-                                    num.iteration = 1, file.name = "mem_CCF.csv")) %>% subset(feature.dimensions < data.dimensions)
[…]

### 2020-01-24 Benchmark_Runtime.R

@@ -16,14 +16,14 @@ vars.circle <- as_tibble(expand.grid(measure = "time", data.type = "circle",
                                      data.dimensions = 2:4, num.points = seq(10, 500, 5),
                                      feature.dimensions = 1:3, 
                                      TDA.library = c("stats", "Dionysus", "GUDHI", "GUDHIalpha"),
-                                     num.iteration = 10, file.name = "timeccf.csv")) %>% subset(feature.dimensions < data.dimensions)
+                                     num.iteration = 10, file.name = "time.csv")) %>% subset(feature.dimensions < data.dimensions)
 
 #Making Annulus Variables for time
 vars.noisycircle <- as_tibble(expand.grid(measure = "time", data.type = "annulus",
                                           data.dimensions = 2:4, num.points = seq(10, 500, 5),
                                           feature.dimensions = 1:3, 
                                           TDA.library = c("stats", "Dionysus", "GUDHI", "GUDHIalpha"),
-                                          num.iteration = 10, file.name = "timeccf.csv")) %>% subset(feature.dimensions < data.dimensions)
+                                          num.iteration = 10, file.name = "time.csv")) %>% subset(feature.dimensions < data.dimensions)
 
 
 #Making Box Variables for time
@@ -31,14 +31,14 @@ vars.box <- as_tibble(expand.grid(measure = "time", data.type = "uniform",
                                   data.dimensions = 2:8, num.points = seq(10, 500, 5),
                                   feature.dimensions = 1:7, 
                                   TDA.library = c("stats", "Dionysus", "GUDHI", "GUDHIalpha"),
-                                  num.iteration = 10, file.name = "timeccf.csv")) %>% subset(feature.dimensions < data.dimensions)
+                                  num.iteration = 10, file.name = "time.csv")) %>% subset(feature.dimensions < data.dimensions)
 
 #Making Torus Variables for time
 vars.torus <- as_tibble(expand.grid(measure = "time", data.type = "torus",
                                     data.dimensions = 3, num.points = seq(10, 500, 5),
                                     feature.dimensions = 1:2, 
                                     TDA.library = c("stats", "Dionysus", "GUDHI", "GUDHIalpha"),
-                                    num.iteration = 10, file.name = "timeccf.csv")) %>% subset(feature.dimensions < data.dimensions)
[…]

### 2020-01-24 Figures/Extra_Mac_Figures.R
No update, just reran the figure code

@@ -13,7 +13,7 @@ library(ggforce)
 #Figure 5 was left out since object size is invariant and always the same  
 
 #Reading in the time data
-data.time.mac <- read.csv("./time_mac_desktop.csv", header = FALSE)
+data.time.mac <- read.csv("./Mac_data/time_mac_desktop.csv", header = FALSE)
 #Adding Column Names
 colnames(data.time.mac) <- c("measure.type", "point.cloud", "point.cloud.dim", 
                          "num.points", "feat.dim", "library", 

### 2020-01-24 Enclosing_Radius_Improvement.R
Not used anywhere in the project

@@ -1,103 +1,0 @@
-library(plyr)
-library(readr)
-library(dplyr)
-library(TDA)
-library(TDAstats)
-library(bench)
-library(pryr)
-library(ggplot2)
-
-source("Functions.R")
-
-#####Function to make enclosing radius for TDA GUDHI and Dionysus (makes runtime fairer)#####
-enclosing_radius <- function(X){
-  # function which finds radius beyond which no homology changes
-  # X is a point cloud data frame
-  d = dist(X)
-  n = nrow(X)
-  return(max(unlist(lapply(X = 1:(nrow(X) - 1), FUN = function(X){ return(min(d[(1+(X-1)*n-(X-1)*X/2):(X*n-X*(X+1)/2)])) }))))
-}
-
-#####alternate bench function
-bench.enc <- function(pointdata, TDA.library, featdim, num.iterations) {
-  rad <- enclosing_radius(pointdata)
-  str.lib <- paste(TDA.library)
-  # TDAstats
-  if (TDA.library == "stats") {
-    time <- mark(calculate_homology(pointdata, dim = featdim, threshold = rad),
-                 iterations = num.iterations)
-    # TDA - Dionysus
-  } else if (TDA.library == "Dionysus") {
-    time <- mark(ripsDiag(pointdata, maxdimension = featdim, maxscale = rad,
-                          location = FALSE, library = "Dionysus"),
-                 iterations = num.iterations)
-    # TDA - GUDHI
-  } else if (TDA.library == "GUDHI") {
-    time <- mark(ripsDiag(pointdata, maxdimension = featdim, maxscale = rad,
-                          location = FALSE, library = "GUDHI"),
-                 iterations = num.iterations)
-    # TDA - GUDHI (alpha complex)
-  } else if (TDA.library == "GUDHIalpha") {
-    time <- mark(alphaComplexDiag(pointdata, maxdimension = featdim,
-                                  location = FALSE, library = "GUDHI"),
-                 iterations = num.iterations)
-    # none of the above
-  } else {
-    stop("Invalid TDA engine selected")
-  }
-  
-  # row 1 column 12 is list of times
-  time.list <- time[1,12]
-  time.runs <- unlist(time.list[[1]])
-  appnd <- tibble()
-  row <- c(rad, time.runs, str.lib)
-  row.appnd <- rbind(appnd, row)
-  colnames(row.appnd) <- seq(1:ncol(row.appnd))
-  return(row.appnd)
-}
-
-
[…]

### 2020-01-24 Supplemental_Regression/Regression_analysis.R
Relocated to appropriate folders

@@ -4,9 +4,75 @@ library(reshape)
 library(ggforce)
 library(knitr)
 
-source("./Figures/Creating_ggplot_Figures.R")
+####Reading in Data####
+#Reading in the time data
+data.time <- rbind(read.csv("./Cluster_data/annulus.csv", header = TRUE),
+                   read.csv("./Cluster_data/circle.csv", header = TRUE),
+                   read.csv("./Cluster_data/torus.csv", header = TRUE),
+                   read.csv("./Cluster_data/uniform.csv", header = TRUE))
+#Adding Column Names
+colnames(data.time) <- c("row", "measure.type", "point.cloud", "point.cloud.dim", 
+                         "num.points", "feat.dim", "library", 
+                         "time1", "time2", "time3", "time4", 
+                         "time5", "time6", "time7",
+                         "time8", "time9", "time10")
+
+#Remove extraneaous row variable
+data.time <- data.time %>% select(-"row")
+
+#Creating average and standard deviation columns for time data
+data.time$avg.time <- apply(data.time[,7:16],1,mean)
+data.time$std <- apply(data.time[,7:16],1,sd)
+data.time$min.time <- apply(data.time[,7:16],1,min)
+data.time$max.time <- apply(data.time[,7:16],1,max)
+
+#Create Factor labels for libraries
+data.time$library <- factor(data.time$library,
+                            levels = c("stats","Dionysus","GUDHI", "GUDHIalpha"),
+                            labels = c("TDAstats", "Dionysus", "GUDHI Rips", "GUDHI Alpha"))
+
+#Read in object size data
+data.mem <- read.csv("./mem1.csv", header = FALSE)
+
+#Name columns
+colnames(data.mem) <- c("measure.type", "point.cloud", "point.cloud.dim", 
+                        "num.points", "feat.dim", "library", "memory")
+
+#Create Factor labels for libraries
+data.mem$library <- factor(data.mem$library,
+                           levels = c("stats","Dionysus","GUDHI", "GUDHIalpha"),
+                           labels = c("TDAstats", "Dionysus", "GUDHI Rips", "GUDHI Alpha"))
+
+####Making Data Subsets####
+#Data subsets identical to the data subset for each figure
[…]

### 2020-01-24 tdabench-paper/tdabench.Rmd
Beefed up captions and modified figure size

@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/Intro_Rips.png}
-  \caption{As $\delta$ continues to grow, more circles intersect filling out the simplicial complex.
+  \caption{Basic Visualization of the Rips Complex.
+For a given parameter, $\delta$, $\delta$ diameter circles are drawn around each point.
+If two circles intersect, a point is drawn between their centers.
+As $\delta$ continues to grow, more circles intersect filling out the simplicial complex.
 Features on the simplicial complex appear and die as $\delta$ increases.
+These features' dimensions, birth, and death are recorded in an n x 3 matrix.
 Eventually, the full convex hull is drawn ending the "filtration" process.}
@@
   \centering
-  \includegraphics[height=2.5in]{../Figures/Final_Figures/Intro_Alpha.png}
-  \caption{As $\alpha$ continues to grow, more circles fill out the Voronoi cells.
-Once two circles in adjacent Voronoi cells fill out the cell, a connecting segment can be drawn.
+  \includegraphics[height=1.5in]{../Figures/Final_Figures/Intro_Alpha.png}
+  \caption{Basic visualization of the Alpha complex.
+For a given $\alpha$, $\alpha$ radius balls are drawn around each point, and the union is taken.
+Then an intersection between this union of $\alpha$ balls and the Vornoi diagram is taken.
+A connecting segment is drawn between points in adjacent voronoi cells once the $\alpha$ ball "fills" out the Voronoi diagram.
+As $\alpha$ continues to grow, more circles fill out the Voronoi cells.
 Eventually, once $\alpha$ is large enough, the Delaunay Triangulation is formed.}
@@
 The x-axis represents the number of points in a 3D torus point cloud, and y-axis represents average run time.
-For these given parameters, the Rips libraries show a polynomial growth curve for increasing point dimension.
-Further details about regression parameters is found in supplemental material.}
+Each library that calculated the Rips complex is differentiated by point color.
+All calculations calculated up to two dimensional features.
+Average runtime is presented as mean +/- standard deviation over 10 iterations.
+ALl three Rips libraries exhibit a polynomial time growth for these parameters.
+Further details about regression parameters is found in supplemental material on the paper Github.
+Some libraries are not calculated for all points due to computational limitations, which are described more in the Computational Limitations section.}
 \label{fig:tor}
@@
 For the conditions tested in our analysis, all Rips libraries resulted in an apparent polynomial growth for run time. 
-The exact values for the log-log regression of this growth pattern is provided in supplemental data.
+The exact values for the log-log regression of this growth pattern is provided in supplemental data on the paper Github.
 
@@
 
-\begin{figure}
+\begin{widefigure}
   \centering
-  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig2.png}
+  \includegraphics[height=2in]{../Figures/Final_Figures/fig2.png}
   \caption{Runtime of persistent homology calculation as a function of underlying engine for different dimension circles.
-Panel 1.
+The x-axis represents the number of points in the point cloud, and y-axis represents average run time.
+Each library that calculated the Rips complex is differentiated by point color.
+All calculations calculated up to two dimensional features.Panel 1.
 Comparing persistent homology calculation run times for a 2D circle.
@@
 Comparing persistent homology calculation run times for a 4D hypersphere.
-Maximum feature dimensions were calculated for each point cloud.}
+Maximum feature dimensions were calculated for each point cloud.
+Average runtime is presented as mean +/- standard deviation over 10 iterations.}
 \label{fig:cir}
-\end{figure}
+\end{widefigure}
 
@@
 
-\begin{figure}
+\begin{widefigure}
   \centering
-  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig3.png}
+  \includegraphics[height=2in]{../Figures/Final_Figures/fig3.png}
   \caption{Comparison of Rips complex persistent homology calculation as a function of feature dimension.
+The x-axis represents the number of features in the point cloud, and y-axis represents average run time.
+Point cloud number was held constant and indicated by color.
+All Rips complexes were calculated on an 8 dimensional box.
 Panel 1.
@@
 Comparing run times for GUDHI.
-Limited point clouds were tested for high dimensional features due to computational limits.}
+Limited point clouds were tested for high dimensional features due to computational limits.
+Average runtime is presented as mean +/- standard deviation over 10 iterations.}
 \label{fig:box}
-\end{figure}
+\end{widefigure}
 
@@
 Another key difference is that Alpha complex runtime seems to have an apparent linear rise in the conditions for our analysis. 
-We provide the linear regression coefficients in the supplemental data.
+We provide the linear regression coefficients in the supplemental data on the paper Github.
 
@@
   \centering
-  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig4.png}
+  \includegraphics[height=2in]{../Figures/Final_Figures/fig4.png}
   \caption{Comparing homology calculation as a function of point cloud dimension for Rips vs Alpha complexes.
+The x-axis represents the number of points in the point cloud, and y-axis represents average run time.
+Different dimensional annulus point clouds are indicaed by color.
 Panel 1 = Comparing run times using GUDHI's Rips complex library.
@@
 For our parameters, Alpha complex run times are more linear in contrast to Rips more polynomial run times.
-Exact regression data is provided in Supplemental material.}
+Exact regression data is provided in supplemental material on the paper Github.
+Average runtime is presented as mean +/- standard deviation over 10 iterations.}
 \label{fig:ann}
@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/fig5a.png}
-  \caption{Persistent Homology Boundary Matrix for Rips vs Alpha complex. Shape of used point cloud is demonstrated by each image. Rips complex vs Alpha complex for object size.}
+  \caption{Persistent Homology Boundary Matrix for Rips vs Alpha complex.
+The x-axis represents the number of points in the point cloud, and y-axis represents object size.
+The type of complex is indicated by color.
+Maximum feature dimensions were calculated.
+Panel 1 = Comparing object size on 3D annulus.
+Panel 2 = Comparing object size on sphere.
+Panel 3 = Comparing object size on torus.
+Panel 4 = Comparing object size on 3D box.
+Similar to runtime growth, Alpha complex showed a linear pattern and Rips complex showed a log-log linear pattern.
+Regression results can be found in supplemental data on the paper Github}
 \label{fig:mem1}
@@
 
-\begin{figure}
+\begin{widefigure}
   \centering
@@
   \caption{Different factors change boundary matrix size for each complex.
+The x-axis represents the number of points in the point cloud, and y-axis represents object size.
 Panel 1 = Rips complex object size depends on feature size and point cloud number alone.
+For Rips complex, the decreasing number of data points for each calculated feature dimension reflects the computational limits of large point cloud sizes.
+Point cloud shape is not defined as it had no impact on boundary matrix size.
 Panel 2 = Alpha complex object size depends on feature size, point cloud number, and point cloud shape.
-Point cloud shape is not defined as it had no impact on boundary matrix size.
-For a given n point cloud, n-1 dimension features were calculated in these examples.}
+Point cloud shape was indicated as it impacted object size.}
 \label{fig:mem2}
-\end{figure}
+\end{widefigure}
 
@@
   \caption{TDAstats Rips complexes vs GUDHI Alpha complex average run times across all 3D shapes for 2D features.
+The x-axis represents the number of points in the point cloud, and y-axis represents average run time.
+The type of complex is indicated by color.
 Panel 1 = Execution time on annulus.
@@
 Panel 3 = Execution time on torus.
-Panel 4 = Execution time on uniform box}
+Panel 4 = Execution time on uniform box.
+Data is presented as mean +/- standard deviation}
 \label{fig:fin}

### 2020-01-26 tdabench-paper/tdabench.Rmd
Added Abstract

@@
 abstract: >
-  An abstract of less than 150 words.
+  With the increasing popularity of topological data analysis, several persistent homology software libraries have been implemented in R.
+The Dionysus and GUDHI C++ libraries have been implemented in the TDA package, and the Ripser C++ library has been implemented in the TDAstats package.
+All libraries feature Rips complex functionality.
+Additionally, within the TDA GUDHI package, the Alpha complex is available.
+These software represent powerful tools for analyzing the underlying topological structure of data.
+However, they are limited by computational restraints.
+High dimensional features and large point cloud sizes are limiting factors for persistent homology calculations.
+These limitations and guidelines for which persistent homology libraries to use have not been specified in R.
+To fill this gap, we benchmarked these libraries on a High-Powered Cluster.
+We benchmarked all Rips complex libraries for runtime.
+In addition, we compared Alpha and Rips for memory burden and runtime.
+We find that Alpha complexes run significantly faster than all Rips complex libraries for all data sets where data dimension is less than or equal to three.
+Alpha complexes were not computable on our systems for greater than 3 dimensions.
+For Rips complex calculations, run time performance was fastest for TDAstats and slowest for Dionysus.
+This difference was markedly clear at high dimensional features and point cloud values.
+From a computational efficiency perspective, if data dimension is less than or equal to 3, we recommend performing persistent homology with GUDHI Alpha complexes.
+If data dimension is higher, we recommend using TDAstats Rips Complex.
 output:

### 2020-01-26 tdabench-paper/tdabench.Rmd
Really struggling with the 150 word count

--- 2020-01-24 tdabench-paper/tdabench.Rmd
@@
 abstract: >
-  An abstract of less than 150 words.
+  Several persistent homology software libraries have been implemented in R.
+The Dionysus (TDA), GUDHI (TDA), and Ripser C++ libraries (TDAstats) provide Rips complex functionality.
+GUDHI  (TDA package) also provides Alpha complex functionality.
+These software represent powerful analysis tools but are limited computationally.
+These limitations and guidelines for which persistent homology libraries to use have not been specified in R.
+We benchmarked the all libraries on a High-Powered Cluster for runtime.
+We also compared Alpha and Rips for memory burden.
+Alpha complexes run significantly faster than all Rips complex libraries for all data sets for data dimension less than or equal to three.
+Alpha complexes were not computable for data dimension greater than 3.
+TDAstats had the fastest runtimes for all Rips complexes.
+We recommend performing persistent homology with GUDHI Alpha complexes for data dimension less than or equal to three.
+For higher dimension, we recommend TDAstats Rips Complex.
 output:

### 2020-01-31 tdabench-paper/tdabench.Rmd
removed unnecessary R specifier

@@
 The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after considering available computational resources.
-For consistency between software libraries, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} R packages.
+For consistency between software libraries, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} packages.
 Within the \CRANpkg{TDA} package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the \CRANpkg{TDAstats} package, benchmark data was collected for the Ripser \citep{Ripser} library.

### 2020-01-31 tdabench-paper/tdabench.Rmd
uncapitalized alpha complex
improved wording

@@
 Within the \CRANpkg{TDA} package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the \CRANpkg{TDAstats} package, benchmark data was collected for the Ripser \citep{Ripser} library.
-As Alpha complex calculation was only implemented in GUDHI, Alpha complex benchmark data was naturally only collected for the single library.
-Measuring memory usage proved challenging since all the libraries calculating persistent homology were implemented in another programming language and then wrapped in R.
+As alpha complex calculation was only implemented in GUDHI, alpha complex benchmark data was naturally only collected for the single library.
+Measuring memory usage proved challenging since all the libraries calculating persistent homology were implemented in either C++ or Java and then wrapped in R as part of a CRAN package.
 Thus, memory burden was indirectly measured by using boundary matrix size as a proxy.

### 2020-01-31 tdabench-paper/tdabench.Rmd
improved wording

@@
 Thus, memory burden was indirectly measured by using boundary matrix size as a proxy.
-Given that Ripser optimizes persistent homology calculation by avoiding calculation of a boundary matrix, memory use benchmarks are not provided for Ripser and, consequently, \CRANpkg{TDAstats}.
+Given that Ripser optimizes computation of persistent homology by avoiding calculation of a boundary matrix, memory use benchmarks are not provided for Ripser and, consequently, \CRANpkg{TDAstats}.
 Fully reproducible code for all numerical results and figures in this report can be found at https://github.com/eashwarsoma/TDA-benchmark.

### 2020-01-31 tdabench-paper/tdabench.Rmd
moved part of computational limitations to methods

@@
 Given that Ripser optimizes computation of persistent homology by avoiding calculation of a boundary matrix, memory use benchmarks are not provided for Ripser and, consequently, \CRANpkg{TDAstats}.
+
+Benchmark data was collected twice - once on a local machine and once on a remote computing node, each of which featured 16 GB random access memory.
+Both datasets were compared for consistency and are publicly available at the repository linked below.
+Data from the remote computing node is visualized in this report.
+The larger point clouds required more than 16 GB of RAM to calculate persistent homology using a subset of the libraries; attempts to compute results resulted in runtime errors, and the corresponding output is missing from the corresponding figures and tables.
 Fully reproducible code for all numerical results and figures in this report can be found at https://github.com/eashwarsoma/TDA-benchmark.

### 2020-01-31 tdabench-paper/tdabench.Rmd
removed unnecessary subsection (was more methods than results)

@@
 \section{Results}
-
-\subsection{Runtime comparison between packages for calculation of V-R complex}
-In our analysis, we compare runtime performances of the three Rips complexes libraries for persistence homology calculation under different conditions.
-We then focus our analysis on comparing Rips complexes with Alpha complexes using the GUDHI package.
 

### 2020-01-31 tdabench-paper/tdabench.Rmd
shortened wording
shortened sentence
removed unnecessary subheading, shortened text

@@
 
-\subsubsection{Prototype/Canonical dataset (Torus)}
-The torus point cloud is a classical shape used to demonstrate persistent homology calculations. 
-A persistent homology calculation of a uniformly distributed torus would reveal a large 1 dimensional feature representing the middle circle of the torus.
+Computing persistent homology of a canonical torus grants quick insight into efficiency of each library (Figure \ref{fig:tor}).
+Calculating persistent homology of points distributed on a torus should reveal a single 1-dimensional feature representing the middle circle of the torus.
 There would also be a large 2 dimensional feature representing the inner cavity of the torus.

### 2020-01-31 tdabench-paper/tdabench.Rmd
removed info not seen in figure
removed duplicate sentence

@@
 Computing persistent homology of a canonical torus grants quick insight into efficiency of each library (Figure \ref{fig:tor}).
-Calculating persistent homology of points distributed on a torus should reveal a single 1-dimensional feature representing the middle circle of the torus.
-There would also be a large 2 dimensional feature representing the inner cavity of the torus.
-In figure \ref{fig:tor}, we calculated up to 2 dimensional features of torus 3d points cloud for an increasing number of points across all TDA libraries.
 

### 2020-01-31 tdabench-paper/tdabench.Rmd
added narrative for torus figure
revised torus figure caption

@@
 Computing persistent homology of a canonical torus grants quick insight into efficiency of each library (Figure \ref{fig:tor}).
+Dionysus exhibits the longest average runtime, and, although Ripser and GUDHI have similar runtimes for smaller point clouds, as the number of points increases Ripser eventually has a significant lead.
 
@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/fig1.png}
-  \caption{Comparison of Rips complex persistent homology calculation runtimes across R packages for a torus.
-The x-axis represents the number of points in a 3D torus point cloud, and y-axis represents average run time.
-Each library that calculated the Rips complex is differentiated by point color.
-All calculations calculated up to two dimensional features.
-Average runtime is presented as mean +/- standard deviation over 10 iterations.
-ALl three Rips libraries exhibit a polynomial time growth for these parameters.
-Further details about regression parameters is found in supplemental material on the paper Github.
-Some libraries are not calculated for all points due to computational limitations, which are described more in the Computational Limitations section.}
+  \caption{\textbf{Calculating persistent homology of a torus with three TDA libraries.}
+  Average runtime (mean $\pm$ SD) for each TDA library (denoted by color) is plotted against point cloud size.
+  Homological features up to 2 dimensions were calculated.
+  Time complexity follows a power law for all three libraries (see GitHub repo for regression details).
+  Although the libraries have similar runtimes for smaller point clouds, Dionysus has a clear disadvantage when the number of points exceeds 100.
+  When the number of points exceeds 200, Ripser has a clear advantage over GUDHI, which maintains its advantage over Dionysus.
+  }
 \label{fig:tor}

### 2020-01-31 Figures/Final_Figures_Reproducible.R
Corrected figures so that shape and color change together. Modified figure legends slightly

@@ -28,7 +28,8 @@ data.time$max.time <- apply(data.time[,7:16],1,max)
 #Create Factor labels for libraries
 data.time$library <- factor(data.time$library,
                             levels = c("stats","Dionysus","GUDHI", "GUDHIalpha"),
-                            labels = c("TDAstats", "Dionysus", "GUDHI Rips", "GUDHI Alpha"))
+                            labels = c("TDAstats (Ripser)", 
+                                       "TDA (Dionysus)", "TDA (GUDHI Rips)", "TDA (GUDHI Alpha)"))
 
 #Read in object size data
 data.mem <- read.csv("./mem1.csv", header = FALSE)
@@ -40,7 +41,8 @@ colnames(data.mem) <- c("measure.type", "point.cloud", "point.cloud.dim",
 #Create Factor labels for libraries
 data.mem$library <- factor(data.mem$library,
                            levels = c("stats","Dionysus","GUDHI", "GUDHIalpha"),
-                           labels = c("TDAstats", "Dionysus", "GUDHI Rips", "GUDHI Alpha"))
+                           labels = c("TDAstats (Ripser)", 
+                                      "TDA (Dionysus)", "TDA (GUDHI Rips)", "TDA (GUDHI Alpha)"))
 
 
 ####Figure_1####
@@ -48,15 +50,16 @@ data.mem$library <- factor(data.mem$library,
 
 #Selected torus data scanning for two dimensional features
 data.fig.1 <- subset(data.time, point.cloud == "torus" & feat.dim == 2 & 
-                       library != "GUDHI Alpha") 
+                       library != "TDA (GUDHI Alpha)") 
 
 #Point graph; X: Num.points; Y: Runtime; Color: TDA library
-fig.1 <- ggplot(data.fig.1, aes(x=num.points, y=avg.time, color=library)) + 
+fig.1 <- ggplot(data.fig.1, aes(x=num.points, y=avg.time, color=library, shape=library)) + 
   geom_point()+
   geom_errorbar(data.fig.1, mapping = aes(x=num.points, 
                                           ymin=avg.time - std, 
                                           ymax=avg.time + std)) +
   labs(color = "TDA Library",
+       shape = "TDA Library",
        x = "Number of Points on Torus",
        y = "Average Run Time",
[…]

### 2020-01-31 tdabench-paper/tdabench.Rmd
Corrected minor fact: Toruses have two 1D features

--- 2020-01-31 tdabench-paper/tdabench.Rmd
@@
 The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after considering available computational resources.
-For consistency between software libraries, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} packages.
+For consistency between software libraries, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} R packages.
 Within the \CRANpkg{TDA} package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the \CRANpkg{TDAstats} package, benchmark data was collected for the Ripser \citep{Ripser} library.
-As alpha complex calculation was only implemented in GUDHI, alpha complex benchmark data was naturally only collected for the single library.
-Measuring memory usage proved challenging since all the libraries calculating persistent homology were implemented in either C++ or Java and then wrapped in R as part of a CRAN package.
+As Alpha complex calculation was only implemented in GUDHI, Alpha complex benchmark data was naturally only collected for the single library.
+Measuring memory usage proved challenging since all the libraries calculating persistent homology were implemented in another programming language and then wrapped in R.
 Thus, memory burden was indirectly measured by using boundary matrix size as a proxy.
-Given that Ripser optimizes computation of persistent homology by avoiding calculation of a boundary matrix, memory use benchmarks are not provided for Ripser and, consequently, \CRANpkg{TDAstats}.
-
-Benchmark data was collected twice - once on a local machine and once on a remote computing node, each of which featured 16 GB random access memory.
-Both datasets were compared for consistency and are publicly available at the repository linked below.
-Data from the remote computing node is visualized in this report.
-The larger point clouds required more than 16 GB of RAM to calculate persistent homology using a subset of the libraries; attempts to compute results resulted in runtime errors, and the corresponding output is missing from the corresponding figures and tables.
+Given that Ripser optimizes persistent homology calculation by avoiding calculation of a boundary matrix, memory use benchmarks are not provided for Ripser and, consequently, \CRANpkg{TDAstats}.
 Fully reproducible code for all numerical results and figures in this report can be found at https://github.com/eashwarsoma/TDA-benchmark.
@@
 
-Computing persistent homology of a canonical torus grants quick insight into efficiency of each library (Figure \ref{fig:tor}).
-Calculating persistent homology of points distributed on a torus should reveal a single 1-dimensional feature representing the middle circle of the torus.
+\subsection{Runtime comparison between packages for calculation of V-R complex}
+In our analysis, we compare runtime performances of the three Rips complexes libraries for persistence homology calculation under different conditions.
+We then focus our analysis on comparing Rips complexes with Alpha complexes using the GUDHI package.
+
+\subsubsection{Prototype/Canonical dataset (Torus)}
+The torus point cloud is a classical shape used to demonstrate persistent homology calculations. 
+A persistent homology calculation of a uniformly distributed torus would reveal two large 1 dimensional features, which are the two circles defined by the major and minor radius.
 There would also be a large 2 dimensional feature representing the inner cavity of the torus.

### 2020-02-20 tdabench-paper/tdabench.Rmd
resolved remaining merge conflict and shortened initial paragraph

--- 2020-01-31 tdabench-paper/tdabench.Rmd
@@
 The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after considering available computational resources.
-For consistency between software libraries, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} packages.
+For consistency between software libraries, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} R packages.
 Within the \CRANpkg{TDA} package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the \CRANpkg{TDAstats} package, benchmark data was collected for the Ripser \citep{Ripser} library.
@@
 Dionysus exhibits the longest average runtime, and, although Ripser and GUDHI have similar runtimes for smaller point clouds, as the number of points increases Ripser eventually has a significant lead.
+Next, we compare performance of these three libraries with multiple canonical datasets to ensure that the noted pattern generalizes.
 

### 2020-02-20 tdabench-paper/tdabench.Rmd
shortened phrasing
removed unnecessary phrase

@@
 The larger point clouds required more than 16 GB of RAM to calculate persistent homology using a subset of the libraries; attempts to compute results resulted in runtime errors, and the corresponding output is missing from the corresponding figures and tables.
-Fully reproducible code for all numerical results and figures in this report can be found at https://github.com/eashwarsoma/TDA-benchmark.
+Fully reproducible code for all numerical results and figures can be found at https://github.com/eashwarsoma/TDA-benchmark.
 
@@
 Dionysus exhibits the longest average runtime, and, although Ripser and GUDHI have similar runtimes for smaller point clouds, as the number of points increases Ripser eventually has a significant lead.
-Next, we compare performance of these three libraries with multiple canonical datasets to ensure that the noted pattern generalizes.
+Next, we compare library performance with multiple canonical datasets to ensure that the noted pattern generalizes.
 

### 2020-02-20 tdabench-paper/tdabench.Rmd
edited narrative for fig cir

@@
 
-\subsubsection{Runtime as a function of data dimension (feature = n - 1)}
-For small point cloud numbers, Gudhi and TDAstats performed similarly while Dionysus lagged behind. 
-At larger point clouds, TDAstats outperforms Gudhi, and the difference among the Rips libraries is more visible.
-This result is especially true for calculating higher feature dimensions on large point clouds. 
-While the runtime values changed for each point cloud condition, the growth pattern with respect to increasing number of points was similar for all Rips complex libraries. 
-For the conditions tested in our analysis, all Rips libraries resulted in an apparent polynomial growth for run time. 
-The exact values for the log-log regression of this growth pattern is provided in supplemental data on the paper Github.
+Tori do not trivially generalize to dimensions other than 3, but circles do.
+Benchmarking on a circular point cloud permits confirmation of the pattern in Figure \ref{fig:tor} while also revealing how the libraries compare as dataset dimension increases.
+Figure \ref{fig:cir} exhibits the resulting data for a 2-dimensional circle (panel A), 3-dimensional sphere (panel B), and 4-dimensional hypersphere (panel C).
+When the dataset dimension equals 2, GUDHI practically matches Ripser's performance in outpacing Dionysus.
+However, in the case of the 3-dimensional sphere, the pattern visualized in Figure \ref{fig:tor} for the 3-dimensional torus is again present.
+By the 4th dimension, the gap between Ripser and GUDHI widens.
+Of note, missing points for larger datasets in Figure \ref{fig:cir} are not plotted if and only if calculating persistent homology caused an error due to insufficient RAM.
+Thus, for the hypersphere, Ripser was able to calculate persistent homology for a dataset with approximately 3 times as many points as Dionysus and over 2 times as many as GUDHI.
+Interestingly, all curves plotted in Figure \ref{fig:cir} grow polynomially with respect to the number of points (see Supplement for regression details).
 
-Each panel in figure \ref{fig:cir} corresponds to an n-dimensional circle (circle, sphere, and hypersphere). 
-For each point cloud, n-1 dimensional features were calculated.
-As the sphere dimensions and feature dimensions increased, the polynomial time growth becomes apparent sooner. 
-At these higher dimensional features, some of the curves seem to cut off. 
-This is due to certain point cloud sizes not being computable on our devices. 
-These limitations are laid out in the "Computation Limits" section.
+For each point cloud, n-1 dimensional features were calculated. *previous sentence will be moved to figure caption*
 

### 2020-02-20 tdabench-paper/tdabench.Rmd
fixed indentation/newlines

@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/fig1.png}
-  \caption{\textbf{Calculating persistent homology of a torus with three TDA libraries.}
+  \caption{
+    \textbf{Calculating persistent homology of a torus with three TDA libraries.}
     Average runtime (mean $\pm$ SD) for each TDA library (denoted by color) is plotted against point cloud size.

### 2020-02-20 tdabench-paper/tdabench.Rmd
revised caption for fig cir

@@
     \textbf{Calculating persistent homology of a torus with three TDA libraries.}
-    Average runtime (mean $\pm$ SD) for each TDA library (denoted by color) is plotted against point cloud size.
+    Average runtime (mean $\pm$ SD, $n=10$) for each TDA library (denoted by color) is plotted against point cloud size.
     Homological features up to 2 dimensions were calculated.
@@
 
-For each point cloud, n-1 dimensional features were calculated. *previous sentence will be moved to figure caption*
-
 \begin{widefigure}
@@
   \includegraphics[height=2in]{../Figures/Final_Figures/fig2.png}
-  \caption{Runtime of persistent homology calculation as a function of underlying engine for different dimension circles.
+  \caption{
+    \textbf{Calculating persistent homology of round point clouds of varying dimensions with three TDA libraries.}
+    Runtime of persistent homology calculation as a function of underlying engine for different dimension circles.
     The x-axis represents the number of points in the point cloud, and y-axis represents average run time.
     Each library that calculated the Rips complex is differentiated by point color.
-All calculations calculated up to two dimensional features.Panel 1.
-Comparing persistent homology calculation run times for a 2D circle.
-Panel 2.
-Comparing persistent homology calculation run times for a 3D sphere.
-Panel 3.
-Comparing persistent homology calculation run times for a 4D hypersphere.
-Maximum feature dimensions were calculated for each point cloud.
-Average runtime is presented as mean +/- standard deviation over 10 iterations.}
+    All calculations calculated up to two dimensional features.
+    Panel 1 compares library performance for a 2-dimensional circular point cloud; panel 2 for a 3-dimensional spherical point cloud; and panel 3 for a 4-dimensional hyperspherical point cloud.
+    Maximum feature dimensions (one less than data dimension) were calculated for each point cloud.
+    Average runtime is presented as mean $\pm$ SD over 10 iterations.
+  }
 \label{fig:cir}

### 2020-02-20 tdabench-paper/tdabench.Rmd
fixed phrasing
revised fig cir caption

@@
 
-Tori do not trivially generalize to dimensions other than 3, but circles do.
+Tori do not trivially generalize to other dimensions, but circles do.
 Benchmarking on a circular point cloud permits confirmation of the pattern in Figure \ref{fig:tor} while also revealing how the libraries compare as dataset dimension increases.
@@
     \textbf{Calculating persistent homology of round point clouds of varying dimensions with three TDA libraries.}
-    Runtime of persistent homology calculation as a function of underlying engine for different dimension circles.
-    The x-axis represents the number of points in the point cloud, and y-axis represents average run time.
-    Each library that calculated the Rips complex is differentiated by point color.
-    All calculations calculated up to two dimensional features.
+    Average runtime (mean $\pm$ SD, $n=10$) for each TDA library (denoted by color) is plotted against point cloud size and faceted by data dimension.
     Panel 1 compares library performance for a 2-dimensional circular point cloud; panel 2 for a 3-dimensional spherical point cloud; and panel 3 for a 4-dimensional hyperspherical point cloud.
-    Maximum feature dimensions (one less than data dimension) were calculated for each point cloud.
-    Average runtime is presented as mean $\pm$ SD over 10 iterations.
+    Maximum feature dimensions (one less than data dimension) were calculated in each case.
   }

### 2020-02-20 tdabench-paper/tdabench.Rmd
updated panel labels

@@
 Benchmarking on a circular point cloud permits confirmation of the pattern in Figure \ref{fig:tor} while also revealing how the libraries compare as dataset dimension increases.
-Figure \ref{fig:cir} exhibits the resulting data for a 2-dimensional circle (panel A), 3-dimensional sphere (panel B), and 4-dimensional hypersphere (panel C).
+Figure \ref{fig:cir} exhibits the resulting data for a 2-dimensional circle (left panel), 3-dimensional sphere (center panel), and 4-dimensional hypersphere (right panel).
 When the dataset dimension equals 2, GUDHI practically matches Ripser's performance in outpacing Dionysus.
@@
     Average runtime (mean $\pm$ SD, $n=10$) for each TDA library (denoted by color) is plotted against point cloud size and faceted by data dimension.
-    Panel 1 compares library performance for a 2-dimensional circular point cloud; panel 2 for a 3-dimensional spherical point cloud; and panel 3 for a 4-dimensional hyperspherical point cloud.
+    Left panel compares library performance for a 2-dimensional circular point cloud; center panel for a 3-dimensional spherical point cloud; and right panel for a 4-dimensional hyperspherical point cloud.
     Maximum feature dimensions (one less than data dimension) were calculated in each case.

### 2020-03-08 tdabench-paper/tdabench.Rmd
caption edits

@@
     \textbf{Calculating persistent homology of a torus with three TDA libraries.}
-    Average runtime (mean $\pm$ SD, $n=10$) for each TDA library (denoted by color) is plotted against point cloud size.
+    Average runtime (mean $\pm$ SD, $n=10$ iterations per data point) for each TDA library (denoted by color) is plotted against point cloud size.
     Homological features up to 2 dimensions were calculated.
@@
     \textbf{Calculating persistent homology of round point clouds of varying dimensions with three TDA libraries.}
-    Average runtime (mean $\pm$ SD, $n=10$) for each TDA library (denoted by color) is plotted against point cloud size and faceted by data dimension.
+    Average runtime (mean $\pm$ SD, $n=10$ iterations per data point) for each TDA library (denoted by color) is plotted against point cloud size and faceted by data dimension.
     Left panel compares library performance for a 2-dimensional circular point cloud; center panel for a 3-dimensional spherical point cloud; and right panel for a 4-dimensional hyperspherical point cloud.

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised text

@@
 
-\subsubsection{Runtime as a function of feature dimension}
+<!--FOLLOWING SENTENCES SHOULD GO IN DISCUSSION
 For an n-dimensional point cloud, up to n-1 feature dimensions can be calculated. 
@@
 Perhaps this is due to difficulty of conceptualizing what high dimensional features represent. 
-However, it might also be due to computational restrictions of calculating high dimensional features. 
-For Figure \ref{fig:box}, we restricted the calculations to very small point clouds due to computational limits.
-
-As feature dimension increased, Rips complexes show a significant rise in run time with increasing feature dimension within an 8 dimensional point cloud.
-With this small sample set, it is not clear whether the growth is exponential or polynomial, so no regression values are provided in the supplement.
+However, it might also be due to computational restrictions of calculating high dimensional features.-->
+Large data and feature dimensionality often restricts persistent homology calculations to small point clouds due to computational limits.
+When calculating persistent homology on a high-dimensional point cloud, as Vietoris-Rips feature dimension increases, there is a corresponding increase in runtime (Figure \ref{fig:box}).
+Due to the limited number of points, it is unclear whether runtime grows polynomially or exponentially.
 
@@
   \includegraphics[height=2in]{../Figures/Final_Figures/fig3.png}
-  \caption{Comparison of Rips complex persistent homology calculation as a function of feature dimension.
+  \caption{
+    \textbf{Comparison of Rips complex persistent homology calculation as a function of feature dimension.}
     The x-axis represents the number of features in the point cloud, and y-axis represents average run time.
@@
     All Rips complexes were calculated on an 8 dimensional box.
-Panel 1.
-Comparing run times for TDAstats.
-Panel 2.
-Comparing run times for Dionysus.
-Panel 3.
-Comparing run times for GUDHI.
+    Panel 1. Comparing run times for TDAstats.
+    Panel 2. Comparing run times for Dionysus.
+    Panel 3. Comparing run times for GUDHI.
     Limited point clouds were tested for high dimensional features due to computational limits.

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised caption

@@
     \textbf{Comparison of Rips complex persistent homology calculation as a function of feature dimension.}
-    The x-axis represents the number of features in the point cloud, and y-axis represents average run time.
-    Point cloud number was held constant and indicated by color.
-    All Rips complexes were calculated on an 8 dimensional box.
-    Panel 1. Comparing run times for TDAstats.
-    Panel 2. Comparing run times for Dionysus.
-    Panel 3. Comparing run times for GUDHI.
-    Limited point clouds were tested for high dimensional features due to computational limits.
-    Average runtime is presented as mean +/- standard deviation over 10 iterations.}
+    Average runtime (mean $\pm$ SD, $n=10$ iterations per data point) for various point cloud sizes (denoted by color) is plotted against calculated feature dimension and faceted by TDA libary.
+    Persistent homology was calculated on a uniformly distributed random sample of points contained within a 1 unit, 8-dimensional cube.
+    Computational limitations of calculating persistent homology for a large number of feature dimensions restricted point clouds to relatively small sizes.}
 \label{fig:box}

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised narrative text

@@
 When calculating persistent homology on a high-dimensional point cloud, as Vietoris-Rips feature dimension increases, there is a corresponding increase in runtime (Figure \ref{fig:box}).
-Due to the limited number of points, it is unclear whether runtime grows polynomially or exponentially.
+Dionysus is clearly outmatched by GUDHI and Ripser as feature dimension increases, with the difference being clearest for larger point clouds; by feature dimension 5, Ripser outpaces GUDHI as well (Figure \ref{fig:box}).
+Due to the limited number of points, it is unclear whether runtime for each library grows polynomially or exponentially.
+<!--RECHECK, should be enough data point to regress on-->
 

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised text

@@
 \subsection{Comparison of alpha and Vietoris-Rips complexes within GUDHI}
+<!--SHOULD BE IN METHODS
 The GUDHI library provided functionality to calculate persistent homology with Rips and Alpha complexes. 
-Therefore, we only used the GUDHI library to compare performance between Rips and Alpha complexes.
+Therefore, we only used the GUDHI library to compare performance between Rips and Alpha complexes.-->
 
@@
 \section{Acknowledgements}
-Emily Dolson, PhD for assistance in using the HPC for collecting data
+Emily Dolson, PhD for assistance with using a high-performance computing node for data collection.
 

### 2020-03-08 tdabench-paper/tdabench.Rmd
updated table content

@@
 \hline
-Feature & TDA Stats & Dionysus & Gudhi Rips & Gudhi Alpha \\
+Feature & Ripser & Dionysus & Gudhi Rips & Gudhi Alpha \\
 Dimension & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit\\
@@
 Limiting factors were point cloud number and feature dimension calculation. 
-For Rips complexes, TDA stats had the highest performing capacity among the three Rips libraries. 
+For Vietoris-Rips complices, Ripser had the best performing capacity among the three libraries. 
 Alpha complexes were not computable for any point cloud greater than 3 dimensions, so dimensional features 3 and higher could not be computed.}

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised text

@@
 
-\subsection{Comparison of alpha and Vietoris-Rips complexes within GUDHI}
 <!--SHOULD BE IN METHODS
@@
 
-\subsubsection{Runtime as a function of data dimension (const feature)}
-In our study, our computers could not calculate persistent homology for Alpha complexes when the point cloud dimension exceeded 3. 
+Even with a constant feature dimension, the underlying data dimension could play a role in runtime of persistent homology calculation.
+Figure \ref{fig:ann} compares handling of this issue by the Vietoris-Rips complex and the alpha complex.
+Since GUDHI is the only library implementing functionality with an alpha complex, we compare its implementations of the Vietoris-Rips and alpha complices.
+Due to computational limitations, an alpha complex could not be calculated for any point clouds with data dimension exceeding 3.
+Two notable aspects of Figure \ref{fig:ann} stand out.
+First, the alpha complex calculation clearly runs faster than the Vietoris-Rips complex calculation, a trend that becomes clearer as point cloud size increases.
+Second, although the Vietoris-Rips complex calculation runtime appears to be independent of underlying data dimension, alpha complex calculation is dependent on it.
+Figure \ref{fig:ann} shows a subtle difference between data dimensions 2 and 3 as point cloud size increases.
+Although unconcerning for a data dimension up to 3, failure to run any alpha complex calculations with a data dimension of 4 could be cause for concern.
+<!--SHOULD BE IN DISCUSSION
 This is understandable as Alpha complex theoretical complexity increases exponentially for increasing point cloud dimension.
-
-Changes in computational performance from changing point cloud dimension is one major difference between calculation of Rips and Alpha complexes.
-This difference can be seen when comparing run times as a function of changing point cloud dimension.
-In figure \ref{fig:ann}, we show that Rips complexes  run time does not change for different dimensional point clouds as long as feature dimension is held constant.
-Even though Alpha complexes perform better overall, their computation time increases with increasing point cloud dimension even when calculating the same feature dimension.
-
-Another key difference is that Alpha complex runtime seems to have an apparent linear rise in the conditions for our analysis. 
-We provide the linear regression coefficients in the supplemental data on the paper Github.
+Changes in computational performance from changing point cloud dimension is one major difference between calculation of Rips and Alpha complexes.-->
 
@@
   \includegraphics[height=2in]{../Figures/Final_Figures/fig4.png}
-  \caption{Comparing homology calculation as a function of point cloud dimension for Rips vs Alpha complexes.
+  \caption{
+    \textbf{Comparing persistent homology calculation between Vietoris-Rips and alpha complices.}
     The x-axis represents the number of points in the point cloud, and y-axis represents average run time.
@@
 
+<!--Include somewhere that differences are b/n implementations (practical) not necessarily mathematical structures (theory), so results can change as implementations change (whether b/n libraries or b/n complices)-->
+
 \begin{figure}

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised text

@@
     \textbf{Comparing persistent homology calculation between Vietoris-Rips and alpha complices.}
-    The x-axis represents the number of points in the point cloud, and y-axis represents average run time.
-    Different dimensional annulus point clouds are indicaed by color.
-    Panel 1 = Comparing run times using GUDHI's Rips complex library.
-    Panel 2 = Comparing run times using GUDHI's Alpha complex library.
-    In all calculations, only 1 dimensional features were calculated.
-    For our parameters, Alpha complex run times are more linear in contrast to Rips more polynomial run times.
-    Exact regression data is provided in supplemental material on the paper Github.
-    Average runtime is presented as mean +/- standard deviation over 10 iterations.}
+    Average runtime (mean $\pm$ SD, $n=10$ iterations per data point) for various data dimensions (denoted by color) are plotted against point cloud size and faceted by type of simplicial complex.
+    Maximum of feature dimension was kept constant at 1.
+    Alpha complex runtimes are linear, in contrast to polynomial Vietoris-Rips runtimes (see Supplement for regression details).}
 \label{fig:ann}
@@
 
-Rips complexes show a rapid growth in object size as the point cloud number increases in figure \label{fig:mem1}. 
+Rips complexes show a rapid growth in object size as the point cloud number increases in figure \ref{fig:mem1}. 
 The growth pattern differences mirror our findings for run time growth with increasing point cloud size. 
@@
 
-Figure \label{fig:mem2}, panel 2 shows that Alpha complex object size depends on the feature dimension, point cloud size, and point cloud shape with torus and uniform shapes having the higher object size.
+Figure \ref{fig:mem2}, panel 2 shows that Alpha complex object size depends on the feature dimension, point cloud size, and point cloud shape with torus and uniform shapes having the higher object size.
 Noisier shapes like the box and annulus increase object size compared to more uniform shapes like the sphere.
-Figure \label{fig:mem2}, panel 1 shows that Rips complex object size depends only on the point cloud size and feature dimensions.
+Figure \ref{fig:mem2}, panel 1 shows that Rips complex object size depends only on the point cloud size and feature dimensions.
 

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised mem1 text

@@
 
-\subsubsection{Memory Benchmarks}
-As point cloud number increases, Rips complexes show a more dramatic rise in run time than Alpha complexes. 
-This is predicted by Rips theoretical exponential complexity and Alpha complex's theoretical polynomial complexity. 
-In both complexes, the boundary matrix is used to calculate persistent homology (explained in detail in the Mathematics section). 
-The library GUDHI provides functions to return the boundary matrix, which allows us to measure object size.
+Vietoris-Rips and alpha complices differ not only in runtime, but also in memory usage.
+GUDHI conveniently allows measurement of the simplicial complex boundary matrix, a reasonable proxy for total memory usage in calculation of persistent homology.
+Since Ripser and Dionysus do not have similar capabilities, they are excluded from memory benchmarks.
+Vietoris-Rips complices show a rapid growth in object size as point cloud size increases; in contrast, alpha complices consistently maintain low, linearly-growing memory use (Figure \ref{fig:mem1}).
+The growth pattern differences mirror findings for runtime growth (Figure \ref{fig:ann}).
+Interestingly, different factors determine memory usage in the two types of simplicial complex.
 
-Rips complexes show a rapid growth in object size as the point cloud number increases in figure \ref{fig:mem1}. 
-The growth pattern differences mirror our findings for run time growth with increasing point cloud size. 
-Alpha complexes show a linear growth in memory complexity whereas Rips complexes show a polynomial growth. 
-However, calculating the object size reveals some interesting differences in what parameters determine the object size of Rips vs Alpha complexes.
+<!--BELONGS IN DISCUSSION
+This is predicted by Vietoris-Rips theoretical exponential complexity and Alpha complex's theoretical polynomial complexity.-->
 
@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/fig5a.png}
-  \caption{Persistent Homology Boundary Matrix for Rips vs Alpha complex.
-The x-axis represents the number of points in the point cloud, and y-axis represents object size.
-The type of complex is indicated by color.
-Maximum feature dimensions were calculated.
-Panel 1 = Comparing object size on 3D annulus.
-Panel 2 = Comparing object size on sphere.
-Panel 3 = Comparing object size on torus.
-Panel 4 = Comparing object size on 3D box.
-Similar to runtime growth, Alpha complex showed a linear pattern and Rips complex showed a log-log linear pattern.
-Regression results can be found in supplemental data on the paper Github}
+  \caption{
+    \textbf{Comparing memory usage of persistent homology calculation using a Vietoris-Rips versus an alpha complex.}
+    Mean boundary matrix object size, a proxy for memory usage, plotted against point cloud size varying by point cloud structure (facet) and simplicial complex type (color, shape).
+    Data dimension and feature dimension were held constant at 3 and 2, respectively.
+    Point cloud structures for which memory usage was calculated were 3-dimensional annulus (top-left), 3-dimensional sphere (top-right), torus (bottom-left), and 3-dimensional cube (bottom-right).
+    Similar to runtime results, alpha complex calculation showed linear growth and Vietoris-Rips complex calculation showed polynomial growth (see Supplement for regression details).}
 \label{fig:mem1}
@@
 
-<!--Include somewhere that differences are b/n implementations (practical) not necessarily mathematical structures (theory), so results can change as implementations change (whether b/n libraries or b/n complices)-->
+<!--Include somewhere that differences are b/n implementations (practical) not necessarily mathematical structures (theory), so results can change as implementations change (whether b/n libraries or b/n complices)
+
+Ripser memory not measured here b/c R, but refer to Bauer Ripser memory measurements.
+-->
 

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised mem text

@@
 Since Ripser and Dionysus do not have similar capabilities, they are excluded from memory benchmarks.
-Vietoris-Rips complices show a rapid growth in object size as point cloud size increases; in contrast, alpha complices consistently maintain low, linearly-growing memory use (Figure \ref{fig:mem1}).
+Vietoris-Rips complices show rapid growth in object size as point cloud size increases; in contrast, alpha complices consistently maintain low, linearly-growing memory use (Figure \ref{fig:mem1}).
 The growth pattern differences mirror findings for runtime growth (Figure \ref{fig:ann}).
 Interestingly, different factors determine memory usage in the two types of simplicial complex.
+Whereas memory use in calculation of a Vietoris-Rips complex is dependent on point cloud size and feature dimension, memory use for an analogous calculation with an alpha complex is also dependent on the structure of the point cloud itself (Figure \ref{fig:mem2}).
+Within the limited set of tested point cloud structures, the sphere requires the least memory.
 
@@
 This is predicted by Vietoris-Rips theoretical exponential complexity and Alpha complex's theoretical polynomial complexity.-->
-
-Figure \ref{fig:mem2}, panel 2 shows that Alpha complex object size depends on the feature dimension, point cloud size, and point cloud shape with torus and uniform shapes having the higher object size.
-Noisier shapes like the box and annulus increase object size compared to more uniform shapes like the sphere.
-Figure \ref{fig:mem2}, panel 1 shows that Rips complex object size depends only on the point cloud size and feature dimensions.
 
@@
 
-<!--Include somewhere that differences are b/n implementations (practical) not necessarily mathematical structures (theory), so results can change as implementations change (whether b/n libraries or b/n complices)
+<!--Include somewhere:
+
+differences are b/n implementations (practical) not necessarily mathematical structures (theory), so results can change as implementations change (whether b/n libraries or b/n complices)
 
 Ripser memory not measured here b/c R, but refer to Bauer Ripser memory measurements.
+
+Rips complex memory more predictable, but alpha complex memory generally better (but what about Ripser?).
+
+Thoughts on why sphere alpha uses less memory than 3-d annulus alpha
 -->

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised limitations text

@@
 Within the limited set of tested point cloud structures, the sphere requires the least memory.
+Table \ref{tbl:limit} lists the maximum point cloud size for which persistent homology calculations were completed for each library.
 
@@
     \textbf{Comparing memory usage of persistent homology calculation using a Vietoris-Rips versus an alpha complex.}
-    Mean boundary matrix object size, a proxy for memory usage, plotted against point cloud size varying by point cloud structure (facet) and simplicial complex type (color, shape).
+    Mean boundary matrix size, a proxy for memory usage, plotted against point cloud size varying by point cloud structure (facet) and simplicial complex type (color, shape).
     Data dimension and feature dimension were held constant at 3 and 2, respectively.
@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/fig5t.png}
-  \caption{Different factors change boundary matrix size for each complex.
-The x-axis represents the number of points in the point cloud, and y-axis represents object size.
-Panel 1 = Rips complex object size depends on feature size and point cloud number alone.
-For Rips complex, the decreasing number of data points for each calculated feature dimension reflects the computational limits of large point cloud sizes.
-Point cloud shape is not defined as it had no impact on boundary matrix size.
-Panel 2 = Alpha complex object size depends on feature size, point cloud number, and point cloud shape.
-Point cloud shape was indicated as it impacted object size.}
+  \caption{
+    \textbf{Factors affecting boundary matrix size in Vietoris-Rips and alpha complices.}
+    Mean boundary matrix size plotted against point cloud size varying by point cloud structure (right, faceted), simplicial complex type (left, right), and feature dimension (color).
+    Whereas the boundary matrix for a Vietoris-Rips complex (left) is independent of point cloud structure, the same is not true for an alpha complex (right, faceted).
+    For the Vietoris-Rips complex, the decreasing number of data points plotted for successive feature dimensions reflects the computational requirements of increasing point cloud size.}
 \label{fig:mem2}
 \end{widefigure}
-
-\subsubsection{Computational Limitations (advice/guidance for users)}
-In benchmarking all of our parameters, each point cloud was given up to 16 GB of RAM. 
-However, memory allocation errors as reported by R caused several of our parameters to not run. 
-This table describes the point cloud size limit for calculating a given feature dimension. 
-Notably, TDAstats has a higher threshold for point cloud size at high dimensions when compared to TDA's GUDHI or Dionysus libraries. 
-Alpha complex's limitation is better described as the point cloud dimension being the limiting factor, which is why 3 and 4 dimensional features were not computable.
 
@@
 \end{tabular}
+\label{tbl:limit}
 \caption{During the benchmarking, certain libraries were not able to calculate persistent homology for sufficiently complex point clouds. 
@@
 Thoughts on why sphere alpha uses less memory than 3-d annulus alpha
+
+Weigh limitations and benefits of each library/complex type based on Table 1
 -->
@@
 
-\section{Acknowledgements}
+\section{Acknowledgments}
 Emily Dolson, PhD for assistance with using a high-performance computing node for data collection.

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised text
fixed gudhi capitalization

@@
 Within the limited set of tested point cloud structures, the sphere requires the least memory.
-Table \ref{tbl:limit} lists the maximum point cloud size for which persistent homology calculations were completed for each library.
+Table \ref{tbl:limit} lists the maximum point cloud size for which persistent homology calculations could be completed for each library.
 
@@
 \hline
-Feature & Ripser & Dionysus & Gudhi Rips & Gudhi Alpha \\
+Feature & Ripser & Dionysus & GUDHI Rips & GUDHI Alpha \\
 Dimension & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit\\
@@
 \end{tabular}
-\label{tbl:limit}
 \caption{During the benchmarking, certain libraries were not able to calculate persistent homology for sufficiently complex point clouds. 
@@
 Alpha complexes were not computable for any point cloud greater than 3 dimensions, so dimensional features 3 and higher could not be computed.}
+\label{tbl:limit}
 \end{table}

### 2020-03-08 tdabench-paper/tdabench.Rmd
added narrative for fig fin to results

@@
 
+In real life scenarios, practitioners likely care most about comparative performance of the fastest Vietoris-Rips complex library and the fastest alpha complex library.
+Figure \ref{fig:fin} illustrates this comparison, showing that for 3-dimensional point clouds, GUDHI's alpha complex library outpaces Ripser's Vietoris-Rips functionality.
+
 \section{Discussion}
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig6.png}
+  \caption{
+    \textbf{Runtime comparison of persistent homology calculation between Ripser's Vietoris-Rips and GUDHI's alpha complex functionality.}
+    Average runtime (mean $\pm$ SD, $n=10$ iterations per data point) for various 3-dimensional point cloud structures (facet) plotted against point cloud size for each library (color).
+    Benchmarking was conducted on an annulus (top-left), a sphere (top-right), a torus (bottom-left), and a cube (bottom-right).
+    Data was not collected for data dimensions greater than 3 due to computational limitations of calculating alpha complices.}
+\label{fig:fin}
+\end{figure}
+
 Based on this theoretical complexity and our results, Alpha complexes are superior for small dimensional point clouds (3 dimensions and below). 
@@
 GUDHI and Dionysus libraries, however, do have the option of returning the boundary matrix, which Ripser does not do as one of its time saving features.
-
-\begin{figure}
-  \centering
-  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig6.png}
-  \caption{TDAstats Rips complexes vs GUDHI Alpha complex average run times across all 3D shapes for 2D features.
-The x-axis represents the number of points in the point cloud, and y-axis represents average run time.
-The type of complex is indicated by color.
-Panel 1 = Execution time on annulus.
-Panel 2 = Execution time on sphere.
-Panel 3 = Execution time on torus.
-Panel 4 = Execution time on uniform box.
-Data is presented as mean +/- standard deviation}
-\label{fig:fin}
-\end{figure}
 
@@
 
+<!--PUT AS TEXT INSTEAD
 \begin{figure}
@@
 \label{decision}
-\end{figure}
+\end{figure}-->
 

### 2020-03-10 Figures/Final_Figures_Reproducible.R
widened errors bars figure 4

@@ -179,7 +179,7 @@ fig.4 <- ggplot(data.fig.4, aes(x=num.points, y=avg.time, color=point.cloud.dim,
   geom_point() + facet_wrap(~library, scales = "free") + 
   geom_errorbar(data.fig.4, mapping = aes(x=num.points, 
                                           ymin=avg.time - std, 
-                                          ymax=avg.time + std)) +
+                                          ymax=avg.time + std), width=28) +
   labs(color = "Annulus \nDimension",
        shape = "Annulus \nDimension",
        x = "Number of Points",

### 2020-03-19 tdabench-paper/tdabench.Rmd
alpha capitalization

@@
 The Dionysus (TDA), GUDHI (TDA), and Ripser C++ libraries (TDAstats) provide Rips complex functionality.
-GUDHI  (TDA package) also provides Alpha complex functionality.
+GUDHI  (TDA package) also provides alpha complex functionality.
 These software represent powerful analysis tools but are limited computationally.
@@
 We benchmarked the all libraries on a High-Powered Cluster for runtime.
-We also compared Alpha and Rips for memory burden.
+We also compared alpha and Rips for memory burden.
 Alpha complexes run significantly faster than all Rips complex libraries for all data sets for data dimension less than or equal to three.
-Alpha complexes were not computable for data dimension greater than 3.
+alpha complexes were not computable for data dimension greater than 3.
 TDAstats had the fastest runtimes for all Rips complexes.
-We recommend performing persistent homology with GUDHI Alpha complexes for data dimension less than or equal to three.
+We recommend performing persistent homology with GUDHI alpha complexes for data dimension less than or equal to three.
 For higher dimension, we recommend TDAstats Rips Complex.
@@
 There are several methods to generate these complexes on point clouds. 
-In this paper, we are using the Rips and Alpha complex for our persistent homology calculations. 
+In this paper, we are using the Rips and alpha complex for our persistent homology calculations. 
 In both methods, a series of data structures, called simplicial complexes, are built on top of a point cloud.
@@
 
-There are several C++ libraries that calculate Alpha and Rips complexes.
-Dionysus provides Alpha and Rips complex functionality \cite{Dionysus}.
+There are several C++ libraries that calculate alpha and Rips complexes.
+Dionysus provides alpha and Rips complex functionality \cite{Dionysus}.
 GUDHI is another C++ library that also provides both functionalities as well as other TDA methodologies \cite{gudhi}.
@@
 We specifically aim to characterize the computational performance (run time) and limits of persistent homology calculations between Rips complexes in the R packages TDAstats and TDA.
-We also explore the differences in computational performance between Rips and Alpha complexes in R to assess which datasets are most appropriate for each algorithm.
+We also explore the differences in computational performance between Rips and alpha complexes in R to assess which datasets are most appropriate for each algorithm.
 
@@
 
-There are several different methods to construct a simplicial complex on a given point cloud S, but this paper focuses on the Rips and Alpha complexes. 
+There are several different methods to construct a simplicial complex on a given point cloud S, but this paper focuses on the Rips and alpha complexes. 
 The Rips complex is perhaps the most common method for constructing a simplicial complex under persistent homology \cite{Rips-Complex}.
@@
 The size of the boundary matrix can describe the memory complexity on the RAM for persistent homology calculations. 
-We compare this memory complexity between Alpha and Rips complexes for this project.
+We compare this memory complexity between alpha and Rips complexes for this project.
 

### 2020-03-19 tdabench-paper/tdabench.Rmd
updated package mentions w/ tags

--- 2020-03-08 tdabench-paper/tdabench.Rmd
@@
 \section{Introduction}
+
 Topological data analysis (TDA) is a broad set of methodologies that characterize structural features of data sets inspired by topological principles.
@@
 In this project, we explore different methodologies to find efficient options that could solve this issue. 
-We specifically aim to characterize the computational performance (run time) and limits of persistent homology calculations between Rips complexes in the R packages TDAstats and TDA.
+We specifically aim to characterize the computational performance (run time) and limits of persistent homology calculations between Rips complexes in the \CRANpkg{TDAstats} and \CRANpkg{TDA} R packages.
 We also explore the differences in computational performance between Rips and Alpha complexes in R to assess which datasets are most appropriate for each algorithm.

### 2020-03-19 tdabench-paper/tdabench.Rmd
removed 1 discussion comment

@@
 
-<!--FOLLOWING SENTENCES SHOULD GO IN DISCUSSION
-For an n-dimensional point cloud, up to n-1 feature dimensions can be calculated. 
-In the field, most work has only included calculations up to 1 or 2 feature dimensions.
-Perhaps this is due to difficulty of conceptualizing what high dimensional features represent. 
-However, it might also be due to computational restrictions of calculating high dimensional features.-->
 Large data and feature dimensionality often restricts persistent homology calculations to small point clouds due to computational limits.

### 2020-03-19 tdabench-paper/tdabench.Rmd
removed a comment that was already addressed
removed 1 comment after discussing

@@
 
-<!--SHOULD BE IN METHODS
-The GUDHI library provided functionality to calculate persistent homology with Rips and Alpha complexes. 
-Therefore, we only used the GUDHI library to compare performance between Rips and Alpha complexes.-->
-
 Even with a constant feature dimension, the underlying data dimension could play a role in runtime of persistent homology calculation.
@@
 Although unconcerning for a data dimension up to 3, failure to run any alpha complex calculations with a data dimension of 4 could be cause for concern.
-<!--SHOULD BE IN DISCUSSION
-This is understandable as Alpha complex theoretical complexity increases exponentially for increasing point cloud dimension.
-Changes in computational performance from changing point cloud dimension is one major difference between calculation of Rips and Alpha complexes.-->
 

### 2020-03-19 tdabench-paper/tdabench.Rmd
added @eashwarsoma youtube playlist to methods
modified phrasing

@@
 Fully reproducible code for all numerical results and figures can be found at https://github.com/eashwarsoma/TDA-benchmark.
+Video explanations of TDA concepts and reproducing all results in this report can be found at https://tinyurl.com/TDABench.
 
@@
 Dionysus is clearly outmatched by GUDHI and Ripser as feature dimension increases, with the difference being clearest for larger point clouds; by feature dimension 5, Ripser outpaces GUDHI as well (Figure \ref{fig:box}).
-Due to the limited number of points, it is unclear whether runtime for each library grows polynomially or exponentially.
-<!--RECHECK, should be enough data point to regress on-->
+It is unclear whether runtime for each library grows polynomially or exponentially (see Supplement).
 

### 2020-03-19 tdabench-paper/tdabench.Rmd
updated parenthetical for consistency
added supplement ref to methods

@@
 Fully reproducible code for all numerical results and figures can be found at https://github.com/eashwarsoma/TDA-benchmark.
+This GitHub repository also contains instructions for generating the Supplement referenced in this report's results.
 Video explanations of TDA concepts and reproducing all results in this report can be found at https://tinyurl.com/TDABench.
@@
 Dionysus is clearly outmatched by GUDHI and Ripser as feature dimension increases, with the difference being clearest for larger point clouds; by feature dimension 5, Ripser outpaces GUDHI as well (Figure \ref{fig:box}).
-It is unclear whether runtime for each library grows polynomially or exponentially (see Supplement).
+It is unclear whether runtime for each library grows polynomially or exponentially (see Supplement for regression details).
 

### 2020-03-19 tdabench-paper/tdabench.Rmd
capitlization

--- 2020-03-19 tdabench-paper/tdabench.Rmd
@@
 We benchmarked the all libraries on a High-Powered Cluster for runtime.
-We also compared Alpha and Rips for memory burden.
+We also compared alpha and Rips for memory burden.
 Alpha complexes run significantly faster than all Rips complex libraries for all data sets for data dimension less than or equal to three.
@@
 TDAstats had the fastest runtimes for all Rips complexes.
-We recommend performing persistent homology with GUDHI Alpha complexes for data dimension less than or equal to three.
+We recommend performing persistent homology with GUDHI alpha complexes for data dimension less than or equal to three.
 For higher dimension, we recommend TDAstats Rips Complex.
@@
 \section{Introduction}
-
 Topological data analysis (TDA) is a broad set of methodologies that characterize structural features of data sets inspired by topological principles.
@@
 In this project, we explore different methodologies to find efficient options that could solve this issue. 
-We specifically aim to characterize the computational performance (run time) and limits of persistent homology calculations between Rips complexes in the \CRANpkg{TDAstats} and \CRANpkg{TDA} R packages.
-We also explore the differences in computational performance between Rips and Alpha complexes in R to assess which datasets are most appropriate for each algorithm.
+We specifically aim to characterize the computational performance (run time) and limits of persistent homology calculations between Rips complexes in the R packages TDAstats and TDA.
+We also explore the differences in computational performance between Rips and alpha complexes in R to assess which datasets are most appropriate for each algorithm.
 
@@
 
+<!--FOLLOWING SENTENCES SHOULD GO IN DISCUSSION
+For an n-dimensional point cloud, up to n-1 feature dimensions can be calculated. 
+In the field, most work has only included calculations up to 1 or 2 feature dimensions.
+Perhaps this is due to difficulty of conceptualizing what high dimensional features represent. 
+However, it might also be due to computational restrictions of calculating high dimensional features.-->
 Large data and feature dimensionality often restricts persistent homology calculations to small point clouds due to computational limits.
@@
 
+<!--SHOULD BE IN METHODS
+The GUDHI library provided functionality to calculate persistent homology with Rips and Alpha complexes. 
+Therefore, we only used the GUDHI library to compare performance between Rips and Alpha complexes.-->
+
 Even with a constant feature dimension, the underlying data dimension could play a role in runtime of persistent homology calculation.
@@
 Although unconcerning for a data dimension up to 3, failure to run any alpha complex calculations with a data dimension of 4 could be cause for concern.
+<!--SHOULD BE IN DISCUSSION
+This is understandable as Alpha complex theoretical complexity increases exponentially for increasing point cloud dimension.
+Changes in computational performance from changing point cloud dimension is one major difference between calculation of Rips and Alpha complexes.-->
 

### 2020-03-19 tdabench-paper/tdabench.Rmd

--- 2020-03-19 tdabench-paper/tdabench.Rmd
@@
 We specifically aim to characterize the computational performance (run time) and limits of persistent homology calculations between Rips complexes in the \CRANpkg{TDAstats} and \CRANpkg{TDA} R packages.
-We also explore the differences in computational performance between Rips and Alpha complexes in R to assess which datasets are most appropriate for each algorithm.
+We also explore the differences in computational performance between Rips and alpha complexes in R to assess which datasets are most appropriate for each algorithm.
 

### 2020-03-19 tdabench-paper/tdabench.Rmd
Edited discussion

@@
   Several persistent homology software libraries have been implemented in R.
-The Dionysus (TDA), GUDHI (TDA), and Ripser C++ libraries (TDAstats) provide Rips complex functionality.
+The Dionysus (TDA), GUDHI (TDA), and Ripser C++ libraries (TDAstats) provide Vietoris-Rips complex functionality.
 GUDHI  (TDA package) also provides alpha complex functionality.
@@
 We benchmarked the all libraries on a High-Powered Cluster for runtime.
-We also compared alpha and Rips for memory burden.
-Alpha complexes run significantly faster than all Rips complex libraries for all data sets for data dimension less than or equal to three.
+We also compared alpha and Vietoris-Rips for memory burden.
+Alpha complexes run significantly faster than all Vietoris-Rips complex libraries for all data sets for data dimension less than or equal to three.
 Alpha complexes were not computable for data dimension greater than 3.
-TDAstats had the fastest runtimes for all Rips complexes.
+TDAstats had the fastest runtimes for all Vietoris-Rips complexes.
 We recommend performing persistent homology with GUDHI alpha complexes for data dimension less than or equal to three.
-For higher dimension, we recommend TDAstats Rips Complex.
+For higher dimension, we recommend TDAstats Vietoris-Rips Complex.
 output:
@@
 There are several methods to generate these complexes on point clouds. 
-In this paper, we are using the Rips and alpha complex for our persistent homology calculations. 
+In this paper, we are using the Vietoris-Rips and alpha complex for our persistent homology calculations. 
 In both methods, a series of data structures, called simplicial complexes, are built on top of a point cloud.
@@
 One limitation of persistent homology is its run time limits for processing large numbers of points ( $> 10^9$ points) and calculating of high dimensional features ($>$ 2 dimensions). 
-Alpha complex calculations have a complexity of $O(n^\frac{d}{2})$ and Rips complex calculations have a complexity of $O(2^n)$ where n is the number of points and d is the dimension of the point cloud \citep{roadmap}.
+Alpha complex calculations have a complexity of $O(n^\frac{d}{2})$ and Vietoris-Rips complex calculations have a complexity of $O(2^n)$ where n is the number of points and d is the dimension of the point cloud \citep{roadmap}.
 The limits and computational resources needed to run complex data sets with major persistent homology C++ software have been well characterized \cite{roadmap}.
 
-There are several C++ libraries that calculate alpha and Rips complexes.
-Dionysus provides alpha and Rips complex functionality \cite{Dionysus}.
+There are several C++ libraries that calculate alpha and Vietoris-Rips complexes.
+Dionysus provides alpha and Vietoris-Rips complex functionality \cite{Dionysus}.
 GUDHI is another C++ library that also provides both functionalities as well as other TDA methodologies \cite{gudhi}.
@@
 
-A new C++ library, Ripser, was developed as a more efficient way to calculate Rips complexes.
+A new C++ library, Ripser, was developed as a more efficient way to calculate Vietoris-Rips complexes.
 The code is supposed to outperform other similar libraries (e.g.
@@
 It calculates persistent homology using the more efficient Ripser algorithm and provides data visualization with integration of ggplot.
-While Ripser is known to be faster than the other libraries for Rips complex calculations in C++, how this translates in the R interface is not clear.
+While Ripser is known to be faster than the other libraries for Vietoris-Rips complex calculations in C++, how this translates in the R interface is not clear.
 While C++ generally runs faster than R, R provides higher level functionality, and appeals to a broad audience of data science users.
@@
 In this project, we explore different methodologies to find efficient options that could solve this issue. 
-We specifically aim to characterize the computational performance (run time) and limits of persistent homology calculations between Rips complexes in the \CRANpkg{TDAstats} and \CRANpkg{TDA} R packages.
-We also explore the differences in computational performance between Rips and alpha complexes in R to assess which datasets are most appropriate for each algorithm.
+We specifically aim to characterize the computational performance (run time) and limits of persistent homology calculations between Vietoris-Rips complexes in the \CRANpkg{TDAstats} and \CRANpkg{TDA} R packages.
+We also explore the differences in computational performance between Vietoris-Rips and alpha complexes in R to assess which datasets are most appropriate for each algorithm.
 
@@
 
-There are several different methods to construct a simplicial complex on a given point cloud S, but this paper focuses on the Rips and alpha complexes. 
-The Rips complex is perhaps the most common method for constructing a simplicial complex under persistent homology \cite{Rips-Complex}.
+There are several different methods to construct a simplicial complex on a given point cloud S, but this paper focuses on the Vietoris-Rips and alpha complexes. 
+The Vietoris-Rips complex is perhaps the most common method for constructing a simplicial complex under persistent homology \cite{Rips-Complex}.
 In a point cloud of k points, a distance parameter, $\delta$, can be used to draw a circle of diameter $\delta$ around every single point in S.
@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/Intro_Rips.png}
-  \caption{Basic Visualization of the Rips Complex.
+  \caption{Basic Visualization of the Vietoris-Rips Complex.
 For a given parameter, $\delta$, $\delta$ diameter circles are drawn around each point.
@@
 Alpha complexes are simplicial complexes that are subsets of the Delaunay Triangulation.
-The parameter, $\alpha$, can describe the radius of a ball of each point in the point cloud S much like $\delta$ describes the diameter of a circle in the Rips complex.
+The parameter, $\alpha$, can describe the radius of a ball of each point in the point cloud S much like $\delta$ describes the diameter of a circle in the Vietoris-Rips complex.
 An Alpha complex at radius $\alpha$ is formed through connecting centers of intersecting regions formed by intersecting the union of $\alpha$ radius balls and the Voronoi diagram. 
 Eventually, once $\alpha$ is large enough, the full Delaunay Triangulation is formed.
-In between these stages, the birth and death of features at certain values of $\alpha$ can be captured in a three column persistent homology matrix much like the Rips complex.
+In between these stages, the birth and death of features at certain values of $\alpha$ can be captured in a three column persistent homology matrix much like the Vietoris-Rips complex.
 
@@
 The size of the boundary matrix can describe the memory complexity on the RAM for persistent homology calculations. 
-We compare this memory complexity between alpha and Rips complexes for this project.
+We compare this memory complexity between alpha and Vietoris-Rips complexes for this project.
 
@@
   \caption{
-    \textbf{Comparison of Rips complex persistent homology calculation as a function of feature dimension.}
+    \textbf{Comparison of Vietoris-Rips complex persistent homology calculation as a function of feature dimension.}
     Average runtime (mean $\pm$ SD, $n=10$ iterations per data point) for various point cloud sizes (denoted by color) is plotted against calculated feature dimension and faceted by TDA libary.
@@
 
-Figure \ref{fig:fin} shows that at high point cloud sizes, GUDHI’s Alpha complex calculations outperforms TDAstats. 
-Theoritically, Alpha complexes gain polynomial complexity as the number of points increases whereas Rips complexes gain exponentially. 
-Specifically, Alpha complexes are $O(n^\frac{d}{2})$ and Rips complexes are $O(2^n)$ where n is the number of points on a point cloud and d is the dimensions on the point cloud. 
-For the conditions in our paper, Rips and Alpha complexes both performed better than their theoritical maximums. 
-Rips complex calculations consistently had a polynomial growth for both runtime and boundary matrix size while Alpha complexes had linear growth.
+Figure \ref{fig:fin} shows that at high point cloud sizes, GUDHI’s alpha complex calculations outperforms TDAstats. 
+Theoretically, alpha complexes gain polynomial complexity as the number of points increases whereas Vietoris-Rips complexes gain exponentially. 
+Specifically, alpha complexes are $O(n^\frac{d}{2})$ and Vietoris-Rips complexes are $O(2^n)$ where n is the number of points on a point cloud and d is the dimensions on the point cloud. 
+For the conditions in our paper, Vietoris-Rips and alpha complexes both performed better than their theoritical maximums. 
+Vietoris-Rips complex calculations consistently had a polynomial growth for both runtime and boundary matrix size while alpha complexes had linear growth.
 
@@
 
-Based on this theoretical complexity and our results, Alpha complexes are superior for small dimensional point clouds (3 dimensions and below). 
+Based on this theoretical complexity and our results, alpha complexes are superior for small dimensional point clouds (3 dimensions and below). 
 This advantage becomes especially clear at high number of points.
-However, with out computational resources, Alpha complexes were not usable for point cloud dimensions greater than 3. 
-If data is more than three dimensions, then it should be dimensionally reduced by a procedure such as PCA for before using Alpha complexes.
+However, without sufficient computational resources, Alpha complexes were not usable for point cloud dimensions greater than 3. 
+If data is more than three dimensions, then it should be dimensionally reduced by a procedure such as PCA for before using alpha complexes.
 
-However, if data dimension cannot be reduced, then Rips complexes should be used. Among the different libraries, TDAstats has the fastest calculation time. 
+However, if data dimension cannot be reduced, then Vietoris-Rips complexes should be used. Among the different libraries, TDAstats has the fastest calculation time. 
 GUDHI and Dionysus significantly fall behind for high feature dimension calculations and high numbers of points. 
@@
 
-While Rips complexes can handle high dimensional data well, it still significantly slows down when looking for higher dimensional topological features. 
+While Vietoris-Rips complexes can handle high dimensional data well, it still significantly slows down when looking for higher dimensional topological features. 
 The size of the boundary matrix rapidly increases and saturates the RAM even for small numbers of points (e.g. ~100 points when looking for 4 dimensional features). 
 Looking for high dimensional features in a large high dimensional point cloud remains a challenge. 
-Other persistent homology complex methods exist such as the Delaunay complex, Witness complex, etc. but they are not currently implemented in R [cite]. 
+Other persistent homology complex methods exist such as the Delaunay complex, Witness complex, etc. but they are not currently implemented in R. 
 Future challenges would be creating and implementing algorithms that reduce the computational complexity of higher dimensional topological feature calculations for R.
@@
 
-Rips complex memory more predictable, but alpha complex memory generally better (but what about Ripser?).
+Vietoris-Rips complex memory more predictable, but alpha complex memory generally better (but what about Ripser?).
 

### 2020-03-19 tdabench-paper/tdabench.Rmd
updated complex labels
updated phrasing

@@
 As persistent homology calculations continue to become a more popular tool to analyze complex multidimensional data, it will be important to understand from a computational perspective which method to use.
-In this paper, we examined two forms of persistent homology complexes: Vietoris Rips and Alpha complexes. 
-Both algorithms describe the topological features through the generation of simplicial complexes. 
+In this paper, we examined two forms of persistent homology complexes: Vietoris-Rips and alpha complexes. 
+Both algorithms describe topological features through the generation of simplicial complexes. 
 The advantage in saving computational time in choosing a particular algorithm depends on the point cloud characteristics.

### 2020-03-19 tdabench-paper/tdabench.Rmd
corrected phrasing
correct library name

@@
 Both algorithms describe topological features through the generation of simplicial complexes. 
-The advantage in saving computational time in choosing a particular algorithm depends on the point cloud characteristics.
+The advantage in saving computational time by choosing a particular algorithm depends on point cloud characteristics.
 
-Figure \ref{fig:fin} shows that at high point cloud sizes, GUDHI’s alpha complex calculations outperforms TDAstats. 
+Figure \ref{fig:fin} shows that at high point cloud sizes, GUDHI’s alpha complex outperforms Ripser. 
 Theoretically, alpha complexes gain polynomial complexity as the number of points increases whereas Vietoris-Rips complexes gain exponentially. 

### 2020-03-19 tdabench-paper/tdabench.Rmd
updated exponent aesthetics

@@
 Figure \ref{fig:fin} shows that at high point cloud sizes, GUDHI’s alpha complex outperforms Ripser. 
-Theoretically, alpha complexes gain polynomial complexity as the number of points increases whereas Vietoris-Rips complexes gain exponentially. 
-Specifically, alpha complexes are $O(n^\frac{d}{2})$ and Vietoris-Rips complexes are $O(2^n)$ where n is the number of points on a point cloud and d is the dimensions on the point cloud. 
+Theoretically, alpha complexes gain polynomial complexity as the number of points increases whereas Vietoris-Rips complexes gain exponentially \citep{roadmap}.
+Specifically, alpha complexes are $O(n^{d/2})$ and Vietoris-Rips complexes are $O(2^n)$ where n is the number of points on a point cloud and d is the dimensions on the point cloud. 
 For the conditions in our paper, Vietoris-Rips and alpha complexes both performed better than their theoritical maximums. 

### 2020-03-19 tdabench-paper/tdabench.Rmd
math mode for vars

@@
 Theoretically, alpha complexes gain polynomial complexity as the number of points increases whereas Vietoris-Rips complexes gain exponentially \citep{roadmap}.
-Specifically, alpha complexes are $O(n^{d/2})$ and Vietoris-Rips complexes are $O(2^n)$ where n is the number of points on a point cloud and d is the dimensions on the point cloud. 
+Specifically, alpha complexes are $O(n^{d/2})$ and Vietoris-Rips complexes are $O(2^n)$ where $n$ is the number of points on a point cloud and $d$ is the dimensions on the point cloud. 
 For the conditions in our paper, Vietoris-Rips and alpha complexes both performed better than their theoritical maximums. 

### 2020-03-19 tdabench-paper/tdabench.Rmd
fixed misspelling
fixed phrasing

@@
 Specifically, alpha complexes are $O(n^{d/2})$ and Vietoris-Rips complexes are $O(2^n)$ where $n$ is the number of points on a point cloud and $d$ is the dimensions on the point cloud. 
-For the conditions in our paper, Vietoris-Rips and alpha complexes both performed better than their theoritical maximums. 
+For the conditions in our paper, Vietoris-Rips and alpha complexes both performed better than their theoretical maximums.
 Vietoris-Rips complex calculations consistently had a polynomial growth for both runtime and boundary matrix size while alpha complexes had linear growth.
@@
 
-Based on this theoretical complexity and our results, alpha complexes are superior for small dimensional point clouds (3 dimensions and below). 
+Based on the theoretical complexity and our results, alpha complexes are superior for point clouds with 3 or fewer dimensions. 
 This advantage becomes especially clear at high number of points.

### 2020-03-19 tdabench-paper/tdabench.Rmd
fixed phrasing

@@
 This advantage becomes especially clear at high number of points.
-However, without sufficient computational resources, Alpha complexes were not usable for point cloud dimensions greater than 3. 
-If data is more than three dimensions, then it should be dimensionally reduced by a procedure such as PCA for before using alpha complexes.
-
-However, if data dimension cannot be reduced, then Vietoris-Rips complexes should be used. Among the different libraries, TDAstats has the fastest calculation time. 
-GUDHI and Dionysus significantly fall behind for high feature dimension calculations and high numbers of points. 
+However, without sufficient computational resources, alpha complexes were not usable for point cloud dimensions greater than 3.
+If a point cloud has more than 3 dimensions, then it could undergo pre-processing with dimension reduction before using alpha complexes.
+However, if data dimension cannot be reduced without significant information loss, then Vietoris-Rips complexes should be used.
+Among the tested libraries, Ripser (wrapped by \CRANpkg{TDAstats}) has the fastest calculation time. 
+GUDHI and Dionysus (wrapped by \CRANpkg{TDA}) significantly fall behind as feature dimension and number of points increase. 
 GUDHI and Dionysus libraries, however, do have the option of returning the boundary matrix, which Ripser does not do as one of its time saving features.

### 2020-03-19 tdabench-paper/tdabench.Rmd
fixed phrasing

@@
 GUDHI and Dionysus (wrapped by \CRANpkg{TDA}) significantly fall behind as feature dimension and number of points increase. 
-GUDHI and Dionysus libraries, however, do have the option of returning the boundary matrix, which Ripser does not do as one of its time saving features.
+GUDHI and Dionysus, however, do have the option of returning the boundary matrix, which Ripser does not store for the sake of optimization.
 
-While Vietoris-Rips complexes can handle high dimensional data well, it still significantly slows down when looking for higher dimensional topological features. 
-The size of the boundary matrix rapidly increases and saturates the RAM even for small numbers of points (e.g. ~100 points when looking for 4 dimensional features). 
-Looking for high dimensional features in a large high dimensional point cloud remains a challenge. 
+While Vietoris-Rips complexes can handle high-dimensional data well, calculation still significantly slows down when looking for higher dimension features. 
+Boundary matrix size rapidly increases and can overwhelm available random access memory even for small numbers of points (e.g. 100 points when looking for 4 dimensional features). 
+Looking for high-dimensional features in a large, high-dimensional point clouds remains a challenge. 
 Other persistent homology complex methods exist such as the Delaunay complex, Witness complex, etc. but they are not currently implemented in R. 

### 2020-03-19 tdabench-paper/tdabench.Rmd
fixed phrasing

@@
 While Vietoris-Rips complexes can handle high-dimensional data well, calculation still significantly slows down when looking for higher dimension features. 
-Boundary matrix size rapidly increases and can overwhelm available random access memory even for small numbers of points (e.g. 100 points when looking for 4 dimensional features). 
+Boundary matrix size rapidly increases and can overwhelm available random access memory even for small numbers of points (e.g. 100 points when looking for 4-dimensional features). 
 Looking for high-dimensional features in a large, high-dimensional point clouds remains a challenge. 
-Other persistent homology complex methods exist such as the Delaunay complex, Witness complex, etc. but they are not currently implemented in R. 
+Methods to calculate persistent homology do exist for other simplicial complexes, such as the Delaunay complex and the Witness complex, but, to our knowledge, they are not currently implemented in CRAN packages.
 Future challenges would be creating and implementing algorithms that reduce the computational complexity of higher dimensional topological feature calculations for R.

### 2020-03-19 tdabench-paper/tdabench.Rmd
fixed phrasing

@@
 
-Vietoris-Rips complex memory more predictable, but alpha complex memory generally better (but what about Ripser?).
-
 Thoughts on why sphere alpha uses less memory than 3-d annulus alpha
@@
 \section{Acknowledgments}
-Emily Dolson, PhD for assistance with using a high-performance computing node for data collection.
+The authors thank Emily Dolson, PhD for assistance with using a high-performance computing node for data collection.
 

### 2020-03-19 tdabench-paper/tdabench.Rmd
added Ripser's mem performance advantage
changes to discussion

@@
 This advantage becomes especially clear at high number of points.
+This difference in performance is clear in both runtime and memory use. 
+Interestingly, while alpha complexes had overall less memory use, the memory use varied depending on the shape. 
+Alpha complexes seem to require more memory for noisier data sets such as the annulus when compared to the sphere.
+
 However, without sufficient computational resources, alpha complexes were not usable for point cloud dimensions greater than 3.
@@
 However, if data dimension cannot be reduced without significant information loss, then Vietoris-Rips complexes should be used.
+It should be stated that if a point cloud is compatible with both complexes, both analyses should be performed as there may be a variation the persistent homology matrix. 
 Among the tested libraries, Ripser (wrapped by \CRANpkg{TDAstats}) has the fastest calculation time. 
@@
 GUDHI and Dionysus, however, do have the option of returning the boundary matrix, which Ripser does not store for the sake of optimization.
+While we did not measure Ripser's boundary matrix object size, Ripser's C++ code (outside its implementation in R) demonstrates memory advantages ~\citep{Ripser}
 

### 2020-03-19 tdabench-paper/tdabench.Rmd
added line about alpha's limits being computational not mathematical
modified discussion of alpha complex limitation

@@
 \section{Methods}
-<!--An exclusively R (v3.6.2) code pipeline was built for this project \citep{Rlang,ihaka1996}.-->
+
 We use \CRANpkg{readr} v1.3.1 to read rectangular data \citep{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2}, \CRANpkg{scatterplot3d} v0.3-41 \citep{scatterplot3d}, \CRANpkg{recexcavAAR} v0.3.0 \citep{recexcavAAR}, and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
@@
 
-<!--BELONGS IN DISCUSSION
-This is predicted by Vietoris-Rips theoretical exponential complexity and Alpha complex's theoretical polynomial complexity.-->
-
 \begin{figure}
@@
 If a point cloud has more than 3 dimensions, then it could undergo pre-processing with dimension reduction before using alpha complexes.
+Note, it is possible an algorithm will eventually be developed to enable alpha complex computation of higher dimensional data. 
 However, if data dimension cannot be reduced without significant information loss, then Vietoris-Rips complexes should be used.
@@
 
-<!--Include somewhere:
-
-differences are b/n implementations (practical) not necessarily mathematical structures (theory), so results can change as implementations change (whether b/n libraries or b/n complices)
-
-Ripser memory not measured here b/c R, but refer to Bauer Ripser memory measurements.
-
-Thoughts on why sphere alpha uses less memory than 3-d annulus alpha
-
-Weigh limitations and benefits of each library/complex type based on Table 1
--->
-
-<!--PUT AS TEXT INSTEAD
-\begin{figure}
-  \centering
-  \includegraphics[height=2.5in]{decision.png}
-    \caption{Basic decision chart for choosing TDA library based on dataset.}
-\label{decision}
-\end{figure}-->
-
 \section{Acknowledgments}

### 2020-03-19 Figures/Final_Figures_Reproducible.R
updated title figure 6

@@ -337,7 +337,7 @@ fig.6 <- ggplot(data.fig.6, aes(x=num.points, y=avg.time, color=library, shape=l
        shape = "Library",
        x = "Number of Points",
        y = "Average Run Time (s)",
-       title = "Rips (TDAstats) vs Alpha Complex (GUDHI) \nRun Times on 3D Point Clouds",
+       title = "Vietoris-Rips (Ripser) vs Alpha Complex (GUDHI) \nRun Times on 3D Point Clouds",
        subtitle = "") + 
   scale_x_continuous(limits=c(5,550), breaks = c(100, 200, 300, 400, 500)) + 
   scale_y_continuous(limits=c(-.05, 84))

### 2020-03-19 tdabench-paper/tdabench.Rmd
updated intro

@@
 
-Topological data analysis (TDA) is a broad set of methodologies that characterize structural features of data sets inspired by topological principles.
-Within the umbrella of TDA, persistent homology represents an algebraic approach to understanding the number, characteristics, and persistence of structural features in an n-dimensional point cloud.
+Topological data analysis (TDA) is a broad set of methodologies that characterizes structural features of datasets inspired by topological principles.
+It has a broad range of usage from viral evolution to physical chemistry \citep{TDA-Viral,TDA-PChem}. 
+Within the umbrella of TDA, persistent homology represents an algebraic approach to understanding the number, characteristics, and persistence of structural features in an $n$-dimensional point cloud.
+In the basic workflow of persistent homology, a series of simplicial complexes are generated on point clouds to characterize topological features.
+There are several methods to generate these complexes on point clouds. 
+In this paper, we focus on persistent homology of the Vietoris-Rips and alpha complexes, which use simplicial complexes to approximate topologic relationships in point clouds.
+The exact method of constructing these complexes is described in the Mathematics section.
+Persistent features more likely represent real geometric patterns, rather than noise.
 
-In the basic workflow of persistent homology, a series of simplicial complexes are generated on point clouds to characterize the topological features. 
-Simplicial complexes are connected sets of points, lines, and n-dimensional shapes.
-There are several methods to generate these complexes on point clouds. 
-In this paper, we are using the Vietoris-Rips and alpha complex for our persistent homology calculations. 
-In both methods, a series of data structures, called simplicial complexes, are built on top of a point cloud.
-These complexes can be viewed as approximating real topologic relationships in the data.
-The exact method of constructing these complexes is described in the Mathematics section. 
-However, the key takeaway is that the simplicial complex attempts to reveal the underlying topology of the point cloud. 
-These features can be described by dimension and "persistence." 
-Long lasting (persistent) features that exist across the series of simplicial complexes intuitively represent "real" geometrical patterns as opposed to noise.
-
-One limitation of persistent homology is its run time limits for processing large numbers of points ( $> 10^9$ points) and calculating of high dimensional features ($>$ 2 dimensions). 
-Alpha complex calculations have a complexity of $O(n^\frac{d}{2})$ and Vietoris-Rips complex calculations have a complexity of $O(2^n)$ where n is the number of points and d is the dimension of the point cloud \citep{roadmap}.
-The limits and computational resources needed to run complex data sets with major persistent homology C++ software have been well characterized \cite{roadmap}.
-
-There are several C++ libraries that calculate alpha and Vietoris-Rips complexes.
-Dionysus provides alpha and Vietoris-Rips complex functionality \cite{Dionysus}.
-GUDHI is another C++ library that also provides both functionalities as well as other TDA methodologies \cite{gudhi}.
-These libraries have been implemented in R with the TDA package \cite{TDA} providing an easier interface to utilize persistent homology calculations.
-
-A new C++ library, Ripser, was developed as a more efficient way to calculate Vietoris-Rips complexes.
-The code is supposed to outperform other similar libraries (e.g.
-Dionysus and GUDHI) by a factor over 40 in runtime \cite{Ripser}.
-This library was given an R interface in the package TDAstats \cite{TDAstats}. 
-TDAstats represents a more specialized package for topological data analysis compared to the TDA package.
-It calculates persistent homology using the more efficient Ripser algorithm and provides data visualization with integration of ggplot.
-While Ripser is known to be faster than the other libraries for Vietoris-Rips complex calculations in C++, how this translates in the R interface is not clear.
-While C++ generally runs faster than R, R provides higher level functionality, and appeals to a broad audience of data science users.
-
-Topological data analysis has had a broad range of usage from viral evolution to physical chemistry\cite{TDA-Viral,TDA-PChem}. 
-Given that persistent homology can apply to any n-dimensional data set, it has wide ranging applicability. 
-Its main limiting factor is computational for calculating high dimensional features in large multi-dimensional data sets. 
-We believe lack of analysis of these higher dimensional features may under utilize persistent homology's analysis potential.
-In this project, we explore different methodologies to find efficient options that could solve this issue. 
-We specifically aim to characterize the computational performance (run time) and limits of persistent homology calculations between Vietoris-Rips complexes in the \CRANpkg{TDAstats} and \CRANpkg{TDA} R packages.
-We also explore the differences in computational performance between Vietoris-Rips and alpha complexes in R to assess which datasets are most appropriate for each algorithm.
+There are several C++ libraries available to researchers that calculate alpha and Vietoris-Rips complexes, such as Dionysus, GUDHI, and Ripser \citep{Dionysus,gudhi,Ripser}.
+These libraries have been wrapped in R by the \CRANpkg{TDA} and \CRANpkg{TDAstats} packages \citep{TDA,TDAstats}.
+Although useful, calculating persistent homology for large datasets is often limited due to computational complexity \citep{roadmap}.
+As a result, researchers often limit persistent homology analysis to lower dimensions.
+However, ignoring features in higher dimensions may cause significant information loss, underutilizing persistent homology's capabilities.
+Here, we aim to benchmark two R packages - \CRANpkg{TDA} and \CRANpkg{TDAstats} - and enable researchers to most efficiently calculate persistent homology in R.
 
 \section{Mathematics of Persistent Homology}
+
 This section summarizes the mathematics behind generating simplicial complexes on point clouds in the workflow of persistent homology. 
@@
 We compare this memory complexity between alpha and Vietoris-Rips complexes for this project.
+Alpha complex calculations have a complexity of $O(n^\frac{d}{2})$ and Vietoris-Rips complex calculations have a complexity of $O(2^n)$ where n is the number of points and d is the dimension of the point cloud \citep{roadmap}.
 

### 2020-03-19 tdabench-paper/tdabench.Rmd
updated citations

@@
 
-We use \CRANpkg{readr} v1.3.1 to read rectangular data \citep{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2}, \CRANpkg{scatterplot3d} v0.3-41 \citep{scatterplot3d}, \CRANpkg{recexcavAAR} v0.3.0 \citep{recexcavAAR}, and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
+We use \CRANpkg{readr} v1.3.1 to read rectangular data \citep{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2}, \CRANpkg{scatterplot3d} v0.3-41 \citep{scatterplot3d}, \CRANpkg{recexcavAAR} v0.3.0 \citep{recexcavAAR}, \CRANpkg{deldir} v0.1 \citep{deldir}, ggtda v0.1 \citep{ggtda}, and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
 Runtime calculations were averaged over 10 iterations and are visualized as mean $\pm$ SD.

### 2020-03-19 tdabench-paper/tdabench.Rmd
revised math section

@@
 \section{Mathematics of Persistent Homology}
+<!--change circles to more generalizable object, e.g. n-sphere or n-ball-->
+An $n$-dimensional simplex is the convex hull of $n+1$ vertices.
+More intuitively, an $n$-dimensional simplex is the simplest $n$-dimensional object (e.g. a 0-simplex is a point, a 1-simplex is a line, a 2-simplex is a triangle, 3-simplex is a tetrahedron).
+These simplices can be glued together on common sub-simplices to form a simplicial complex (e.g. two triangles sharing a common side).
+In a simplicial complex, certain features will arise that can be characterized by Betti number: 0-dimensional features refer to connected components, 1-dimensional features refer to loops, 2-dimensional features refer to voids, and so on \citep{phom-survey}.
 
-This section summarizes the mathematics behind generating simplicial complexes on point clouds in the workflow of persistent homology. 
-An n-dimensional simplex is the convex hull of n+1 vertices.
-More intuitively, n-dimensional simplex is the simplest n-dimensional object (e.g. a 1 simplex is a line, a 2-simplex is a triangle, 3-simplex is a tetrahedron, etc.).
-These simplices can be glued together on a common subsimplices to form a simplicial complex (e.g. two triangles sharing a common side).
-In a simplicial complex, certain features will arise that can be characterized by dimension number (Betti number): 0-dimensional features refer to connected components, 1-dimensional features refer to holes, 2-dimensional features refer to cavities, etc.
-\cite{phom-survey}
-
-There are several different methods to construct a simplicial complex on a given point cloud S, but this paper focuses on the Vietoris-Rips and alpha complexes. 
-The Vietoris-Rips complex is perhaps the most common method for constructing a simplicial complex under persistent homology \cite{Rips-Complex}.
-In a point cloud of k points, a distance parameter, $\delta$, can be used to draw a circle of diameter $\delta$ around every single point in S.
-If the distance parameter is non-zero, then some of the resulting circles may intersect.
-If they intersect, a line is drawn to connect the points (which represent the centers of the growing circles).
-This group of points and lines form a simplicial complex.
-For each distance parameter, $\delta$, there will be some simplicial complex associated with it.
-As $\delta$ increases, different topological features may appear, persist, then disappear once $\delta$ has reached a threshold.
-Eventually, once $\delta$ reaches the maximum euclidean distance between any pair of points in the point cloud, a convex hull will form around all k points creating a (k-1)-dimensional simplex.
-At this stage, all features should have disappeared.
-A three column matrix can be created recording the Betti number of a certain feature, the distance, $\delta$, at which that feature appeared, and the $\delta$ at which that feature disappeared.
+There are several different methods to construct a simplicial complex on a given point cloud $S$, but this paper focuses on the Vietoris-Rips and alpha complexes.
+The Vietoris-Rips complex is perhaps the most common method for constructing a simplicial complex to calculate persistent homology \citep{Rips-Complex}.
+In a point cloud of $k$ points, a distance parameter, $\delta$, can be used to draw a circle of diameter $\delta$ around every point in $S$.
+If $\delta$ is sufficiently large, then some of the resulting circles may intersect.
+In this case, a line is drawn to connect the points at the center of the intersecting circles.
+This group of points and lines form the skeleton of a simplicial complex.
+For each distance parameter, $\delta$, there will be a single simplicial complex associated with it.
+As $\delta$ increases, different topological features may appear, persist, and eventually disappear.
+<!--Is below theoretical max accurate phrasing?-->
+Once $\delta$ reaches the maximum Euclidean distance between any pair of points in the point cloud, a convex hull will form around all $k$ points creating a $(k-1)$-dimensional simplex.
+A 3-column matrix can be created recording the Betti number of each feature, the $\delta$ at which that feature appeared, and the $\delta$ at which it disappeared.
 This matrix characterizes the persistent homology of that point cloud.
@@
   \includegraphics[height=2.5in]{../Figures/Final_Figures/Intro_Rips.png}
-  \caption{Basic Visualization of the Vietoris-Rips Complex.
-For a given parameter, $\delta$, $\delta$ diameter circles are drawn around each point.
+  \caption{\textbf{Basic Visualization of the Vietoris-Rips Complex.}
+  For a given parameter, $\delta$, $\delta$-diameter circles are drawn around each point.
   If two circles intersect, a point is drawn between their centers.
@@
   Features on the simplicial complex appear and die as $\delta$ increases.
-These features' dimensions, birth, and death are recorded in an n x 3 matrix.
+  These features' dimensions, birth, and death are recorded in an $n$x3 matrix.
   Eventually, the full convex hull is drawn ending the "filtration" process.}
@@
 
-Alpha complexes provide another method to generate simplicial complexes on the point cloud S.
+Alpha complexes provide another method to generate simplicial complexes on the point cloud $S$.
 Alpha complexes are closely linked with Voronoi diagrams and Delaunay triangulations.
-The Voronoi diagram gives each point in S a cell such that each every point in the cell is closest to that point in S.
-The nerve of a Voronoi diagram is equivalent to the Delaunay Triangulation\cite{alpha-complex}.
+The Voronoi diagram gives each point in $S$ a cell such that each every point in the cell is closest to that point in $S$.
+The nerve of a Voronoi diagram is equivalent to the Delaunay Triangulation \citep{alpha-complex}.
 Alpha complexes are simplicial complexes that are subsets of the Delaunay Triangulation.
-The parameter, $\alpha$, can describe the radius of a ball of each point in the point cloud S much like $\delta$ describes the diameter of a circle in the Vietoris-Rips complex.
-An Alpha complex at radius $\alpha$ is formed through connecting centers of intersecting regions formed by intersecting the union of $\alpha$ radius balls and the Voronoi diagram. 
-Eventually, once $\alpha$ is large enough, the full Delaunay Triangulation is formed.
-In between these stages, the birth and death of features at certain values of $\alpha$ can be captured in a three column persistent homology matrix much like the Vietoris-Rips complex.
+The parameter, $\alpha$, can describe the radius of a ball of each point in the point cloud $S$ much like $\delta$ describes the diameter of a circle in the Vietoris-Rips complex.
+An alpha complex at radius $\alpha$ is formed through connecting centers of intersecting regions formed by intersecting the union of $\alpha$-radius balls and the Voronoi diagram. 
+Once $\alpha$ is large enough, the full Delaunay Triangulation is formed.
+In between these stages, the birth and death of features at certain values of $\alpha$ can be captured in a 3-column persistent homology matrix much like the Vietoris-Rips complex.
 
@@
 From this matrix, the persistent homology matrix is calculated. 
-The size of the boundary matrix can describe the memory complexity on the RAM for persistent homology calculations. 
-We compare this memory complexity between alpha and Vietoris-Rips complexes for this project.
-Alpha complex calculations have a complexity of $O(n^\frac{d}{2})$ and Vietoris-Rips complex calculations have a complexity of $O(2^n)$ where n is the number of points and d is the dimension of the point cloud \citep{roadmap}.
+The size of the boundary matrix can describe the memory complexity on the random access memory (RAM) for persistent homology calculations. 
+We compare memory complexity between alpha and Vietoris-Rips complexes in this paper.
+Alpha complex calculations have a complexity of $O(n^{d/2})$ and Vietoris-Rips complex calculations have a complexity of $O(2^n)$ where $n$ is the number of points and $d$ is the point cloud dimension \citep{roadmap}.
 
@@
   \includegraphics[height=1.5in]{../Figures/Final_Figures/Intro_Alpha.png}
-  \caption{Basic visualization of the Alpha complex.
-For a given $\alpha$, $\alpha$ radius balls are drawn around each point, and the union is taken.
-Then an intersection between this union of $\alpha$ balls and the Vornoi diagram is taken.
-A connecting segment is drawn between points in adjacent voronoi cells once the $\alpha$ ball "fills" out the Voronoi diagram.
-As $\alpha$ continues to grow, more circles fill out the Voronoi cells.
-Eventually, once $\alpha$ is large enough, the Delaunay Triangulation is formed.}
+  \caption{\textbf{Basic visualization of the Alpha complex.}
+  For a given $\alpha$, $\alpha$-radius balls are drawn around each point, and the union of the balls is taken.
+  Then, an intersection between this union of $\alpha$-balls and the Vornoi diagram is taken.
+  A connecting segment is drawn between points in adjacent Voronoi cells once the $\alpha$-ball fills out the Voronoi diagram.
+  As $\alpha$ grows, more circles fill out the Voronoi cells.
+  Once $\alpha$ is large enough, the Delaunay Triangulation is formed.}
 \end{figure}
@@
 
-Benchmark data was collected twice - once on a local machine and once on a remote computing node, each of which featured 16 GB random access memory.
+Benchmark data was collected twice - once on a local machine and once on a remote computing node, each of which featured 16 GB RAM.
 Both datasets were compared for consistency and are publicly available at the repository linked below.
@@
 While Vietoris-Rips complexes can handle high-dimensional data well, calculation still significantly slows down when looking for higher dimension features. 
-Boundary matrix size rapidly increases and can overwhelm available random access memory even for small numbers of points (e.g. 100 points when looking for 4-dimensional features). 
+Boundary matrix size rapidly increases and can overwhelm available RAM even for small numbers of points (e.g. 100 points when looking for 4-dimensional features). 
 Looking for high-dimensional features in a large, high-dimensional point clouds remains a challenge. 

### 2020-03-24 tdabench-paper/tdabench.Rmd
Integrated Shael's edits!

@@
 The exact method of constructing these complexes is described in the Mathematics section.
-Persistent features more likely represent real geometric patterns, rather than noise.
+Essentially, we measures features that are "discovered" by the algorithm at a particular stage and disappear at a later stage. 
+The difference between these stages is the persistence. 
+Features with larger persistence more likely represent real geometric patterns, rather than noise.
 
@@
 <!--change circles to more generalizable object, e.g. n-sphere or n-ball-->
-An $n$-dimensional simplex is the convex hull of $n+1$ vertices.
+An $n$-dimensional simplex is the convex hull of $n+1$ points in some Euclidean space.
 More intuitively, an $n$-dimensional simplex is the simplest $n$-dimensional object (e.g. a 0-simplex is a point, a 1-simplex is a line, a 2-simplex is a triangle, 3-simplex is a tetrahedron).
 These simplices can be glued together on common sub-simplices to form a simplicial complex (e.g. two triangles sharing a common side).
-In a simplicial complex, certain features will arise that can be characterized by Betti number: 0-dimensional features refer to connected components, 1-dimensional features refer to loops, 2-dimensional features refer to voids, and so on \citep{phom-survey}.
+In a simplicial complex, topological features will arise that can be characterized by Betti numbers.
+Each Betti number $k$ counts the number of features in dimension $k$.
+Betti $0$ counts the number of connected components, Betti $1$ counts refer to loops, Betti $2$ counts the number of voids, and so on \citep{phom-survey}.
 
@@
 The Vietoris-Rips complex is perhaps the most common method for constructing a simplicial complex to calculate persistent homology \citep{Rips-Complex}.
-In a point cloud of $k$ points, a distance parameter, $\delta$, can be used to draw a circle of diameter $\delta$ around every point in $S$.
+In a point cloud of $k$ points in 2 dimensions, a distance parameter, $\delta > 0$, can be used to draw a circle of diameter $\delta$ around every point in $S$.
+For point clouds in 3 dimensions, spheres of diameter $\delta$ are drawn around each point, and hyperspheres of appropriate dimensions are drawn in higher dimensional spaces. 
+The remainder of this explanation will consider the 2 dimensional case of circles. 
 If $\delta$ is sufficiently large, then some of the resulting circles may intersect.
 In this case, a line is drawn to connect the points at the center of the intersecting circles.
+When a triple of points is connected, we add a triangle (2-simplex). When a quadruple of points is connected, we add tetrahedron (3-simplex) and so forth. 
+However, we only add simplices at most the dimension of the space of the point clouds (e.g. only up to 3-simplices are added in a 3-dimensional point cloud). 
 This group of points and lines form the skeleton of a simplicial complex.
@@
 Once $\delta$ reaches the maximum Euclidean distance between any pair of points in the point cloud, a convex hull will form around all $k$ points creating a $(k-1)$-dimensional simplex.
-A 3-column matrix can be created recording the Betti number of each feature, the $\delta$ at which that feature appeared, and the $\delta$ at which it disappeared.
+A 3-column matrix can be created recording the dimension of each feature, the $\delta$ at which that feature appeared, and the $\delta$ at which it disappeared.
 This matrix characterizes the persistent homology of that point cloud.
@@
 Alpha complexes provide another method to generate simplicial complexes on the point cloud $S$.
-Alpha complexes are closely linked with Voronoi diagrams and Delaunay triangulations.
-The Voronoi diagram gives each point in $S$ a cell such that each every point in the cell is closest to that point in $S$.
+Fo alpha complexes, we partition the whole space in which the data resides into cells such that each cell contains exactly one data point x, and the cell of that data point is the set of all points closer to x than any other data points.
+Such a partition is also known as a Voronoi diagram.
 The nerve of a Voronoi diagram is equivalent to the Delaunay Triangulation \citep{alpha-complex}.
 Alpha complexes are simplicial complexes that are subsets of the Delaunay Triangulation.
-The parameter, $\alpha$, can describe the radius of a ball of each point in the point cloud $S$ much like $\delta$ describes the diameter of a circle in the Vietoris-Rips complex.
-An alpha complex at radius $\alpha$ is formed through connecting centers of intersecting regions formed by intersecting the union of $\alpha$-radius balls and the Voronoi diagram. 
+The parameter, $\alpha$, can describe the radius of a ball (dimension matches dimension of the space) of each point in the point cloud $S$ much like $\delta$ describes the diameter of a circle in the Vietoris-Rips complex.
+We first intersect the $\alpha$ radius balls with their own Voronoi cell and then search for intersections of these subsetted balls to form simplices. 
 Once $\alpha$ is large enough, the full Delaunay Triangulation is formed.
 In between these stages, the birth and death of features at certain values of $\alpha$ can be captured in a 3-column persistent homology matrix much like the Vietoris-Rips complex.
+One key difference from the Vietoris-Rips complex is that edges can only form between neighboring points in the alpha complex.
 
-In both methods, the boundary matrix records all simplicial complexes for a given parameter value. 
-From this matrix, the persistent homology matrix is calculated. 
+In both methods, the boundary matrix records all simplicial complexes for each parameter value ($\delta$ for Vietoris-Rips complexes and $\alpha$ for alpha complexes). 
+Calculating persistent homology is divided into two steps: (1) forming the boundary matrix, and (2) reducing the boundary matrix to be able to read off the topological features of each dimension and their birth/death values of the parameter.
+The second step can be computed in at most O(k^3) steps where k is the number of rows (and columns) of the boundary matrix. 
 The size of the boundary matrix can describe the memory complexity on the random access memory (RAM) for persistent homology calculations. 
 We compare memory complexity between alpha and Vietoris-Rips complexes in this paper.
-Alpha complex calculations have a complexity of $O(n^{d/2})$ and Vietoris-Rips complex calculations have a complexity of $O(2^n)$ where $n$ is the number of points and $d$ is the point cloud dimension \citep{roadmap}.
+
+Alpha complex calculations have a run time complexity of $O(n^{d/2})$ and Vietoris-Rips complex calculations have a run time complexity of $O(2^n)$ where $n$ is the number of points and $d$ is the point cloud dimension \citep{roadmap}.
+Vietoris-Rips's run time and memory are exponential with regards to point number (but constant with data dimension) in contrast to alpha complexes where run time and memory are polynomial wih point number (but exponential with data dimension).
+Therefore, we can predict that low dimensional point clouds favor alpha complexes but fewer points in higher dimension favors Vietoris-Rips complexes.
 
@@
 We use \CRANpkg{readr} v1.3.1 to read rectangular data \citep{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2}, \CRANpkg{scatterplot3d} v0.3-41 \citep{scatterplot3d}, \CRANpkg{recexcavAAR} v0.3.0 \citep{recexcavAAR}, \CRANpkg{deldir} v0.1 \citep{deldir}, ggtda v0.1 \citep{ggtda}, and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
-Runtime calculations were averaged over 10 iterations and are visualized as mean $\pm$ SD.
+Runtime calculations were averaged over 10 iterations and are visualized as mean $\pm$ standard deviation.
 Datasets were generated by sampling functions in base R to generate points uniformly distributed over circles (dimension = 2), spheres (dimension = 3), filled squares (dimension = 2), filled cubes (dimension = 3, 4), and tori (dimension = 3).
-The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after considering available computational resources.
+The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empirical limits chosen after considering available computational resources.
 For consistency between software libraries, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} R packages.
@@
 Figure \ref{fig:fin} shows that at high point cloud sizes, GUDHI’s alpha complex outperforms Ripser. 
-Theoretically, alpha complexes gain polynomial complexity as the number of points increases whereas Vietoris-Rips complexes gain exponentially \citep{roadmap}.
+Theoretically, alpha complexes gain polynomial run time complexity as the number of points increases whereas Vietoris-Rips complexes gain exponential run time complexity \citep{roadmap}.
 Specifically, alpha complexes are $O(n^{d/2})$ and Vietoris-Rips complexes are $O(2^n)$ where $n$ is the number of points on a point cloud and $d$ is the dimensions on the point cloud. 
@@
 It should be stated that if a point cloud is compatible with both complexes, both analyses should be performed as there may be a variation the persistent homology matrix. 
+This is because alpha complexes satisfy the Nerve theorem ~\citep{alpha-complex}. 
+This essentially means that alpha complexes are topologically equivalent to the true underlying topology of the datase whereas Rips complexes are approximations ~\citep{Rips-Complex}.
 Among the tested libraries, Ripser (wrapped by \CRANpkg{TDAstats}) has the fastest calculation time. 

### 2020-03-24 tdabench-paper/tdabench.Rmd

@@
 The exact method of constructing these complexes is described in the Mathematics section.
-Essentially, we measures features that are "discovered" by the algorithm at a particular stage and disappear at a later stage. 
+Essentially, we measure features that are discovered by the algorithm at a particular stage and disappear at a later stage. 
 The difference between these stages is the persistence. 
@@
 <!--change circles to more generalizable object, e.g. n-sphere or n-ball-->
-An $n$-dimensional simplex is the convex hull of $n+1$ points in some Euclidean space.
+An $n$-dimensional simplex is the convex hull of $n+1$ points in a Euclidean space.
 More intuitively, an $n$-dimensional simplex is the simplest $n$-dimensional object (e.g. a 0-simplex is a point, a 1-simplex is a line, a 2-simplex is a triangle, 3-simplex is a tetrahedron).
@@
 In a simplicial complex, topological features will arise that can be characterized by Betti numbers. 
-Each Betti number $k$ counts the number of features in dimension $k$.
-Betti $0$ counts the number of connected components, Betti $1$ counts refer to loops, Betti $2$ counts the number of voids, and so on \citep{phom-survey}.
+Each Betti number, denoted by $B_k$, $k$ counts the number of features in dimension $k$. 
+$B_0$ counts the number of connected components, $B_1$ counts refer to loops, $B_2$ counts the number of voids, and so on \citep{phom-survey}.
 
@@
 In a point cloud of $k$ points in 2 dimensions, a distance parameter, $\delta > 0$, can be used to draw a circle of diameter $\delta$ around every point in $S$.
-For point clouds in 3 dimensions, spheres of diameter $\delta$ are drawn around each point, and hyperspheres of appropriate dimensions are drawn in higher dimensional spaces. 
-The remainder of this explanation will consider the 2 dimensional case of circles. 
+For point clouds in 3 dimensions, spheres of diameter $\delta$ are drawn around each point. 
+For dimensions $k$ greater than 3, a $k$-dimensional hypersphere is drawn around each point. 
+The remainder of this explanation will focus on the 2-dimensional case. 
 If $\delta$ is sufficiently large, then some of the resulting circles may intersect.
 In this case, a line is drawn to connect the points at the center of the intersecting circles.
-When a triple of points is connected, we add a triangle (2-simplex). When a quadruple of points is connected, we add tetrahedron (3-simplex) and so forth. 
+When a triple of points is connected, we add a triangle (2-simplex). 
+When a quadruple of points is connected, we add a tetrahedron (3-simplex) and so forth. 
 However, we only add simplices at most the dimension of the space of the point clouds (e.g. only up to 3-simplices are added in a 3-dimensional point cloud). 
@@
 As $\delta$ increases, different topological features may appear, persist, and eventually disappear.
-<!--Is below theoretical max accurate phrasing?-->
+
 Once $\delta$ reaches the maximum Euclidean distance between any pair of points in the point cloud, a convex hull will form around all $k$ points creating a $(k-1)$-dimensional simplex.
@@
 Alpha complexes provide another method to generate simplicial complexes on the point cloud $S$.
-Fo alpha complexes, we partition the whole space in which the data resides into cells such that each cell contains exactly one data point x, and the cell of that data point is the set of all points closer to x than any other data points.
+For alpha complexes, we partition the whole space in which the data resides into cells such that each cell contains exactly one data point $x$, and the cell of that data point is the set of all points closer to $x$ than any other data points.
 Such a partition is also known as a Voronoi diagram.
@@
 It should be stated that if a point cloud is compatible with both complexes, both analyses should be performed as there may be a variation the persistent homology matrix. 
-This is because alpha complexes satisfy the Nerve theorem ~\citep{alpha-complex}. 
-This essentially means that alpha complexes are topologically equivalent to the true underlying topology of the datase whereas Rips complexes are approximations ~\citep{Rips-Complex}.
+This is because alpha complexes satisfy the Nerve Theorem \citep{alpha-complex}, which implies that they are topologically equivalent to the true underlying topology of the dataset; in contrast, Vietoris-Rips complexes only approximate the underlying topology \citep{Rips-Complex}.
 Among the tested libraries, Ripser (wrapped by \CRANpkg{TDAstats}) has the fastest calculation time. 
@@
 GUDHI and Dionysus, however, do have the option of returning the boundary matrix, which Ripser does not store for the sake of optimization.
-While we did not measure Ripser's boundary matrix object size, Ripser's C++ code (outside its implementation in R) demonstrates memory advantages ~\citep{Ripser}
+While we did not measure Ripser's boundary matrix object size, Ripser's C++ code (outside its implementation in R) demonstrates memory advantages \citep{Ripser}
 

### 2020-04-05 tdabench-paper/tdabench.Rmd
updated citation and title
revised abstract

@@
 ---
-title: Comparing R packages for calculation of persistent homology
+title: Benchmarking R packages for calculation of persistent homology
 author:
@@
   Several persistent homology software libraries have been implemented in R.
-The Dionysus (TDA), GUDHI (TDA), and Ripser C++ libraries (TDAstats) provide Vietoris-Rips complex functionality.
-GUDHI  (TDA package) also provides alpha complex functionality.
-These software represent powerful analysis tools but are limited computationally.
-These limitations and guidelines for which persistent homology libraries to use have not been specified in R.
-We benchmarked the all libraries on a High-Powered Cluster for runtime.
-We also compared alpha and Vietoris-Rips for memory burden.
-Alpha complexes run significantly faster than all Vietoris-Rips complex libraries for all data sets for data dimension less than or equal to three.
-Alpha complexes were not computable for data dimension greater than 3.
-TDAstats had the fastest runtimes for all Vietoris-Rips complexes.
-We recommend performing persistent homology with GUDHI alpha complexes for data dimension less than or equal to three.
-For higher dimension, we recommend TDAstats Vietoris-Rips Complex.
+Specifically, the Dionysus, GUDHI, and Ripser libraries have been wrapped in R by the TDA and TDAstats packages on CRAN.
+They provide Vietoris-Rips and alpha complex functionality.
+These software represent powerful analysis tools that are currently computationally expensive and, to our knowledge, have not been formally benchmarked in R.
+Here, we compare the 2 R packages and the 3 underlying libraries for runtime and memory.
+We find that datasets with less than 3 dimensions can be evaluated with persistent homology fastest by the GUDHI library in the TDA package.
+For larger dimensional datasets, the Ripser library in the TDAstats package is best.
+Analysis of runtime and memory growth is also included in this report.
 output:

### 2020-12-14 Figures/Final_Figures_Reproducible.R
updated based on reviewer comments

@@ -2,6 +2,8 @@ library(dplyr)
 library(ggplot2)
 library(reshape)
 library(ggforce)
+library(scales)
+
 
 ####Reading in Data####
 #Reading in the time data
@@ -24,12 +26,15 @@ data.time$avg.time <- apply(data.time[,7:16],1,mean)
 data.time$std <- apply(data.time[,7:16],1,sd)
 data.time$min.time <- apply(data.time[,7:16],1,min)
 data.time$max.time <- apply(data.time[,7:16],1,max)
+data.time$med.time <- apply(data.time[,7:16],1,median)
+
 
 #Create Factor labels for libraries
 data.time$library <- factor(data.time$library,
                             levels = c("stats","Dionysus","GUDHI", "GUDHIalpha"),
                             labels = c("TDAstats (Ripser)", 
-                                       "TDA (Dionysus)", "TDA (GUDHI Rips)", "TDA (GUDHI Alpha)"))
+                                       "TDA (Dionysus)", "TDA (GUDHI Rips)", 
+                                       "TDA (GUDHI Alpha)"))
 
 #Read in object size data
 data.mem <- read.csv("./mem1.csv", header = FALSE)
@@ -42,7 +47,8 @@ colnames(data.mem) <- c("measure.type", "point.cloud", "point.cloud.dim",
 data.mem$library <- factor(data.mem$library,
                            levels = c("stats","Dionysus","GUDHI", "GUDHIalpha"),
                            labels = c("TDAstats (Ripser)", 
-                                      "TDA (Dionysus)", "TDA (GUDHI Rips)", "TDA (GUDHI Alpha)"))
+                                      "TDA (Dionysus)", "TDA (GUDHI Rips)", 
+                                      "TDA (GUDHI Alpha)"))
 
 
 ####Figure_1####
@@ -53,27 +59,31 @@ data.fig.1 <- subset(data.time, point.cloud == "torus" & feat.dim == 2 &
                        library != "TDA (GUDHI Alpha)") 
 
 #Point graph; X: Num.points; Y: Runtime; Color: TDA library
-fig.1 <- ggplot(data.fig.1, aes(x=num.points, y=avg.time, color=library, shape=library)) + 
+fig.1 <- ggplot(data.fig.1, aes(x=num.points, y=med.time, color=library, shape=library)) + 
   geom_point()+
   geom_errorbar(data.fig.1, mapping = aes(x=num.points, 
-                                          ymin=avg.time - std, 
[…]

### 2020-12-14 Figures/Final_Figures_Reproducible.R
changed average to median

@@ -67,7 +67,7 @@ fig.1 <- ggplot(data.fig.1, aes(x=num.points, y=med.time, color=library, shape=l
   labs(color = "TDA Library",
        shape = "TDA Library",
        x = "Number of Points on Torus",
-       y = "Average Runtime (s)",
+       y = "Median Runtime (s)",
        title = "Rips Complex Runtimes on 3D Torus Point Clouds",
        subtitle = "") + 
   scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), 
@@ -110,7 +110,7 @@ fig.2 <- ggplot(data.fig.2, aes(x=num.points, y=med.time, color=library, shape=l
   labs(color = "TDA Library",
        shape = "TDA Library",
        x = "Number of Points on N-Sphere",
-       y = "Average Runtime (s)",
+       y = "Median Runtime (s)",
        title = "Rips Complex Runtimes For n-Dimensional Spheres",
        subtitle = "") +
   scale_x_continuous(limits=c(0,550), breaks = c(100, 200, 300, 400, 500)) + 
@@ -155,7 +155,7 @@ fig.3 <- ggplot(data.fig.3, aes(x=feat.dim, y=med.time, color=num.points, shape=
   labs(color = "Number of \nPoints",
        shape = "Number of \nPoints",
        x = "Feature Dimensions",
-       y = "Average Runtime (s)",
+       y = "Median Runtime (s)",
        title = "Runtimes For Extracting Dimensional \nFeatures on an 8 Dimensional Box",
        subtitle = "") + scale_x_continuous(limits=c(0, 8), breaks=seq(0,8,1)) +
   scale_y_continuous(limits=c(0, 3)) +
@@ -199,7 +199,7 @@ fig.4 <- ggplot(data.fig.4, aes(x=num.points, y=med.time, color=point.cloud.dim,
   labs(color = "Annulus \nDimension",
        shape = "Annulus \nDimension",
        x = "Number of Points",
-       y = "Average Runtime (s)",
+       y = "Median Runtime (s)",
        title = "Rips vs Alpha Complex Runtimes For Extracting \n1 Dimensional Features on N-dimensional Annuluses",
        subtitle = "") + 
   scale_x_continuous(limits=c(0,550), breaks = c(100, 200, 300, 400, 500)) + 
@@ -363,7 +363,7 @@ fig.6 <- ggplot(data.fig.6, aes(x=num.points, y=med.time, color=library, shape=l
[…]

### 2020-12-14 tdabench-paper/tdabench.Rmd
Updated abstract
replaced instances of average/mean with median to reflect new figures

@@
 These software represent powerful analysis tools that are currently computationally expensive and, to our knowledge, have not been formally benchmarked in R.
-Here, we compare the 2 R packages and the 3 underlying libraries for runtime and memory.
+Here, we provide an analysis of runtime and memory growth for the 2 R packages and the 3 underlying libraries.
 We find that datasets with less than 3 dimensions can be evaluated with persistent homology fastest by the GUDHI library in the TDA package.
 For larger dimensional datasets, the Ripser library in the TDAstats package is best.
-Analysis of runtime and memory growth is also included in this report.
 output:
@@
 We use \CRANpkg{readr} v1.3.1 to read rectangular data \citep{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2}, \CRANpkg{scatterplot3d} v0.3-41 \citep{scatterplot3d}, \CRANpkg{recexcavAAR} v0.3.0 \citep{recexcavAAR}, \CRANpkg{deldir} v0.1 \citep{deldir}, ggtda v0.1 \citep{ggtda}, and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
-Runtime calculations were averaged over 10 iterations and are visualized as mean $\pm$ standard deviation.
+Median runtime calculations are shown along with the minimum and maximum of 10 iterations per benchmark.
 Datasets were generated by sampling functions in base R to generate points uniformly distributed over circles (dimension = 2), spheres (dimension = 3), filled squares (dimension = 2), filled cubes (dimension = 3, 4), and tori (dimension = 3).
@@
 Computing persistent homology of a canonical torus grants quick insight into efficiency of each library (Figure \ref{fig:tor}).
-Dionysus exhibits the longest average runtime, and, although Ripser and GUDHI have similar runtimes for smaller point clouds, as the number of points increases Ripser eventually has a significant lead.
+Dionysus exhibits the longest median runtime, and, although Ripser and GUDHI have similar runtimes for smaller point clouds, as the number of points increases Ripser eventually has a significant lead.
 Next, we compare library performance with multiple canonical datasets to ensure that the noted pattern generalizes.
@@
     \textbf{Calculating persistent homology of a torus with three TDA libraries.}
-    Average runtime (mean $\pm$ SD, $n=10$ iterations per data point) for each TDA library (denoted by color) is plotted against point cloud size.
+    Median runtime (min to max, $n=10$ iterations per data point) for each TDA library (denoted by color) is plotted against point cloud size.
     Homological features up to 2 dimensions were calculated.
@@
     \textbf{Calculating persistent homology of round point clouds of varying dimensions with three TDA libraries.}
-    Average runtime (mean $\pm$ SD, $n=10$ iterations per data point) for each TDA library (denoted by color) is plotted against point cloud size and faceted by data dimension.
+    Median runtime (min to max, $n=10$ iterations per data point) for each TDA library (denoted by color) is plotted against point cloud size and faceted by data dimension.
     Left panel compares library performance for a 2-dimensional circular point cloud; center panel for a 3-dimensional spherical point cloud; and right panel for a 4-dimensional hyperspherical point cloud.
@@
     \textbf{Comparison of Vietoris-Rips complex persistent homology calculation as a function of feature dimension.}
-    Average runtime (mean $\pm$ SD, $n=10$ iterations per data point) for various point cloud sizes (denoted by color) is plotted against calculated feature dimension and faceted by TDA libary.
+    Median runtime (min to max, $n=10$ iterations per data point) for various point cloud sizes (denoted by color) is plotted against calculated feature dimension and faceted by TDA libary.
     Persistent homology was calculated on a uniformly distributed random sample of points contained within a 1 unit, 8-dimensional cube.
@@
     \textbf{Comparing persistent homology calculation between Vietoris-Rips and alpha complices.}
-    Average runtime (mean $\pm$ SD, $n=10$ iterations per data point) for various data dimensions (denoted by color) are plotted against point cloud size and faceted by type of simplicial complex.
+    Median runtime (min to max, $n=10$ iterations per data point) for various data dimensions (denoted by color) are plotted against point cloud size and faceted by type of simplicial complex.
     Maximum of feature dimension was kept constant at 1.
@@
     \textbf{Comparing memory usage of persistent homology calculation using a Vietoris-Rips versus an alpha complex.}
-    Mean boundary matrix size, a proxy for memory usage, plotted against point cloud size varying by point cloud structure (facet) and simplicial complex type (color, shape).
+    Boundary matrix size, a proxy for memory usage, plotted against point cloud size varying by point cloud structure (facet) and simplicial complex type (color, shape).
     Data dimension and feature dimension were held constant at 3 and 2, respectively.
@@
     \textbf{Factors affecting boundary matrix size in Vietoris-Rips and alpha complices.}
-    Mean boundary matrix size plotted against point cloud size varying by point cloud structure (right, faceted), simplicial complex type (left, right), and feature dimension (color).
+    Boundary matrix size plotted against point cloud size varying by point cloud structure (right, faceted), simplicial complex type (left, right), and feature dimension (color).
     Whereas the boundary matrix for a Vietoris-Rips complex (left) is independent of point cloud structure, the same is not true for an alpha complex (right, faceted).
@@
     \textbf{Runtime comparison of persistent homology calculation between Ripser's Vietoris-Rips and GUDHI's alpha complex functionality.}
-    Average runtime (mean $\pm$ SD, $n=10$ iterations per data point) for various 3-dimensional point cloud structures (facet) plotted against point cloud size for each library (color).
+    Median runtime (min to max $n=10$ iterations per data point) for various 3-dimensional point cloud structures (facet) plotted against point cloud size for each library (color).
     Benchmarking was conducted on an annulus (top-left), a sphere (top-right), a torus (bottom-left), and a cube (bottom-right).

### 2020-12-28 tdabench-paper/tdabench.Rmd
added memory benchmarks

--- 2020-04-05 tdabench-paper/tdabench.Rmd
@@
 These software represent powerful analysis tools that are currently computationally expensive and, to our knowledge, have not been formally benchmarked in R.
-Here, we compare the 2 R packages and the 3 underlying libraries for runtime and memory.
+Here, we provide an analysis of runtime and memory growth for the 2 R packages and the 3 underlying libraries.
 We find that datasets with less than 3 dimensions can be evaluated with persistent homology fastest by the GUDHI library in the TDA package.
-For larger dimensional datasets, the Ripser library in the TDAstats package is best.
-Analysis of runtime and memory growth is also included in this report.
+For higher dimensional datasets, the Ripser library in the TDAstats package is best.
 output:
@@
 
-\section{Introduction}
+```{r setup, echo = FALSE}
+suppressPackageStartupMessages(library(ggplot2))
+suppressPackageStartupMessages(library(magrittr))
+suppressPackageStartupMessages(library(dplyr))
+suppressPackageStartupMessages(library(broom))
+```
+
+# Introduction
 
@@
 
-\section{Mathematics of Persistent Homology}
+# Mathematics of Persistent Homology
 <!--change circles to more generalizable object, e.g. n-sphere or n-ball-->
@@
 
-\section{Methods}
+# Methods
 
@@
 
-\section{Results}
+# Results
 
@@
 
-Vietoris-Rips and alpha complices differ not only in runtime, but also in memory usage.
-GUDHI conveniently allows measurement of the simplicial complex boundary matrix, a reasonable proxy for total memory usage in calculation of persistent homology.
-Since Ripser and Dionysus do not have similar capabilities, they are excluded from memory benchmarks.
-Vietoris-Rips complices show rapid growth in object size as point cloud size increases; in contrast, alpha complices consistently maintain low, linearly-growing memory use (Figure \ref{fig:mem1}).
-The growth pattern differences mirror findings for runtime growth (Figure \ref{fig:ann}).
-Interestingly, different factors determine memory usage in the two types of simplicial complex.
-Whereas memory use in calculation of a Vietoris-Rips complex is dependent on point cloud size and feature dimension, memory use for an analogous calculation with an alpha complex is also dependent on the structure of the point cloud itself (Figure \ref{fig:mem2}).
-Within the limited set of tested point cloud structures, the sphere requires the least memory.
-Table \ref{tbl:limit} lists the maximum point cloud size for which persistent homology calculations could be completed for each library.
+```{r powerlaw, echo=FALSE}
+# input data
+df_mem <- read.csv("../memory-bench/memory-use.csv") %>%
+  mutate(Memory = Memory / 1024)
 
-\begin{figure}
-  \centering
-  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig5a.png}
-  \caption{
-    \textbf{Comparing memory usage of persistent homology calculation using a Vietoris-Rips versus an alpha complex.}
-    Mean boundary matrix size, a proxy for memory usage, plotted against point cloud size varying by point cloud structure (facet) and simplicial complex type (color, shape).
-    Data dimension and feature dimension were held constant at 3 and 2, respectively.
-    Point cloud structures for which memory usage was calculated were 3-dimensional annulus (top-left), 3-dimensional sphere (top-right), torus (bottom-left), and 3-dimensional cube (bottom-right).
-    Similar to runtime results, alpha complex calculation showed linear growth and Vietoris-Rips complex calculation showed polynomial growth (see Supplement for regression details).}
-\label{fig:mem1}
-\end{figure}
+# setup df
+pl_exponents <- expand.grid(c("TDAstats", "GUDHI", "Dionysus"),
+                            2:4,
+                            c("circle", "annulus", "torus", "uniform"))
 
-\begin{widefigure}
-  \centering
-  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig5t.png}
-  \caption{
-    \textbf{Factors affecting boundary matrix size in Vietoris-Rips and alpha complices.}
-    Mean boundary matrix size plotted against point cloud size varying by point cloud structure (right, faceted), simplicial complex type (left, right), and feature dimension (color).
-    Whereas the boundary matrix for a Vietoris-Rips complex (left) is independent of point cloud structure, the same is not true for an alpha complex (right, faceted).
-    For the Vietoris-Rips complex, the decreasing number of data points plotted for successive feature dimensions reflects the computational requirements of increasing point cloud size.}
-\label{fig:mem2}
-\end{widefigure}
+# return power law exponent from data
+calc_pl_exp <- function(all_data, engine, dim, shape) {
+  # only >= 100 points
+  all_data <- all_data %>%
+    filter(NumPoints >= 100 &
+           Engine == engine &
+           Dimension == dim &
+           DataType == shape) %>%
+    mutate(LogPts = log10(NumPoints),
+           LogMem = log10(Memory))
 
-\begin{table}
-\centering
-\small
-\begin{tabular}{|c|c|c|c|c|}
-\hline
-Feature & Ripser & Dionysus & GUDHI Rips & GUDHI Alpha \\
-Dimension & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit & Point Cloud Limit\\
-\hline
-1 & Not Reached & Not Reached & Not Reached & Not Reached \\ 
-2 & Not Reached & 200 points & 280 points & Not Reached \\ 
-3 & 300 points & 100 points & 125 points & Could Not Compute \\
-4 & 100 points & 50 points & 75 points & Could Not Compute \\ 
-\hline
-\end{tabular}
-\caption{During the benchmarking, certain libraries were not able to calculate persistent homology for sufficiently complex point clouds. 
-Limiting factors were point cloud number and feature dimension calculation. 
-For Vietoris-Rips complices, Ripser had the best performing capacity among the three libraries. 
-Alpha complexes were not computable for any point cloud greater than 3 dimensions, so dimensional features 3 and higher could not be computed.}
-\label{tbl:limit}
-\end{table}
+  lm_mod <- lm(LogMem ~ LogPts, data = all_data)
+  tidy_mod <- tidy(lm_mod)
 
-In real life scenarios, practitioners likely care most about comparative performance of the fastest Vietoris-Rips complex library and the fastest alpha complex library.
-Figure \ref{fig:fin} illustrates this comparison, showing that for 3-dimensional point clouds, GUDHI's alpha complex library outpaces Ripser's Vietoris-Rips functionality.
+  return(tidy_mod$estimate[2])
+}
+```
 
-\section{Discussion}
+In addition to runtime differences, the three Vietoris-Rips homology engines differ in memory use.
+After crossing a threshold between 50 and 100 points, all three engines appeared to follow power law growth, with a linear trend on log-log plots (Figure \ref{fig:mem}).
+However, for all combinations of point cloud dimension and shape, TDAstats used the least memory and Dionysus used the most, with TDAstats also growing with the smallest power law exponent as the number of points increased.
+As an example, for 3-dimensional annulus point clouds, GUDHI memory use grew with a power law exponent of
+`r calc_pl_exp(df_mem, "GUDHI", 3, "annulus") %>% signif(digits = 2)`,
+Dionysus use grew with an exponent of
+`r calc_pl_exp(df_mem, "Dionysus", 3, "annulus") %>% signif(digits = 2)`,
+and TDAstats use grew with an exponent of
+`r calc_pl_exp(df_mem, "TDAstats", 3, "annulus") %>% signif(digits = 2)`.
+The entire table of exponents can be found in the Supplement (GitHub), but Figure \ref{fig:mem} clearly illustrates that TDAstats was, for all tested cases, the optimal Vietoris-Rips engine in the context of memory use.
+
+```{r memory, echo=FALSE, fig.cap="\\label{fig:mem}\\textbf{Comparing memory use of Vietoris-Rips persistent homology engines.} Each column title corresponds to point cloud dimension; each row title lists point cloud shape; each persistent homology engine is represented by points of a distinct shape and color.
+For point clouds containing more than 50 points, there appears to be a linear trend on the log-log axes.
+Data for 2- and 4-dimensional tori were not collected because a torus does not trivially generalize to dimensions other than 3.
+Missing points for GUDHI and Dionysus in the 4-dim plots indicate that persistent homology calculation was terminated since memory requirement exceeded available RAM (32 GB)."}
+df_mem <- df_mem %>%
+  mutate(Dimension = factor(Dimension, levels = 2:4, labels = c("2-dim", "3-dim", "4-dim")))
+
+ggplot(df_mem, aes(x = NumPoints, y = Memory, color = Engine, shape = Engine)) +
+  geom_point() +
+  scale_x_continuous(trans = "log10") +
+  scale_y_continuous(trans = "log10") +
+  facet_grid(DataType ~ Dimension) +
+  xlab("Number of points") +
+  ylab("Memory (GB)") +
+  theme_bw()
+```
+
+# Discussion
+
 As persistent homology calculations continue to become a more popular tool to analyze complex multidimensional data, it will be important to understand from a computational perspective which method to use.

### 2020-12-28 tdabench-paper/tdabench.R

@@ -1,0 +1,4 @@
+## ----memory-------------------------------------------------------------------
+df_mem <- read.csv("../memory-bench/memory-use.csv")
+head(df_mem)
+

### 2020-12-28 tdabench-paper/tdabench.Rmd
added comparison between TDA and TDAstats

@@
 
+On average, Ripser wrapped by \CRANpkg{TDAstats} computed persistent homology of a Vietoris-Rips complex faster and with less memory than either GUDHI or Dionysus wrapped by \CRANpkg{TDA}.
+Thus, when efficiency is critical, useRs would likely find \CRANpkg{TDAstats} the appropriate library.
+However, \CRANpkg{TDA} contains an entire library of features not available in \CRANpkg{TDAstats}.
+Specifically, \CRANpkg{TDA} allows kNN density estimation, kernel density estimation, density-based clustering, and dendrogram visualization, among other useful functionality.
+When computational resources are plenty, when point clouds are small and low-dimensional, or when the aforementioned functionality is required, \CRANpkg{TDA} will likely be more appropriate than \CRANpkg{TDAstats}.
+Both packages are hosted on CRAN.
+
 While Vietoris-Rips complexes can handle high-dimensional data well, calculation still significantly slows down when looking for higher dimension features. 
-Boundary matrix size rapidly increases and can overwhelm available RAM even for small numbers of points (e.g. 100 points when looking for 4-dimensional features). 
+Boundary matrix size rapidly increases and can overwhelm available RAM even for small numbers of points. 
 Looking for high-dimensional features in a large, high-dimensional point clouds remains a challenge. 
@@
 
-\section{Acknowledgments}
+# Acknowledgments
+
 The authors thank Emily Dolson, PhD for assistance with using a high-performance computing node for data collection.

### 2020-12-28 tdabench-paper/tdabench.Rmd
updated discussion for added memory benchmarks

@@
 This is because alpha complexes satisfy the Nerve Theorem \citep{alpha-complex}, which implies that they are topologically equivalent to the true underlying topology of the dataset; in contrast, Vietoris-Rips complexes only approximate the underlying topology \citep{Rips-Complex}.
-Among the tested libraries, Ripser (wrapped by \CRANpkg{TDAstats}) has the fastest calculation time. 
+Among the tested Vietoris-Rips engines, Ripser (wrapped by \CRANpkg{TDAstats}) has the fastest calculation time. 
 GUDHI and Dionysus (wrapped by \CRANpkg{TDA}) significantly fall behind as feature dimension and number of points increase.
-GUDHI and Dionysus, however, do have the option of returning the boundary matrix, which Ripser does not store for the sake of optimization.
-While we did not measure Ripser's boundary matrix object size, Ripser's C++ code (outside its implementation in R) demonstrates memory advantages \citep{Ripser}
 
-On average, Ripser wrapped by \CRANpkg{TDAstats} computed persistent homology of a Vietoris-Rips complex faster and with less memory than either GUDHI or Dionysus wrapped by \CRANpkg{TDA}.
+On average, Ripser computed persistent homology of a Vietoris-Rips complex with less memory than either GUDHI or Dionysus.
 Thus, when efficiency is critical, useRs would likely find \CRANpkg{TDAstats} the appropriate library.

### 2020-12-29 tdabench-paper/review-response.txt
added most of reviewer response

        ---------- Reviewer ----------

        ## Overview ##

        I cannot comment a great deal on the overall suitability to the R
        community, as TDA methods and persistent homology are not areas I have
        a great deal of expertise in. The paper discussion and benchmarking
        analysis both seem sensible and well done. In general the methods they
        used are up to date, however the author's approach has enough
        limitations (detailed below) I think it should be revised before
        acceptance.

        ### Benchmarking measures

        For synthetic benchmarking generally the minimum and median are more
        informative measures then the mean or standard deviation. Both the
        mean and standard deviation (used in this paper) are highly positively
        skewed by outliers. For benchmarks generally any noise _must_ be
        positive, e.g. there are no random processes that will cause the code
        to run _faster_ than the minimum. See <URL> for further discussion on
        these points. The mean and standard deviation are poor enough
        indicators I think the figures should be redone, possibly as box plot
        with an explicit minimum or as point clouds with all measurements. I 
        believe it was mentioned there were only 10 iterations per run, so
        showing all points might be feasible.

The reviewer is correct, the data would be better visualized with more robust
measures. To correct our previous figures, we have visualized the runtime
data using minimum, median, and maximum. Unfortunately, visualizing each
iteration (even with jitter) resulted in cluttered plots that were challening
to interpret. We thank the reviewer for including a URL that discussed
appropriate measures for our data; our correction conforms with the reviewer's
suggestion and the linked document.

        In addition I would suggest the runtime figures (and possibly the memory
        as well) be shown with logarithmic scale. This is because generally you
        want to compare the relative runtime between two points, not the
        absolute runtime. On a logorithmic scale a 2x increase in runtime is the
        same distance if a benchmark changes from 10ms to 20ms and from 100ms to
        200ms, but in an absolute scale the second change looks much larger.
        Using a logorithmic scale would also make the plots more descriptive for
        the smaller datasets.

We agree fully with the reviewer's point and corresponding suggestion. As such,
the runtime and memory data have been visualized on a logarithmic scale.

        ### Benchmarking Memory

        The authors mention issues measuring the memory of the processes,
        because most of the computation occurs outside of R's managed heap.
        They work around this by using a proxy measurement of memory usage,
        however this has the severe limitation that it only applies to one of
        the 3 implementations benchmarked. This limitation is extremely severe,
        and I think must be changed before the paper can be accepted.

        Fortunately there are tools the authors can use to accomplish this.
        `bench::bench_process_memory()` can be used to sample the memory for
        the entire R process (including child processes and memory allocated
        in C++). If the authors run this after their benchmark is run and look
        at the `max` column it will give them the maximum memory usage of the
        process (and its children). Another alternative is running their
        processes with the GNU time command (https://www.gnu.org/software/time/)
        and looking at the maximum resident set size in KB. e.g.
        `time -f '%M' R -f my-benchmark-script.R`. This does basically the same
        thing as `bench::bench_process_memory()`.

        Using the suggested approach would work for all of the benchmarks in
        this paper, and would make it much more informative. Memory usage is a
        very important consideration, particularly for R which is generally
        very memory hungry.

We agree with the reviewer's point and have used `bench::bench_process_memory`
to collect memory benchmark data for all point cloud size, dimension, and
shape combinations. The resulting data has been visualized in Figure 7. Text
in the results and discussion sections has been adjusted to reflect the newly
added data. We thank the reviewer for the improvement and for providing a clear
method to correct our previous limitation.

        I think more discussion should be give to the big O `O()` runtime of the
        implementations. It is mentioned in the Fig. 3 legend that they all
        follow a power law, but at least from the data shown Ripser seems to
        have much lower coefficients, if it is indeed the same big O growth
        rate.

***

        There are a number of benchmarks that could not be calculated due to
        insufficient ram, however the authors are only benchmarking with 16 GB
        of memory. In 2020 16 GB of memory is not a great deal for computational
        work, commodity laptops often have more memory. Amazon Web Services has
        remote machines that have memory limits up to 24 Terabytes!
        (https://aws.amazon.com/ec2/instance-types/high-memory/) and you can
        reserve machines with more modest memory (e.g. 64 GB) machines very cost
        effectively. The paper would be much stronger if machine capacity was
        not a limiting factor. I would strongly urge the authors to look into
        using more powerful infrastructure to run these benchmarks so their
        results were not so limited.

We agree that 16 GB RAM is a benchmarking limitation. We gained access to 32 GB
RAM machines at our academic institution and have collected memory benchmark
data on these. Although increased memory availability would increase the point
cloud sizes we would potentially explore, we stick to 32 GB for 2 reasons.
First, the memory data already illustrates a clear power law trend whose
coefficients are calculated and provided. There is very little "noise" in the
visualized data, which leads us to believe that increased memory would only
confirm the same trend with larger point clouds. Since the memory requirement
grows superlinearly, even doubling the available memory to 64 GB would still
restrict point cloud size to within the same order of magnitude. Second,
although machines with more memory are available cost-effectively, this is
currently a no-cost study. Acquiring and then spending funding would take time
and consume monetary resources. We believe the additions made to the memory
benchmark results provide sufficient evidence for our high-level conclusions
(e.g. relative package efficiency, growth rates for each package) with the
increased RAM used (32 GB instead of the previous 16 GB).

        ### Usability discussion

        I think a (possibly short) section comparing the usability of the two
        packages (TDA and TDAstats) would be useful. Computational efficiency
        should not be the only thing users weigh when deciding what package to
        use, and if there are major differences in the interfaces, ease of use,
        feature set, open bugs or maintainer responsiveness that should be
        discussed. If the authors feel the two packages are the same in terms
        of these aspects that should be mentioned. It might also be worth
        having a short code example of using each package.

We agree that computational efficiency should not be the only deciding factor
between packages. Although there are no major differences in the ease of use,
interface, or maintainer responsiveness between the TDA and TDAstats packages,
there is a significant different in the feature set between the two. We have
added a paragraph in the Discussion section that compares the functionality
of the two packages. We refrained from adding code examples as it risks
unnecessary redundancy - both packages are well documented and have associated
vignettes and publications with extensive sample code. We have referenced the
packages appropriately and any interested users can easily find the code via
the References section (in addition to CRAN and GitHub).

        ## Conclusion ##

        Overall I think the authors did an admirable job analyzing the
        benchmarks and in writing the paper. However the benchmarking
        methodology was flawed enough in terms of metrics, display, measurement
        of memory and memory limits that I think it should be revised and
        redone before the paper is accepted.

We thank the reviewer for their kind words and agree that the submitted
manuscript had room for improvement. We have made revisions to address the
reviewer's concerns and hope that the revised manuscript is now worthy of
acceptance.

### 2020-12-29 tdabench-paper/tdabench.Rmd

@@
   \centering
-  \includegraphics[height=1.5in]{../Figures/Final_Figures/Intro_Alpha.png}
+  \includegraphics[width=4in]{../Figures/Final_Figures/Intro_Alpha.png}
   \caption{\textbf{Basic visualization of the Alpha complex.}
@@
   \centering
-  \includegraphics[height=2in]{../Figures/Final_Figures/fig4.png}
+  \includegraphics[width=6in]{../Figures/Final_Figures/fig4.png}
   \caption{
@@
   geom_point() +
-  scale_x_continuous(trans = "log10") +
   scale_y_continuous(trans = "log10") +
@@
   \centering
-  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig6.png}
+  \includegraphics[width=6in]{../Figures/Final_Figures/fig6.png}
   \caption{

### 2021-01-05 tdabench-paper/review-response.txt
added responses
added big O stuff

@@
         acceptance.
+
+We thank the reviewer for a thorough and clear review. We address each of their
+concerns below.
 
@@
         and I think must be changed before the paper can be accepted.
-
         Fortunately there are tools the authors can use to accomplish this.
@@
         thing as `bench::bench_process_memory()`.
-
         Using the suggested approach would work for all of the benchmarks in
@@
 
-***
+We agree with the reviewer's point. To address this, we have added a faceted
+bar plot figure that illustrates the growth rate of each engine's calculation
+function with various point cloud dimensions and shapes. Additionally, text in
+the results and discussion sections has been edited/added to discuss the big O
+complexity (memory and runtime).
 

### 2021-01-05 tdabench-paper/tdabench.Rmd
added big O stuff

--- 2020-12-28 tdabench-paper/tdabench.Rmd
@@
 suppressPackageStartupMessages(library(dplyr))
+suppressPackageStartupMessages(library(tidyr))
 suppressPackageStartupMessages(library(broom))
+suppressPackageStartupMessages(library(readr))
 ```
@@
 
+\begin{figure}[b]
+  \centering
+  \includegraphics[width=4in]{../Figures/Final_Figures/Intro_Alpha.png}
+  \caption{\textbf{Basic visualization of the Alpha complex.}
+  For a given $\alpha$, $\alpha$-radius balls are drawn around each point, and the union of the balls is taken.
+  Then, an intersection between this union of $\alpha$-balls and the Vornoi diagram is taken.
+  A connecting segment is drawn between points in adjacent Voronoi cells once the $\alpha$-ball fills out the Voronoi diagram.
+  As $\alpha$ grows, more circles fill out the Voronoi cells.
+  Once $\alpha$ is large enough, the Delaunay Triangulation is formed.}
+\end{figure}
+
 In both methods, the boundary matrix records all simplicial complexes for each parameter value ($\delta$ for Vietoris-Rips complexes and $\alpha$ for alpha complexes). 
@@
 
-\begin{figure}
+\begin{figure}[tb]
   \centering
-  \includegraphics[height=1.5in]{../Figures/Final_Figures/Intro_Alpha.png}
-  \caption{\textbf{Basic visualization of the Alpha complex.}
-  For a given $\alpha$, $\alpha$-radius balls are drawn around each point, and the union of the balls is taken.
-  Then, an intersection between this union of $\alpha$-balls and the Vornoi diagram is taken.
-  A connecting segment is drawn between points in adjacent Voronoi cells once the $\alpha$-ball fills out the Voronoi diagram.
-  As $\alpha$ grows, more circles fill out the Voronoi cells.
-  Once $\alpha$ is large enough, the Delaunay Triangulation is formed.}
+  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig1.png}
+  \caption{
+    \textbf{Calculating persistent homology of a torus with three TDA libraries.}
+    Median runtime (min to max, $n=10$ iterations per data point) for each TDA library (denoted by color) is plotted against point cloud size.
+    Homological features up to 2 dimensions were calculated.
+    Time complexity follows a power law for all three libraries (see GitHub repo for regression details).
+    Although the libraries have similar runtimes for smaller point clouds, Dionysus has a clear disadvantage when the number of points exceeds 100.
+    When the number of points exceeds 200, Ripser has a clear advantage over GUDHI, which maintains its advantage over Dionysus.
+  }
+\label{fig:tor}
 \end{figure}
@@
 
+\begin{widefigure}[bt]
+  \centering
+  \includegraphics[height=2in]{../Figures/Final_Figures/fig2.png}
+  \caption{
+    \textbf{Calculating persistent homology of round point clouds of varying dimensions with three TDA libraries.}
+    Median runtime (min to max, $n=10$ iterations per data point) for each TDA library (denoted by color) is plotted against point cloud size and faceted by data dimension.
+    Left panel compares library performance for a 2-dimensional circular point cloud; center panel for a 3-dimensional spherical point cloud; and right panel for a 4-dimensional hyperspherical point cloud.
+    Maximum feature dimensions (one less than data dimension) were calculated in each case.
+  }
+\label{fig:cir}
+\end{widefigure}
+
 # Results
@@
 
-\begin{figure}
+\begin{widefigure}
   \centering
-  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig1.png}
+  \includegraphics[height=2in]{../Figures/Final_Figures/fig3.png}
   \caption{
-    \textbf{Calculating persistent homology of a torus with three TDA libraries.}
-    Median runtime (min to max, $n=10$ iterations per data point) for each TDA library (denoted by color) is plotted against point cloud size.
-    Homological features up to 2 dimensions were calculated.
-    Time complexity follows a power law for all three libraries (see GitHub repo for regression details).
-    Although the libraries have similar runtimes for smaller point clouds, Dionysus has a clear disadvantage when the number of points exceeds 100.
-    When the number of points exceeds 200, Ripser has a clear advantage over GUDHI, which maintains its advantage over Dionysus.
-  }
-\label{fig:tor}
-\end{figure}
+    \textbf{Comparison of Vietoris-Rips complex persistent homology calculation as a function of feature dimension.}
+    Median runtime (min to max, $n=10$ iterations per data point) for various point cloud sizes (denoted by color) is plotted against calculated feature dimension and faceted by TDA libary.
+    Persistent homology was calculated on a uniformly distributed random sample of points contained within a 1 unit, 8-dimensional cube.
+    Computational limitations of calculating persistent homology for a large number of feature dimensions restricted point clouds to relatively small sizes.}
+\label{fig:box}
+\end{widefigure}
 
@@
 
-\begin{widefigure}
-  \centering
-  \includegraphics[height=2in]{../Figures/Final_Figures/fig2.png}
-  \caption{
-    \textbf{Calculating persistent homology of round point clouds of varying dimensions with three TDA libraries.}
-    Median runtime (min to max, $n=10$ iterations per data point) for each TDA library (denoted by color) is plotted against point cloud size and faceted by data dimension.
-    Left panel compares library performance for a 2-dimensional circular point cloud; center panel for a 3-dimensional spherical point cloud; and right panel for a 4-dimensional hyperspherical point cloud.
-    Maximum feature dimensions (one less than data dimension) were calculated in each case.
-  }
-\label{fig:cir}
-\end{widefigure}
-
 Large data and feature dimensionality often restricts persistent homology calculations to small point clouds due to computational limits.
@@
 It is unclear whether runtime for each library grows polynomially or exponentially (see Supplement for regression details).
-
-\begin{widefigure}
-  \centering
-  \includegraphics[height=2in]{../Figures/Final_Figures/fig3.png}
-  \caption{
-    \textbf{Comparison of Vietoris-Rips complex persistent homology calculation as a function of feature dimension.}
-    Median runtime (min to max, $n=10$ iterations per data point) for various point cloud sizes (denoted by color) is plotted against calculated feature dimension and faceted by TDA libary.
-    Persistent homology was calculated on a uniformly distributed random sample of points contained within a 1 unit, 8-dimensional cube.
-    Computational limitations of calculating persistent homology for a large number of feature dimensions restricted point clouds to relatively small sizes.}
-\label{fig:box}
-\end{widefigure}
 
@@
                             2:4,
-                            c("circle", "annulus", "torus", "uniform"))
+                            c("circle", "annulus", "torus", "uniform"),
+                            stringsAsFactors = FALSE) %>%
+  filter(Var3!="torus" | Var2 == 3) %>%
+  rename(Engine = Var1,
+         Dimension = Var2,
+         Shape = Var3)
 
@@
   all_data <- all_data %>%
-    filter(NumPoints >= 100 &
+    filter(NumPoints >= 25 &
            Engine == engine &
@@
 }
+
+calc_pl_exp_summ <- function(size, value) {
+  log_pts <- log10(size)
+  log_mem <- log10(value)
+
+  lm_mod <- lm(log_mem ~ log_pts)
+  tidy_mod <- tidy(lm_mod)
+
+  return(tidy_mod$estimate[2])
+}
+
+# add calculated values to pl_exponents as an additional column
+exp_vec <- vapply(X = seq_len(nrow(pl_exponents)),
+                      FUN.VALUE = numeric(1),
+                      FUN = function(row_num) {
+                        calc_pl_exp(df_mem,
+                                    pl_exponents$Engine[row_num],
+                                    pl_exponents$Dimension[row_num],
+                                    pl_exponents$Shape[row_num])
+                      }
+                  )
+pl_exponents <- pl_exponents %>%
+  mutate(Exp = exp_vec)
 ```
@@
 In addition to runtime differences, the three Vietoris-Rips homology engines differ in memory use.
-After crossing a threshold between 50 and 100 points, all three engines appeared to follow power law growth, with a linear trend on log-log plots (Figure \ref{fig:mem}).
-However, for all combinations of point cloud dimension and shape, TDAstats used the least memory and Dionysus used the most, with TDAstats also growing with the smallest power law exponent as the number of points increased.
-As an example, for 3-dimensional annulus point clouds, GUDHI memory use grew with a power law exponent of
-`r calc_pl_exp(df_mem, "GUDHI", 3, "annulus") %>% signif(digits = 2)`,
-Dionysus use grew with an exponent of
-`r calc_pl_exp(df_mem, "Dionysus", 3, "annulus") %>% signif(digits = 2)`,
-and TDAstats use grew with an exponent of
-`r calc_pl_exp(df_mem, "TDAstats", 3, "annulus") %>% signif(digits = 2)`.
-The entire table of exponents can be found in the Supplement (GitHub), but Figure \ref{fig:mem} clearly illustrates that TDAstats was, for all tested cases, the optimal Vietoris-Rips engine in the context of memory use.
+All three engines appeared to follow power law growth, with a linear trend on log-log plots (Figure \ref{fig:mem}).
+However, for nearly all combinations of point cloud dimension and shape, TDAstats used the least memory and Dionysus used the most, with TDAstats also growing with the smallest power law exponent as the number of points increased for most point clouds.
+For most point clouds, runtime and memory complexity for TDAstats (Ripser) grew with a power function at least one degree less than the other engines (Figure \ref{fig:bigO}).
 
@@
 
+```{r bigoh, message=FALSE, warning=FALSE,echo=FALSE, fig.height=3, fig.cap="\\label{fig:bigO}\\textbf{Big-O exponents for runtime and memory complexity as point cloud size varies.} In the majority of cases, Ripser (and consequently, TDAstats) has the lowest exponent, indicating slowest growth in complexity as point cloud size increases.
+Due to shape constraints, torus only has data available in the third dimension."}
+# collect data for runtime
+get_runtime_data <- function(shape) {
+  filepath <- paste0("../Cluster_data/",
+                     shape,
+                     ".csv")
+
+  # read and transform data
+  read_csv(filepath) %>%
+    select(-c(X1, V1)) %>%
+    rename(Shape = V2,
+           Dimension = V3,
+           Size = V4,
+           Calc_Dim = V5,
+           Engine = V6) %>%
+    filter(Calc_Dim == Dimension - 1) %>%
+    filter(Engine != "GUDHIalpha") %>%
+    pivot_longer(cols = paste0("V", 7:16)) %>%
+    select(-c(name, Calc_Dim)) %>%
+    group_by(Shape, Dimension, Size, Engine) %>%
+    summarise(Value = median(value)) %>%
+    mutate(Engine = gsub(x = Engine,
+                         pattern = "stats",
+                         replacement = "TDAstats")) %>%
+    rename(DataType = Shape)
+}
+
+# collect data
+annulus <- get_runtime_data("annulus")
+circle <- get_runtime_data("circle")
+torus <- get_runtime_data("torus")
+uniform <- get_runtime_data("uniform")
+
+runtime_data <- rbind(annulus, circle, torus, uniform) %>%
+  filter(Dimension %in% 2:4) %>%
+  group_by(Engine, Dimension, DataType) %>%
+  summarise(Exp = calc_pl_exp_summ(Size, Value)) %>%
+  mutate(Type = "runtime")
+
+memory_data <- pl_exponents %>%
+  rename(DataType = Shape) %>%
+  mutate(Type = "memory")
+
+combo_data <- rbind(runtime_data, memory_data) %>%
+  mutate(Engine = gsub(Engine, pattern = "TDAstats", replacement = "Ripser"))
+
+# plot big O exponents for memory
+ggplot(combo_data, aes(x = Dimension, y = Exp, fill = Engine)) +
+  facet_grid(Type ~ DataType) +
+  geom_bar(stat = "identity", position = "dodge") +
+  ylab("Big-O exponent") +
+  scale_y_continuous() +
+  scale_fill_manual(values = c("red", "blue", "brown")) +
+  theme_bw()
+```
+
 # Discussion
@@
 For the conditions in our paper, Vietoris-Rips and alpha complexes both performed better than their theoretical maximums.
-Vietoris-Rips complex calculations consistently had a polynomial growth for both runtime and boundary matrix size while alpha complexes had linear growth.
+Vietoris-Rips complex calculations consistently had a polynomial growth for both runtime and memory, while alpha complexes had linear runtime growth.
 
@@
 While Vietoris-Rips complexes can handle high-dimensional data well, calculation still significantly slows down when looking for higher dimension features.
-Boundary matrix size rapidly increases and can overwhelm available RAM even for small numbers of points. 
-Looking for high-dimensional features in a large, high-dimensional point clouds remains a challenge. 
+This is evidenced by the big-O polynomial growth for runtime and memory that have degree less than 4 for most 2-dimensional point clouds, but degree between 4 and 6 for most 4-dimensional point clouds (Figure \ref{fig:bigO}); even higher degree complexities should be expected as point cloud dimension increases.
+Thus, finding high-dimensional topological features in high-dimensional point clouds remains a challenge. 
 Methods to calculate persistent homology do exist for other simplicial complexes, such as the Delaunay complex and the Witness complex, but, to our knowledge, they are not currently implemented in CRAN packages.
@@
 The authors thank Emily Dolson, PhD for assistance with using a high-performance computing node for data collection.
+JGS was supported by NIH grant R37 CA244613.
 

### 2021-01-12 tdabench-paper/review-response.txt
updated reviewer response

--- 2020-12-29 tdabench-paper/review-response.txt
@@
 in the results and discussion sections has been adjusted to reflect the newly
-added data. We thank the reviewer for the improvement and for providing a clear
+added data.
+        Examples of text revisions/additions
+        - Results: "In addition to runtime differences...than the other engines"
+        - Discussion: "While Vietoris-Rips complexes can...dimension increases."
+We thank the reviewer for the improvement and for providing a clear
 method to correct our previous limitation.
@@
 
-***
+We agree with the reviewer's point of discussing the big O runtime of the
+various persistent homology engines. However, instead of focusing on the
+coefficients, we added results and discussion pertaining to the exponent of the
+power law (matters more as dimension increases). These additions include Figure
+8 in the results (and corresponding text) and text in the last paragraph of the
+discussion.
 
@@
 (e.g. relative package efficiency, growth rates for each package) with the
-increased RAM used (32 GB instead of the previous 16 GB).
+increased RAM used.
 
@@
 there is a significant different in the feature set between the two. We have
-added a paragraph in the Discussion section that compares the functionality
-of the two packages. We refrained from adding code examples as it risks
-unnecessary redundancy - both packages are well documented and have associated
-vignettes and publications with extensive sample code. We have referenced the
-packages appropriately and any interested users can easily find the code via
-the References section (in addition to CRAN and GitHub).
+added the following paragraph in the Discussion section that compares the
+functionality of the two packages:
+        "On average, Ripser computed...packages are hosted on CRAN."
+
+We refrained from adding code examples as it risks unnecessary redundancy - both
+packages are well documented and have associated vignettes and publications with
+extensive sample code. We have referenced the packages appropriately and any
+interested users can easily find the code via the References section (in
+addition to CRAN and GitHub).
 

### 2021-01-13 tdabench-paper/review-response.txt
very minor

@@
 coefficients, we added results and discussion pertaining to the exponent of the
-power law (matters more as dimension increases). These additions include Figure
-8 in the results (and corresponding text) and text in the last paragraph of the
-discussion.
+power law (which matters more as dimension increases). To address this, we
+have added a faceted bar plot figure that illustrates the growth rate of each
+engine's calculation function with various point cloud dimensions and shapes.
+Additionally, text in the results and discussion sections has been edited/added
+to discuss the big O complexity (memory and runtime, last paragraph in dicussion).
 
